# quaternion

## TaylorCatAlice

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1PV411P7w4&autoplay=0"></iframe>

<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1qa411B792&autoplay=0"></iframe>
:::

https://en.wikipedia.org/wiki/Blackboard_bold

### complex / dionion / bionion

$$
\begin{aligned}
 & c=a+b\mathrm{i}=a+\mathrm{i}b,\begin{cases}
c\in\mathbb{C}\\
\mathrm{i}^{2}=-1\\
a,b\in\mathbb{R} & \Leftrightarrow\left\langle a,b\right\rangle \in\mathbb{R}^{2}
\end{cases}\\
= & z=x+y\mathrm{i}=x+\mathrm{i}y,\begin{cases}
z\in\mathbb{C}\\
\mathrm{i}^{2}=-1\\
x,y\in\mathbb{R} & \Leftrightarrow\left\langle x,y\right\rangle \in\mathbb{R}^{2}
\end{cases}\\
= & \sqrt{x^{2}+y^{2}}\left(\dfrac{x}{\sqrt{x^{2}+y^{2}}}+\dfrac{y}{\sqrt{x^{2}+y^{2}}}\mathrm{i}\right)=r\left(\cos\theta+\mathrm{i}\sin\theta\right)=r\mathrm{e}^{\mathrm{i}\theta}
\end{aligned}
$$

Also see [complex group representation]^[\@ref(complex-group-representation)]^.

### trionion / triernion / triplex / ternion

https://zh.wikipedia.org/zh-tw/%E4%B8%89%E5%85%83%E6%95%B8

https://math.stackexchange.com/questions/1784166/why-are-there-no-triernions-3-dimensional-analogue-of-complex-numbers-quate

https://math.stackexchange.com/questions/32100/is-there-a-third-dimension-of-numbers/4453131

$$
\begin{aligned}
 & t=a+b\mathrm{i}+c\mathrm{j}=a+\mathrm{i}b+\mathrm{j}c,\begin{cases}
t\in\mathbb{T}\\
\mathrm{i}^{2}=-1\\
\mathrm{j}^{2}=-1
\end{cases}\\
= & w=x+y\mathrm{i}+z\mathrm{j}=x+\mathrm{i}y+\mathrm{j}z,\begin{cases}
w\in\mathbb{T}\\
\mathrm{i}^{2}=-1\\
\mathrm{j}^{2}=-1
\end{cases}\\
= & \sqrt{x^{2}+y^{2}+z^{2}}\left(\dfrac{x}{\sqrt{x^{2}+y^{2}+z^{2}}}+\dfrac{y}{\sqrt{x^{2}+y^{2}+z^{2}}}\mathrm{i}+\dfrac{z}{\sqrt{x^{2}+y^{2}+z^{2}}}\mathrm{j}\right)=?
\end{aligned}
$$

***

$$
\begin{cases}
A\left(BC\right)=\left(AB\right)C & \left(a\right)\text{associativity}\\
A\left(B+C\right)=AB+AC & \left(d\right)\text{distributivity}
\end{cases}
$$

$$
\begin{aligned}
\mathbb{T}\ni\mathrm{i}\mathrm{j}= & X+Y\mathrm{i}+Z\mathrm{j}\in\mathbb{T}\\
-\mathrm{j}=\left(\mathrm{i}^{2}\right)\mathrm{j}\overset{\left(a\right)}{=}\mathrm{i}\left(\mathrm{i}\mathrm{j}\right)= & \mathrm{i}\left(X+Y\mathrm{i}+Z\mathrm{j}\right)\overset{\left(d\right)}{=}-Y+X\mathrm{i}+Z\mathrm{i}\mathrm{j}\\
\mathrm{i}\mathrm{j}= & \dfrac{Y}{Z}-\dfrac{X}{Z}\mathrm{i}-\dfrac{1}{Z}\mathrm{j}\Rightarrow\begin{cases}
X=\dfrac{Y}{Z}\\
Y=-\dfrac{X}{Z}\\
Z=-\dfrac{1}{Z} & \Rightarrow Z^{2}=-1\Rightarrow Z\notin\mathbb{R}
\end{cases}\\
-\mathrm{i}=\mathrm{i}\left(\mathrm{j}^{2}\right)\overset{\left(a\right)}{=}\left(\mathrm{i}\mathrm{j}\right)\mathrm{j}= & \left(X+Y\mathrm{i}+Z\mathrm{j}\right)\mathrm{j}\overset{\left(d\right)}{=}-Z+X\mathrm{j}+Y\mathrm{i}\mathrm{j}\\
\mathrm{i}\mathrm{j}= & \dfrac{Z}{Y}-\dfrac{1}{Y}\mathrm{i}-\dfrac{X}{Y}\mathrm{j}\Rightarrow\begin{cases}
X=\dfrac{Z}{Y}\\
Y=-\dfrac{1}{Y} & \Rightarrow Y^{2}=-1\Rightarrow Y\notin\mathbb{R}\\
Z=-\dfrac{X}{Y}
\end{cases}
\end{aligned}
$$

### quaternion

https://en.wikipedia.org/wiki/Quaternion

$$
\begin{aligned}
 & q=a+b\mathrm{i}+c\mathrm{j}+d\mathrm{k}=a+\mathrm{i}b+\mathrm{j}c+\mathrm{k}d,\begin{cases}
q\in\mathbb{H}\\
a,b,c,d\in\mathbb{R} & \Leftrightarrow\left\langle a,b,c,d\right\rangle \in\mathbb{R}^{4}
\end{cases}\\
= & w=t+x\mathrm{i}+y\mathrm{j}+z\mathrm{k}=t+\mathrm{i}x+\mathrm{j}y+\mathrm{k}z,\begin{cases}
w\in\mathbb{H}\\
t,x,y,z\in\mathbb{R} & \Leftrightarrow\left\langle t,x,y,z\right\rangle \in\mathbb{R}^{4}
\end{cases}\\
= & ?
\end{aligned}
$$

***

The quaternion set is denoted $\mathbb{H}$ for Sir R.W. **Hamilton**, because he suddenly and strikingly realized

$$
\begin{cases}
\mathrm{i}\mathrm{j}=\mathrm{k}\\
\mathrm{k}\in\mathbb{H}
\end{cases}\Rightarrow\mathrm{i}\mathrm{j}\in\mathbb{H}\text{ for closure property}
$$

***

for the sake of rigorosity, see [group theory]^[\@ref(group-theory)]^

$$
\mathrm{k}^{2}=-1
$$

$$
\begin{aligned}
\mathrm{i}\mathrm{j} & =\mathrm{k}\\
\mathrm{i}\mathrm{j}\mathrm{k}=\mathrm{i}\left(\mathrm{j}\mathrm{k}\right)\overset{\left(a\right)}{=}\left(\mathrm{i}\mathrm{j}\right)\mathrm{k} & =\mathrm{k}\mathrm{k}=\mathrm{k}^{2}=-1\\
\mathrm{k}\mathrm{i}\mathrm{j}=\left(\mathrm{k}\mathrm{i}\right)\mathrm{j}\overset{\left(a\right)}{=}\mathrm{k}\left(\mathrm{i}\mathrm{j}\right) & =\mathrm{k}\mathrm{k}=\mathrm{k}^{2}=-1
\end{aligned}
$$

$$
\begin{aligned}
\mathrm{i}\mathrm{j} & =\mathrm{k}\\
-\mathrm{j}=\left(\mathrm{i}^{2}\right)\mathrm{j}\overset{\left(a\right)}{=}\mathrm{i}\left(\mathrm{i}\mathrm{j}\right) & =\mathrm{i}\mathrm{k}\\
-\mathrm{i}=\mathrm{i}\left(\mathrm{j}^{2}\right)\overset{\left(a\right)}{=}\left(\mathrm{i}\mathrm{j}\right)\mathrm{j} & =\mathrm{k}\mathrm{j}
\end{aligned}
$$

$$
\begin{aligned}
-\mathrm{j}=\left(\mathrm{i}^{2}\right)\mathrm{j}\overset{\left(a\right)}{=}\mathrm{i}\left(\mathrm{i}\mathrm{j}\right) & =\mathrm{i}\mathrm{k}\\
1=-\mathrm{j}^{2}=\mathrm{j}\left(-\mathrm{j}\right) & =\mathrm{j}\left(\mathrm{i}\mathrm{k}\right)\overset{\left(a\right)}{=}\left(\mathrm{j}\mathrm{i}\right)\mathrm{k}\\
\mathrm{k}=\left[1\right]\mathrm{k} & =\left[\left(\mathrm{j}\mathrm{i}\right)\mathrm{k}\right]\mathrm{k}\overset{\left(a\right)}{=}\left(\mathrm{j}\mathrm{i}\right)\left(\mathrm{k}^{2}\right)=\left(\mathrm{j}\mathrm{i}\right)\left(-1\right)\\
-\mathrm{k} & =\mathrm{j}\mathrm{i}\\
-\mathrm{i}=\mathrm{i}\left(\mathrm{j}^{2}\right)\overset{\left(a\right)}{=}\left(\mathrm{i}\mathrm{j}\right)\mathrm{j} & =\mathrm{k}\mathrm{j}\\
1=\left(-\mathrm{i}\right)\mathrm{i} & =\left(\mathrm{k}\mathrm{j}\right)\mathrm{i}\overset{\left(a\right)}{=}\mathrm{k}\left(\mathrm{j}\mathrm{i}\right)=\mathrm{k}\mathrm{j}\mathrm{i}\\
1 & =\mathrm{k}\mathrm{j}\mathrm{i}
\end{aligned}
$$

There is no more [commutativity]^[\@ref(commutativity)]^, i.e.

$$
AB\not\equiv BA
$$

but

$$
AB+BA=0\Leftrightarrow AB=-BA
$$
satisfying [anticommutativity]^[\@ref(anticommutativity)]^.

$$
\begin{aligned}
\begin{cases}
\mathrm{i}\mathrm{j}=\mathrm{k}\\
\mathrm{j}\mathrm{i}=-\mathrm{k}
\end{cases}\Leftrightarrow & \mathrm{j}\mathrm{i}=-\mathrm{k}=-\mathrm{i}\mathrm{j}\\
\Rightarrow & \mathrm{j}\mathrm{i}=-\mathrm{i}\mathrm{j}\\
\Leftrightarrow & \mathrm{i}\mathrm{j}+\mathrm{j}\mathrm{i}=0
\end{aligned}
$$

$$
\begin{aligned}
\begin{cases}
\mathrm{k}\mathrm{i}\mathrm{j}=-1\\
\mathrm{k}\mathrm{j}\mathrm{i}=1
\end{cases}\Leftrightarrow & \mathrm{k}\mathrm{i}\mathrm{j}=-1=-\mathrm{k}\mathrm{j}\mathrm{i}\\
\Rightarrow & \mathrm{k}\mathrm{i}\mathrm{j}=-\mathrm{k}\mathrm{j}\mathrm{i}\\
\Leftrightarrow & \mathrm{k}\mathrm{i}\mathrm{j}+\mathrm{k}\mathrm{j}\mathrm{i}=0
\end{aligned}
$$

***

$$
\begin{aligned}
 & q=a+b\mathrm{i}+c\mathrm{j}+d\mathrm{k}=a+\mathrm{i}b+\mathrm{j}c+\mathrm{k}d,\begin{cases}
q\in\mathbb{H}\\
\mathrm{i}^{2}=\mathrm{j}^{2}=\mathrm{k}^{2}=-1=\mathrm{i}\mathrm{j}\mathrm{k} & \Rightarrow\mathrm{i}\mathrm{j}=\mathrm{k}\\
a,b,c,d\in\mathbb{R} & \Leftrightarrow\left\langle a,b,c,d\right\rangle \in\mathbb{R}^{4}
\end{cases}\\
= & w=t+x\mathrm{i}+y\mathrm{j}+z\mathrm{k}=t+\mathrm{i}x+\mathrm{j}y+\mathrm{k}z,\begin{cases}
w\in\mathbb{H}\\
\mathrm{i}^{2}=\mathrm{j}^{2}=\mathrm{k}^{2}=-1=\mathrm{i}\mathrm{j}\mathrm{k} & \Rightarrow\mathrm{i}\mathrm{j}=\mathrm{k}\\
t,x,y,z\in\mathbb{R} & \Leftrightarrow\left\langle t,x,y,z\right\rangle \in\mathbb{R}^{4}
\end{cases}\\
= & t+\begin{pmatrix}\mathrm{i} & \mathrm{j} & \mathrm{k}\end{pmatrix}\begin{pmatrix}x\\
y\\
z
\end{pmatrix}=x_{{\scriptscriptstyle 0}}+\begin{pmatrix}\boldsymbol{i} & \boldsymbol{j} & \boldsymbol{k}\end{pmatrix}\begin{pmatrix}x\\
y\\
z
\end{pmatrix}=x_{{\scriptscriptstyle 0}}+\begin{pmatrix}\boldsymbol{e}_{{\scriptscriptstyle 1}} & \boldsymbol{e}_{{\scriptscriptstyle 2}} & \boldsymbol{e}_{{\scriptscriptstyle 3}}\end{pmatrix}\begin{pmatrix}x_{{\scriptscriptstyle 1}}\\
x_{{\scriptscriptstyle 2}}\\
x_{{\scriptscriptstyle 3}}
\end{pmatrix}=x_{{\scriptscriptstyle 0}}+\boldsymbol{x},\begin{cases}
\boldsymbol{e}_{{\scriptscriptstyle 1}}=\boldsymbol{i}=\mathrm{i}\\
\boldsymbol{e}_{{\scriptscriptstyle 2}}=\boldsymbol{j}=\mathrm{j}\\
\boldsymbol{e}_{{\scriptscriptstyle 3}}=\boldsymbol{k}=\mathrm{k}
\end{cases}
\end{aligned}
$$
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '40%', fig.cap='quaternion multiplication table'}
\begin{tikzpicture}
  \node {
    \begin{tabular}{|c|c|c|c|c|}
    \hline 
    $\cdot$ & $1$ & $\mathrm{i}$ & $\mathrm{j}$ & $\mathrm{k}$\tabularnewline
    \hline 
    $1$ & $1$ & $\mathrm{i}$ & $\mathrm{j}$ & $\mathrm{k}$\tabularnewline
    \hline 
    $\mathrm{i}$ & $\mathrm{i}$ & $-1$ & $\mathrm{k}$ & $-\mathrm{j}$\tabularnewline
    \hline 
    $\mathrm{j}$ & $\mathrm{j}$ & $-\mathrm{\mathrm{k}}$ & $-1$ & $\mathrm{i}$\tabularnewline
    \hline 
    $\mathrm{k}$ & $\mathrm{k}$ & $\mathrm{j}$ & $-\mathrm{i}$ & $-1$\tabularnewline
    \hline 
    \end{tabular}
  };
\end{tikzpicture}
```

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '40%', fig.cap='quaternion basis group table'}
\begin{tikzpicture}
  \node {
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
    $\cdot$ & $1$ & $\mathrm{i}$ & $\mathrm{j}$ & $\mathrm{k}$ & $-1$ & $-\mathrm{i}$ & $-\mathrm{j}$ & $-\mathrm{k}$\tabularnewline
    \hline 
    $1$ & $1$ & $\mathrm{i}$ & $\mathrm{j}$ & $\mathrm{k}$ &  &  &  & \tabularnewline
    \hline 
    $\mathrm{i}$ & $\mathrm{i}$ & $-1$ & $\mathrm{k}$ & $-\mathrm{j}$ &  &  &  & \tabularnewline
    \hline 
    $\mathrm{j}$ & $\mathrm{j}$ & $-\mathrm{\mathrm{k}}$ & $-1$ & $\mathrm{i}$ &  &  &  & \tabularnewline
    \hline 
    $\mathrm{k}$ & $\mathrm{k}$ & $\mathrm{j}$ & $-\mathrm{i}$ & $-1$ &  &  &  & \tabularnewline
    \hline 
    $-1$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{i}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{j}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{k}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    \end{tabular}
  };
\end{tikzpicture}
```

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '40%', fig.cap='quaternion basis group table 2'}
\begin{tikzpicture}
  \node {
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
    $\cdot$ & $1$ & $-1$ & $\mathrm{i}$ & $-\mathrm{i}$ & $\mathrm{j}$ & $-\mathrm{j}$ & $\mathrm{k}$ & $-\mathrm{k}$\tabularnewline
    \hline 
    $1$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-1$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $\mathrm{i}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{i}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $\mathrm{j}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{j}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $\mathrm{k}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    $-\mathrm{k}$ &  &  &  &  &  &  &  & \tabularnewline
    \hline 
    \end{tabular}
  };
\end{tikzpicture}
```

#### true origin of ( dot product & cross product ) / ( inner product & outer product )

product of two pure imaginary quaternions

$$
\begin{aligned}
\boldsymbol{x}_{{\scriptscriptstyle 1}}\boldsymbol{x}_{{\scriptscriptstyle 2}}= & \left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}\mathrm{i}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}\mathrm{j}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}\mathrm{k}\right)\left(x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\mathrm{i}+x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\mathrm{j}+x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\mathrm{k}\right)\\
= & x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\mathrm{i}^{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\mathrm{i}\mathrm{j}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\mathrm{i}\mathrm{k}\\
+ & x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\mathrm{j}\mathrm{i}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\mathrm{j}^{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\mathrm{j}\mathrm{k}\\
+ & x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\mathrm{k}\mathrm{i}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\mathrm{k}\mathrm{j}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\mathrm{k}^{{\scriptscriptstyle 2}}\\
= & -\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\right)\\
+ & \left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\right)\mathrm{j}\mathrm{k}+\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\right)\mathrm{k}\mathrm{i}+\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\right)\mathrm{i}\mathrm{j}\\
= & -\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\right)\\
+ & \left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}\right)\mathrm{i}+\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\right)\mathrm{j}+\left(x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}-x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}\right)\mathrm{k}\\
= & -\left(\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}\right)+\left(\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}\right),\begin{cases}
\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}=x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}\\
\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}=\begin{vmatrix}\mathrm{i} & \mathrm{j} & \mathrm{k}\\
x_{{\scriptscriptstyle 1}{\scriptscriptstyle 1}} & x_{{\scriptscriptstyle 1}{\scriptscriptstyle 2}} & x_{{\scriptscriptstyle 1}{\scriptscriptstyle 3}}\\
x_{{\scriptscriptstyle 2}{\scriptscriptstyle 1}} & x_{{\scriptscriptstyle 2}{\scriptscriptstyle 2}} & x_{{\scriptscriptstyle 2}{\scriptscriptstyle 3}}
\end{vmatrix}
\end{cases}\\
= & -\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}
\end{aligned}
$$

product of two general quaternions / ordinary quaternions

$$
\begin{aligned}
q_{{\scriptscriptstyle 1}}q_{{\scriptscriptstyle 2}}= & \left(q_{{\scriptscriptstyle 10}}+q_{{\scriptscriptstyle 11}}\mathrm{i}+q_{{\scriptscriptstyle 12}}\mathrm{j}+q_{{\scriptscriptstyle 13}}\mathrm{k}\right)\left(q_{{\scriptscriptstyle 20}}+q_{{\scriptscriptstyle 21}}\mathrm{i}+q_{{\scriptscriptstyle 22}}\mathrm{j}+q_{{\scriptscriptstyle 23}}\mathrm{k}\right)\\
= & \left(x_{{\scriptscriptstyle 10}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\right)\left(x_{{\scriptscriptstyle 20}}+\boldsymbol{x}_{{\scriptscriptstyle 2}}\right),\begin{cases}
x_{{\scriptscriptstyle i\mu}}=q_{{\scriptscriptstyle i\mu}} & \mu\in\left\{ 0\right\} \cup\left(\mathbb{N}\cap\left[1,3\right]\right)\\
\boldsymbol{x}_{{\scriptscriptstyle i}}=x_{{\scriptscriptstyle ij}}\boldsymbol{e}_{{\scriptscriptstyle j}} & i,j\in\mathbb{N}\cap\left[1,3\right],\begin{cases}
\boldsymbol{e}_{{\scriptscriptstyle 1}}=\mathrm{i}\\
\boldsymbol{e}_{{\scriptscriptstyle 2}}=\mathrm{j}\\
\boldsymbol{e}_{{\scriptscriptstyle 3}}=\mathrm{k}
\end{cases}
\end{cases}\\
= & x_{{\scriptscriptstyle 10}}x_{{\scriptscriptstyle 20}}+x_{{\scriptscriptstyle 10}}\boldsymbol{x}_{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 20}}\boldsymbol{x}_{{\scriptscriptstyle 1}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\boldsymbol{x}_{{\scriptscriptstyle 2}}\\
= & x_{{\scriptscriptstyle 10}}x_{{\scriptscriptstyle 20}}+x_{{\scriptscriptstyle 10}}\boldsymbol{x}_{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 20}}\boldsymbol{x}_{{\scriptscriptstyle 1}}-\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}\\
= & \left(x_{{\scriptscriptstyle 10}}x_{{\scriptscriptstyle 20}}-\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}\right)+\left(x_{{\scriptscriptstyle 10}}\boldsymbol{x}_{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 20}}\boldsymbol{x}_{{\scriptscriptstyle 1}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}\right)\\
x_{{\scriptscriptstyle 10}}x_{{\scriptscriptstyle 20}}-\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}= & \begin{pmatrix}q_{{\scriptscriptstyle 10}} & q_{{\scriptscriptstyle 11}} & q_{{\scriptscriptstyle 12}} & q_{{\scriptscriptstyle 13}}\end{pmatrix}\begin{pmatrix}1\\
 & -1\\
 &  & -1\\
 &  &  & -1
\end{pmatrix}\begin{pmatrix}q_{{\scriptscriptstyle 20}}\\
q_{{\scriptscriptstyle 21}}\\
q_{{\scriptscriptstyle 22}}\\
q_{{\scriptscriptstyle 23}}
\end{pmatrix}=q_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle \mu}}\eta_{{\scriptscriptstyle \mu\nu}}q_{{\scriptscriptstyle 2}}^{{\scriptscriptstyle \nu}}\\
= & \begin{pmatrix}q_{{\scriptscriptstyle 10}} & q_{{\scriptscriptstyle 11}} & q_{{\scriptscriptstyle 12}} & q_{{\scriptscriptstyle 13}}\end{pmatrix}\begin{pmatrix}1 & 0 & 0 & 0\\
0 & -1 & 0 & 0\\
0 & 0 & -1 & 0\\
0 & 0 & 0 & -1
\end{pmatrix}\begin{pmatrix}q_{{\scriptscriptstyle 20}}\\
q_{{\scriptscriptstyle 21}}\\
q_{{\scriptscriptstyle 22}}\\
q_{{\scriptscriptstyle 23}}
\end{pmatrix}=\boldsymbol{q}_{{\scriptscriptstyle 1}}^{\intercal}H\boldsymbol{q}_{{\scriptscriptstyle 2}},H=\left[\eta_{{\scriptscriptstyle \mu\nu}}\right]_{4\times4}=\eta_{{\scriptscriptstyle \mu\nu}}
\end{aligned}
$$

$$
\begin{aligned}
q_{{\scriptscriptstyle 1}}q_{{\scriptscriptstyle 2}}= & \left(x_{{\scriptscriptstyle 10}}x_{{\scriptscriptstyle 20}}-\boldsymbol{x}_{{\scriptscriptstyle 1}}\cdot\boldsymbol{x}_{{\scriptscriptstyle 2}}\right)+\left(x_{{\scriptscriptstyle 10}}\boldsymbol{x}_{{\scriptscriptstyle 2}}+x_{{\scriptscriptstyle 20}}\boldsymbol{x}_{{\scriptscriptstyle 1}}+\boldsymbol{x}_{{\scriptscriptstyle 1}}\times\boldsymbol{x}_{{\scriptscriptstyle 2}}\right)\\
QP= & \left(Q_{{\scriptscriptstyle 0}}P_{{\scriptscriptstyle 0}}-\boldsymbol{Q}\cdot\boldsymbol{P}\right)+\left(Q_{{\scriptscriptstyle 0}}\boldsymbol{P}+P_{{\scriptscriptstyle 0}}\boldsymbol{Q}+\boldsymbol{Q}\times\boldsymbol{P}\right)
\end{aligned}
$$

$$
ab=\left(a_{{\scriptscriptstyle 0}}b_{{\scriptscriptstyle 0}}-\boldsymbol{a}\cdot\boldsymbol{b}\right)+\left(a_{{\scriptscriptstyle 0}}\boldsymbol{b}+b_{{\scriptscriptstyle 0}}\boldsymbol{a}+\boldsymbol{a}\times\boldsymbol{b}\right)
$$

Minkowski metric tensor

$$
\eta=H=\left[\eta_{{\scriptscriptstyle \mu\nu}}\right]_{4\times4}=\begin{pmatrix}1 & 0 & 0 & 0\\
0 & -1 & 0 & 0\\
0 & 0 & -1 & 0\\
0 & 0 & 0 & -1
\end{pmatrix}=\begin{pmatrix}1\\
 & -1\\
 &  & -1\\
 &  &  & -1
\end{pmatrix}=\eta_{{\scriptscriptstyle \mu\nu}}
$$

and quaternions as 4-vectors or four-vectors

$$
\boldsymbol{q}_{{\scriptscriptstyle 1}}^{\intercal}=\begin{pmatrix}q_{{\scriptscriptstyle 10}} & q_{{\scriptscriptstyle 11}} & q_{{\scriptscriptstyle 12}} & q_{{\scriptscriptstyle 13}}\end{pmatrix}
$$

$$
\boldsymbol{q}_{{\scriptscriptstyle 2}}=\begin{pmatrix}q_{{\scriptscriptstyle 20}}\\
q_{{\scriptscriptstyle 21}}\\
q_{{\scriptscriptstyle 22}}\\
q_{{\scriptscriptstyle 23}}
\end{pmatrix}
$$

#### commutativity vs. anticommutativity

##### commutativity

::: {show-in="html"}
交換律 = 交換性 = 對易性
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
交換律 = 交換性 = 對易性
\end{CJK}
:::

$$
AB=BA\Leftrightarrow AB-BA=0
$$

$$
AB=BA\Rightarrow AB\equiv BA
$$

##### anticommutativity

::: {show-in="html"}
反交換律 = 反交換性 = 反對易性
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
反交換律 = 反交換性 = 反對易性
\end{CJK}
:::

$$
AB+BA=0\Leftrightarrow AB=-BA
$$

$$
AB=-BA\Rightarrow AB\not\equiv BA
$$

#### bracket

##### self-invented bracket

::: {show-in="html"}
自創括號 = 自創括
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
自創括號 = 自創括
\end{CJK}
:::

###### commutative bracket

::: {show-in="html"}
交換括號 = 對易式
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
交換括號 = 對易式
\end{CJK}
:::

$$
\left[X,Y\right]=\dfrac{XY-YX}{2}
$$

###### anticommutative bracket

::: {show-in="html"}
反交換括號 = 反對易式
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
反交換括號 = 反對易式
\end{CJK}
:::

$$
\left\{ X,Y\right\} =\dfrac{XY+YX}{2}
$$

##### Poisson bracket

https://en.wikipedia.org/wiki/Poisson_bracket

##### Lagrange bracket

##### Lie bracket

#### triple product

product = double product

$$
ab=\left(a_{{\scriptscriptstyle 0}}b_{{\scriptscriptstyle 0}}-\boldsymbol{a}\cdot\boldsymbol{b}\right)+\left(a_{{\scriptscriptstyle 0}}\boldsymbol{b}+b_{{\scriptscriptstyle 0}}\boldsymbol{a}+\boldsymbol{a}\times\boldsymbol{b}\right)
$$
pure imaginary

$$
\begin{aligned}
ab= & \left(a_{{\scriptscriptstyle 0}}b_{{\scriptscriptstyle 0}}-\boldsymbol{a}\cdot\boldsymbol{b}\right)+\left(a_{{\scriptscriptstyle 0}}\boldsymbol{b}+b_{{\scriptscriptstyle 0}}\boldsymbol{a}+\boldsymbol{a}\times\boldsymbol{b}\right)\\
\overset{\begin{cases}
a_{0}=0\\
b_{0}=0
\end{cases}}{=} & \left(00-\boldsymbol{a}\cdot\boldsymbol{b}\right)+\left(0\boldsymbol{b}+0\boldsymbol{a}+\boldsymbol{a}\times\boldsymbol{b}\right)\\
= & -\boldsymbol{a}\cdot\boldsymbol{b}+\boldsymbol{a}\times\boldsymbol{b}
\end{aligned}
$$

pure imaginary product can get both ( real & imaginary ) / ( scalar & vector ) parts

$$
ab=\boldsymbol{a}\boldsymbol{b}=-\boldsymbol{a}\cdot\boldsymbol{b}+\boldsymbol{a}\times\boldsymbol{b},\begin{cases}
a=0+\boldsymbol{a}=\boldsymbol{a}\\
b=0+\boldsymbol{b}=\boldsymbol{b}
\end{cases}
$$

***

triple product

https://en.wikipedia.org/wiki/Triple_product

pure imaginary

$$
\begin{cases}
a=0+\boldsymbol{a}=\boldsymbol{a}\\
b=0+\boldsymbol{b}=\boldsymbol{b}\\
c=0+\boldsymbol{c}=\boldsymbol{c}
\end{cases}
$$

$$
\begin{aligned}
abc=\boldsymbol{a}\boldsymbol{b}\boldsymbol{c}= & \left(\boldsymbol{a}\boldsymbol{b}\right)\boldsymbol{c}=\left(-\boldsymbol{a}\cdot\boldsymbol{b}+\boldsymbol{a}\times\boldsymbol{b}\right)\boldsymbol{c}\\
= & \boldsymbol{a}\left(\boldsymbol{b}\boldsymbol{c}\right)=\boldsymbol{a}\left(-\boldsymbol{b}\cdot\boldsymbol{c}+\boldsymbol{b}\times\boldsymbol{c}\right)
\end{aligned}
$$

$$
\begin{aligned}
\boldsymbol{a}\boldsymbol{b}\boldsymbol{c}=\left(\boldsymbol{a}\boldsymbol{b}\right)\boldsymbol{c}= & \left(-\boldsymbol{a}\cdot\boldsymbol{b}+\boldsymbol{a}\times\boldsymbol{b}\right)\boldsymbol{c}\\
= & -\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}+\left(\boldsymbol{a}\times\boldsymbol{b}\right)\boldsymbol{c}\\
= & -\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}+\left(-\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}+\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}\right)\\
= & \left[-\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}\right]+\left[\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}-\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}\right]\\
=\boldsymbol{a}\left(\boldsymbol{b}\boldsymbol{c}\right)= & \boldsymbol{a}\left(-\boldsymbol{b}\cdot\boldsymbol{c}+\boldsymbol{b}\times\boldsymbol{c}\right)\\
= & -\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)+\boldsymbol{a}\left(\boldsymbol{b}\times\boldsymbol{c}\right)\\
= & -\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)+\left(-\boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right)+\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)\right)\\
= & \left[-\boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right)\right]+\left[\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)-\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)\right]
\end{aligned}
$$

by comparing ( real & imaginary ) / ( scalar & vector ) parts,

$$
\begin{aligned}
 & \begin{cases}
-\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}=-\boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right)\\
\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}-\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}=\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)-\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)
\end{cases}\\
\Rightarrow & \begin{cases}
\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}=\boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right) & \left(s\right)\\
\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}-\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}=\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)-\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right) & \left(v\right)
\end{cases}
\end{aligned}
$$

***

permutation

$$
\sigma=\begin{pmatrix}x_{{\scriptscriptstyle 1}} & x_{{\scriptscriptstyle 2}} & \cdots\\
\sigma\left(x_{{\scriptscriptstyle 1}}\right) & \sigma\left(x_{{\scriptscriptstyle 2}}\right) & \cdots
\end{pmatrix}
$$

$$
\begin{pmatrix}a & b & c\\
a & b & c
\end{pmatrix},\begin{pmatrix}a & b & c\\
b & c & a
\end{pmatrix},\begin{pmatrix}a & b & c\\
c & a & b
\end{pmatrix},\begin{pmatrix}a & b & c\\
a & c & b
\end{pmatrix},\begin{pmatrix}a & b & c\\
b & a & c
\end{pmatrix},\begin{pmatrix}a & b & c\\
c & b & a
\end{pmatrix}
$$

***

$$
\begin{aligned}
\left(s\right)\Rightarrow & \begin{cases}
\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}=\boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right) & \begin{pmatrix}a & b & c\\
a & b & c
\end{pmatrix},s_{{\scriptscriptstyle 1}}\\
\left(\boldsymbol{b}\times\boldsymbol{c}\right)\cdot\boldsymbol{a}=\boldsymbol{b}\cdot\left(\boldsymbol{c}\times\boldsymbol{a}\right) & \begin{pmatrix}a & b & c\\
b & c & a
\end{pmatrix},s_{{\scriptscriptstyle 2}}\\
\left(\boldsymbol{c}\times\boldsymbol{a}\right)\cdot\boldsymbol{b}=\boldsymbol{c}\cdot\left(\boldsymbol{a}\times\boldsymbol{b}\right) & \begin{pmatrix}a & b & c\\
c & a & b
\end{pmatrix},s_{{\scriptscriptstyle 3}}\\
\left(\boldsymbol{b}\times\boldsymbol{a}\right)\cdot\boldsymbol{c}=\boldsymbol{b}\cdot\left(\boldsymbol{a}\times\boldsymbol{c}\right) & \begin{pmatrix}a & b & c\\
b & a & c
\end{pmatrix},s_{{\scriptscriptstyle 4}}\\
\left(\boldsymbol{a}\times\boldsymbol{c}\right)\cdot\boldsymbol{b}=\boldsymbol{a}\cdot\left(\boldsymbol{c}\times\boldsymbol{b}\right) & \begin{pmatrix}a & b & c\\
a & c & b
\end{pmatrix},s_{{\scriptscriptstyle 5}}\\
\left(\boldsymbol{c}\times\boldsymbol{b}\right)\cdot\boldsymbol{a}=\boldsymbol{c}\cdot\left(\boldsymbol{b}\times\boldsymbol{a}\right) & \begin{pmatrix}a & b & c\\
c & b & a
\end{pmatrix},s_{{\scriptscriptstyle 6}}
\end{cases}\\
\overset{\cdot\text{ commutative}}{\Rightarrow} & \left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}\overset{s_{{\scriptscriptstyle 1}}}{=}\left(\boldsymbol{b}\times\boldsymbol{c}\right)\cdot\boldsymbol{a}\overset{s_{{\scriptscriptstyle 2}}}{=}\left(\boldsymbol{c}\times\boldsymbol{a}\right)\cdot\boldsymbol{b}\overset{s_{{\scriptscriptstyle 3}}}{=}\left(\boldsymbol{a}\times\boldsymbol{b}\right)\cdot\boldsymbol{c}\\
\overset{\times\text{ anticommutative}}{=} & -\left(\boldsymbol{b}\times\boldsymbol{a}\right)\cdot\boldsymbol{c}\overset{s_{{\scriptscriptstyle 6}}}{=}-\left(\boldsymbol{c}\times\boldsymbol{b}\right)\cdot\boldsymbol{a}\overset{s_{{\scriptscriptstyle 5}}}{=}-\left(\boldsymbol{a}\times\boldsymbol{c}\right)\cdot\boldsymbol{b}\overset{s_{{\scriptscriptstyle 4}}}{=}-\left(\boldsymbol{b}\times\boldsymbol{a}\right)\cdot\boldsymbol{c}\\
\Leftrightarrow & \boldsymbol{a}\cdot\left(\boldsymbol{b}\times\boldsymbol{c}\right)=\boldsymbol{b}\cdot\left(\boldsymbol{c}\times\boldsymbol{a}\right)=\boldsymbol{c}\cdot\left(\boldsymbol{a}\times\boldsymbol{b}\right)\\
= & -\boldsymbol{a}\cdot\left(\boldsymbol{c}\times\boldsymbol{b}\right)=-\boldsymbol{b}\cdot\left(\boldsymbol{a}\times\boldsymbol{c}\right)=-\boldsymbol{c}\cdot\left(\boldsymbol{b}\times\boldsymbol{a}\right)
\end{aligned}
$$

***

$$
\begin{aligned}
\left(v\right)\Rightarrow & \begin{cases}
\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}-\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}=\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)-\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right) & \begin{pmatrix}a & b & c\\
a & b & c
\end{pmatrix},v_{{\scriptscriptstyle 1}}\\
\left(\boldsymbol{b}\times\boldsymbol{c}\right)\times\boldsymbol{a}-\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)\boldsymbol{a}=\boldsymbol{b}\times\left(\boldsymbol{c}\times\boldsymbol{a}\right)-\boldsymbol{b}\left(\boldsymbol{c}\cdot\boldsymbol{a}\right) & \begin{pmatrix}a & b & c\\
b & c & a
\end{pmatrix},v_{{\scriptscriptstyle 2}}\\
\left(\boldsymbol{c}\times\boldsymbol{a}\right)\times\boldsymbol{b}-\left(\boldsymbol{c}\cdot\boldsymbol{a}\right)\boldsymbol{b}=\boldsymbol{c}\times\left(\boldsymbol{a}\times\boldsymbol{b}\right)-\boldsymbol{c}\left(\boldsymbol{a}\cdot\boldsymbol{b}\right) & \begin{pmatrix}a & b & c\\
c & a & b
\end{pmatrix},v_{{\scriptscriptstyle 3}}\\
\left(\boldsymbol{b}\times\boldsymbol{a}\right)\times\boldsymbol{c}-\left(\boldsymbol{b}\cdot\boldsymbol{a}\right)\boldsymbol{c}=\boldsymbol{b}\times\left(\boldsymbol{a}\times\boldsymbol{c}\right)-\boldsymbol{b}\left(\boldsymbol{a}\cdot\boldsymbol{c}\right) & \begin{pmatrix}a & b & c\\
b & a & c
\end{pmatrix},v_{{\scriptscriptstyle 4}}\\
\left(\boldsymbol{a}\times\boldsymbol{c}\right)\times\boldsymbol{b}-\left(\boldsymbol{a}\cdot\boldsymbol{c}\right)\boldsymbol{b}=\boldsymbol{a}\times\left(\boldsymbol{c}\times\boldsymbol{b}\right)-\boldsymbol{a}\left(\boldsymbol{c}\cdot\boldsymbol{b}\right) & \begin{pmatrix}a & b & c\\
a & c & b
\end{pmatrix},v_{{\scriptscriptstyle 5}}\\
\left(\boldsymbol{c}\times\boldsymbol{b}\right)\times\boldsymbol{a}-\left(\boldsymbol{c}\cdot\boldsymbol{b}\right)\boldsymbol{a}=\boldsymbol{c}\times\left(\boldsymbol{b}\times\boldsymbol{a}\right)-\boldsymbol{c}\left(\boldsymbol{b}\cdot\boldsymbol{a}\right) & \begin{pmatrix}a & b & c\\
c & b & a
\end{pmatrix},v_{{\scriptscriptstyle 6}}
\end{cases}\\
\Rightarrow & \begin{cases}
-Z-C=X-A & v_{{\scriptscriptstyle 1}}\\
-X-A=Y-B & v_{{\scriptscriptstyle 2}}\\
-Y-B=Z-C & v_{{\scriptscriptstyle 3}}\\
Z-C=-Y-B & v_{{\scriptscriptstyle 4}}\\
Y-B=-X-A & v_{{\scriptscriptstyle 5}}\\
X-A=-Z-C & v_{{\scriptscriptstyle 6}}
\end{cases},\begin{cases}
\cdot\text{ and scalar-vector product} & \text{commutative}\\
\times & \text{anticommutative}
\end{cases},\\
 & \begin{cases}
X=\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)=-\left(\boldsymbol{b}\times\boldsymbol{c}\right)\times\boldsymbol{a}=-\boldsymbol{a}\times\left(\boldsymbol{c}\times\boldsymbol{b}\right)=\left(\boldsymbol{c}\times\boldsymbol{b}\right)\times\boldsymbol{a}\\
Y=\boldsymbol{b}\times\left(\boldsymbol{c}\times\boldsymbol{a}\right)=-\left(\boldsymbol{c}\times\boldsymbol{a}\right)\times\boldsymbol{b}=-\boldsymbol{b}\times\left(\boldsymbol{a}\times\boldsymbol{c}\right)=\left(\boldsymbol{a}\times\boldsymbol{c}\right)\times\boldsymbol{b}\\
Z=\boldsymbol{c}\times\left(\boldsymbol{a}\times\boldsymbol{b}\right)=-\left(\boldsymbol{a}\times\boldsymbol{b}\right)\times\boldsymbol{c}=-\boldsymbol{c}\times\left(\boldsymbol{b}\times\boldsymbol{a}\right)=\left(\boldsymbol{b}\times\boldsymbol{a}\right)\times\boldsymbol{c}\\
A=\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)=\boldsymbol{a}\left(\boldsymbol{c}\cdot\boldsymbol{b}\right)=\left(\boldsymbol{c}\cdot\boldsymbol{b}\right)\boldsymbol{a}=\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)\boldsymbol{a}\\
B=\boldsymbol{b}\left(\boldsymbol{c}\cdot\boldsymbol{a}\right)=\boldsymbol{b}\left(\boldsymbol{a}\cdot\boldsymbol{c}\right)=\left(\boldsymbol{a}\cdot\boldsymbol{c}\right)\boldsymbol{b}=\left(\boldsymbol{c}\cdot\boldsymbol{a}\right)\boldsymbol{b}\\
C=\boldsymbol{c}\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)=\boldsymbol{c}\left(\boldsymbol{b}\cdot\boldsymbol{a}\right)=\left(\boldsymbol{b}\cdot\boldsymbol{a}\right)\boldsymbol{c}=\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\boldsymbol{c}
\end{cases}\\
\Rightarrow & \begin{cases}
-Z-C=X-A & v_{{\scriptscriptstyle 1}}=v_{{\scriptscriptstyle 6}}\\
-X-A=Y-B & v_{{\scriptscriptstyle 2}}=v_{{\scriptscriptstyle 5}}\\
-Y-B=Z-C & v_{{\scriptscriptstyle 3}}=v_{{\scriptscriptstyle 4}}
\end{cases}\Leftrightarrow\begin{cases}
Z+X=A-C\\
X+Y=B-A\\
Y+Z=C-B
\end{cases}\Leftrightarrow\begin{cases}
X+Y=B-A\\
Y+Z=C-B\\
Z+X=A-C
\end{cases}\\
\Leftrightarrow & \begin{cases}
2\left(X+Y+Z\right)=0 & \Rightarrow X+Y+Z=0\\
Y+Z=C-B & \Rightarrow X=B-C\Leftrightarrow\boldsymbol{a}\times\left(\boldsymbol{b}\times\boldsymbol{c}\right)=\boldsymbol{b}\left(\boldsymbol{c}\cdot\boldsymbol{a}\right)-\boldsymbol{c}\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)\text{ "back cab"}\\
Z+X=A-C & \Rightarrow Y=C-A\Leftrightarrow\boldsymbol{b}\times\left(\boldsymbol{c}\times\boldsymbol{a}\right)=\boldsymbol{c}\left(\boldsymbol{a}\cdot\boldsymbol{b}\right)-\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)\\
X+Y=B-A & \Rightarrow Z=A-B\Leftrightarrow\boldsymbol{c}\times\left(\boldsymbol{a}\times\boldsymbol{b}\right)=\boldsymbol{a}\left(\boldsymbol{b}\cdot\boldsymbol{c}\right)-\boldsymbol{b}\left(\boldsymbol{c}\cdot\boldsymbol{a}\right)
\end{cases}
\end{aligned}
$$

#### differential operator

https://en.wikipedia.org/wiki/Differential_operator

##### 4-differential operator

4-differential operator / four-differential operator = d'Alembert operator

$$
\begin{aligned}
\mathrm{D}= & \dfrac{\partial}{\partial t}+\mathrm{i}\dfrac{\partial}{\partial x}+\mathrm{j}\dfrac{\partial}{\partial y}+\mathrm{k}\dfrac{\partial}{\partial z}=\partial_{{\scriptscriptstyle t}}+\mathrm{i}\partial_{{\scriptscriptstyle x}}+\mathrm{j}\partial_{{\scriptscriptstyle y}}+\mathrm{k}\partial_{z}\\
= & \dfrac{\partial}{\partial t}+\boldsymbol{i}\dfrac{\partial}{\partial x}+\boldsymbol{j}\dfrac{\partial}{\partial y}+\boldsymbol{k}\dfrac{\partial}{\partial z}=\partial_{{\scriptscriptstyle t}}+\boldsymbol{i}\partial_{{\scriptscriptstyle x}}+\boldsymbol{j}\partial_{{\scriptscriptstyle y}}+\boldsymbol{k}\partial_{z}\\
= & \dfrac{\partial}{\partial x_{{\scriptscriptstyle 0}}}+\boldsymbol{e}_{{\scriptscriptstyle 1}}\dfrac{\partial}{\partial x_{{\scriptscriptstyle 1}}}+\boldsymbol{e}_{{\scriptscriptstyle 2}}\dfrac{\partial}{\partial x_{{\scriptscriptstyle 2}}}+\boldsymbol{e}_{{\scriptscriptstyle 3}}\dfrac{\partial}{\partial x_{{\scriptscriptstyle 3}}}=\partial_{{\scriptscriptstyle 0}}+\boldsymbol{e}_{{\scriptscriptstyle i}}\partial_{{\scriptscriptstyle i}}=\partial_{{\scriptscriptstyle 0}}+\boldsymbol{\nabla}
\end{aligned}
$$

$$
\mathrm{D}=\partial_{{\scriptscriptstyle 0}}+\mathrm{i}\partial_{{\scriptscriptstyle 1}}+\mathrm{j}\partial_{{\scriptscriptstyle 2}}+\mathrm{k}\partial_{{\scriptscriptstyle 3}}=\partial_{{\scriptscriptstyle 0}}+\boldsymbol{\nabla}=\mathrm{D}_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}
$$

##### nabla

nabla = spatial differential operator = 3-differential operator / three-differential operator

$$
\boldsymbol{\nabla}=\boldsymbol{e}_{{\scriptscriptstyle i}}\partial_{{\scriptscriptstyle i}}=\sum_{i=1}^{3}\boldsymbol{e}_{{\scriptscriptstyle i}}\partial_{{\scriptscriptstyle i}}=\sum_{i=1}^{3}\boldsymbol{e}_{{\scriptscriptstyle i}}\dfrac{\partial}{\partial x_{{\scriptscriptstyle i}}}=\begin{pmatrix}\dfrac{\partial}{\partial x} & \dfrac{\partial}{\partial y} & \dfrac{\partial}{\partial z}\end{pmatrix}^{\intercal}=\begin{pmatrix}\dfrac{\partial}{\partial x}\\
\dfrac{\partial}{\partial y}\\
\dfrac{\partial}{\partial z}
\end{pmatrix}
$$

##### Laplace operator

Laplace operator = Laplacian

$$
\triangle=\boldsymbol{\nabla}^{2}=\boldsymbol{\nabla}\cdot\boldsymbol{\nabla}=\dfrac{\partial^{2}}{\partial x^{2}}+\dfrac{\partial^{2}}{\partial y^{2}}+\dfrac{\partial^{2}}{\partial z^{2}}
$$

##### d'Alembert operator

$$
\square=\square_{{\scriptscriptstyle c}}=\dfrac{1}{c^{2}}\dfrac{\partial^{2}}{\partial t^{2}}-\dfrac{\partial^{2}}{\partial x^{2}}-\dfrac{\partial^{2}}{\partial y^{2}}-\dfrac{\partial^{2}}{\partial z^{2}}=\dfrac{1}{c^{2}}\dfrac{\partial^{2}}{\partial t^{2}}-\triangle=\dfrac{1}{c^{2}}\dfrac{\partial^{2}}{\partial t^{2}}-\boldsymbol{\nabla}^{2}
$$

$$
\square_{{\scriptscriptstyle 1}}=\square_{{\scriptscriptstyle c=1}}=\dfrac{\partial^{2}}{\partial t^{2}}-\dfrac{\partial^{2}}{\partial x^{2}}-\dfrac{\partial^{2}}{\partial y^{2}}-\dfrac{\partial^{2}}{\partial z^{2}}=\dfrac{\partial^{2}}{\partial t^{2}}-\triangle=\dfrac{\partial^{2}}{\partial t^{2}}-\boldsymbol{\nabla}^{2}
$$

#### electromagnetism

Maxwell

##### 4-potential

electromagnetic 4-potential / four-potential

$$
A=A_{{\scriptscriptstyle 0}}++\mathrm{i}A_{{\scriptscriptstyle 1}}+\mathrm{j}A_{{\scriptscriptstyle 2}}+\mathrm{k}A_{{\scriptscriptstyle 3}}=A_{0}+\boldsymbol{A}
$$

[4-differential operator]^[\@ref(differential-operator-1)]^

$$
\mathrm{D}=\partial_{{\scriptscriptstyle 0}}+\mathrm{i}\partial_{{\scriptscriptstyle 1}}+\mathrm{j}\partial_{{\scriptscriptstyle 2}}+\mathrm{k}\partial_{{\scriptscriptstyle 3}}=\partial_{{\scriptscriptstyle 0}}+\boldsymbol{\nabla}=\mathrm{D}_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}
$$

$$
QP=\left(Q_{{\scriptscriptstyle 0}}P_{{\scriptscriptstyle 0}}-\boldsymbol{Q}\cdot\boldsymbol{P}\right)+\left(Q_{{\scriptscriptstyle 0}}\boldsymbol{P}+P_{{\scriptscriptstyle 0}}\boldsymbol{Q}+\boldsymbol{Q}\times\boldsymbol{P}\right)
$$

[commutative bracket]^[\@ref(commutative-bracket)]^

$$
\begin{aligned}
\left[\mathrm{D},A\right]= & \dfrac{\mathrm{D}A-A\mathrm{D}}{2}\\
2\left[\mathrm{D},A\right]= & \mathrm{D}A-A\mathrm{D}\\
= & \left(\partial_{{\scriptscriptstyle 0}}+\mathrm{i}\partial_{{\scriptscriptstyle 1}}+\mathrm{j}\partial_{{\scriptscriptstyle 2}}+\mathrm{k}\partial_{{\scriptscriptstyle 3}}\right)\left(A_{{\scriptscriptstyle 0}}++\mathrm{i}A_{{\scriptscriptstyle 1}}+\mathrm{j}A_{{\scriptscriptstyle 2}}+\mathrm{k}A_{{\scriptscriptstyle 3}}\right)\\
- & \left(A_{{\scriptscriptstyle 0}}++\mathrm{i}A_{{\scriptscriptstyle 1}}+\mathrm{j}A_{{\scriptscriptstyle 2}}+\mathrm{k}A_{{\scriptscriptstyle 3}}\right)\left(\partial_{{\scriptscriptstyle 0}}+\mathrm{i}\partial_{{\scriptscriptstyle 1}}+\mathrm{j}\partial_{{\scriptscriptstyle 2}}+\mathrm{k}\partial_{{\scriptscriptstyle 3}}\right)\\
\mathrm{D}A= & \left(\mathrm{D}_{{\scriptscriptstyle 0}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}\right)+\left(\mathrm{D}_{{\scriptscriptstyle 0}}\boldsymbol{A}+A_{{\scriptscriptstyle 0}}\boldsymbol{\mathrm{D}}+\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\right)\\
= & \left(\mathrm{D}_{{\scriptscriptstyle 0}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}\right)+\left(\mathrm{D}_{{\scriptscriptstyle 0}}\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\right)\\
= & \mathrm{D}_{{\scriptscriptstyle 0}}\left(A_{{\scriptscriptstyle 0}}+\boldsymbol{A}\right)-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\\
= & \mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\\
= & \partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}+\boldsymbol{\nabla}\times\boldsymbol{A}\\
A\mathrm{D}= & \left(A_{{\scriptscriptstyle 0}}\mathrm{D}_{{\scriptscriptstyle 0}}-\boldsymbol{A}\cdot\boldsymbol{\mathrm{D}}\right)+\left(A_{{\scriptscriptstyle 0}}\boldsymbol{\mathrm{D}}+\mathrm{D}_{{\scriptscriptstyle 0}}\boldsymbol{A}+\boldsymbol{A}\times\boldsymbol{\mathrm{D}}\right)\\
= & \left(\mathrm{D}_{{\scriptscriptstyle 0}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}\right)+\left(\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}+\mathrm{D}_{{\scriptscriptstyle 0}}\boldsymbol{A}-\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\right)\\
= & \left(\mathrm{D}_{{\scriptscriptstyle 0}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}\right)+\left(\mathrm{D}_{{\scriptscriptstyle 0}}\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\right)\\
= & \mathrm{D}_{{\scriptscriptstyle 0}}\left(A_{{\scriptscriptstyle 0}}+\boldsymbol{A}\right)-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\\
= & \mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\times\boldsymbol{A}\\
= & \partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\times\boldsymbol{A}
\end{aligned}
$$

$$
\begin{aligned}
\mathrm{D}A= & \mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}+\boldsymbol{\mathrm{D}}\times\boldsymbol{A}=\partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}+\boldsymbol{\nabla}\times\boldsymbol{A}\\
A\mathrm{D}= & \mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}-\boldsymbol{\mathrm{D}}\times\boldsymbol{A}=\partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\times\boldsymbol{A}
\end{aligned}
$$

$$
\begin{aligned}
 & \mathrm{D}A-A\mathrm{D}=2\boldsymbol{\mathrm{D}}\times\boldsymbol{A}=2\boldsymbol{\nabla}\times\boldsymbol{A}\\
\left[\mathrm{D},A\right]= & \dfrac{\mathrm{D}A-A\mathrm{D}}{2}=\boldsymbol{\mathrm{D}}\times\boldsymbol{A}=\boldsymbol{\nabla}\times\boldsymbol{A}
\end{aligned}
$$

[anticommutative bracket]^[\@ref(anticommutative-bracket)]^

$$
\begin{aligned}
 & \mathrm{D}A+A\mathrm{D}=2\left(\mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}\right)=2\left(\partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}\right)\\
\left\{ \mathrm{D},A\right\} = & \dfrac{\mathrm{D}A+A\mathrm{D}}{2}=\mathrm{D}_{{\scriptscriptstyle 0}}A-\boldsymbol{\mathrm{D}}\cdot\boldsymbol{A}+\boldsymbol{\mathrm{D}}A_{{\scriptscriptstyle 0}}=\partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}
\end{aligned}
$$

commutation and anticommutation on differential operator and any quaternion

$$
\begin{aligned}
\left[\mathrm{D},Q\right]= & \boldsymbol{\nabla}\times\boldsymbol{Q}\\
\left\{ \mathrm{D},Q\right\} = & \partial_{{\scriptscriptstyle 0}}Q-\boldsymbol{\nabla}\cdot\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}
\end{aligned}
$$

or more evident

$$
\begin{aligned}
\left[\mathrm{D},Q\right]= & \boldsymbol{\nabla}\times\boldsymbol{Q}\\
\left\{ \mathrm{D},Q\right\} = & \partial_{{\scriptscriptstyle 0}}Q-\boldsymbol{\nabla}\cdot\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\\
= & \partial_{{\scriptscriptstyle 0}}\left(Q_{{\scriptscriptstyle 0}}+\boldsymbol{Q}\right)-\boldsymbol{\nabla}\cdot\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\\
= & \left(\partial_{{\scriptscriptstyle 0}}Q_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\cdot\boldsymbol{Q}\right)+\left(\partial_{{\scriptscriptstyle 0}}\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\right)\\
= & \left(\dfrac{\partial Q_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{Q}\right)+\left(\dfrac{\partial\boldsymbol{Q}}{\partial t}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\right)
\end{aligned}
$$

##### Maxwell compromise for both quaternion and 3-vector

electric potential and vector potential

$$
A=A_{{\scriptscriptstyle 0}}++\mathrm{i}A_{{\scriptscriptstyle 1}}+\mathrm{j}A_{{\scriptscriptstyle 2}}+\mathrm{k}A_{{\scriptscriptstyle 3}}=A_{0}+\boldsymbol{A}=U+\boldsymbol{A}
$$

electric quaternion and electric field

$$
\begin{aligned}
E= & -\left\{ \mathrm{D},A\right\} \\
= & -\left(\partial_{{\scriptscriptstyle 0}}A-\boldsymbol{\nabla}\cdot\boldsymbol{A}+\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}\right)\\
= & -\partial_{{\scriptscriptstyle 0}}A+\boldsymbol{\nabla}\cdot\boldsymbol{A}-\boldsymbol{\nabla}A_{{\scriptscriptstyle 0}}\\
= & -\partial_{{\scriptscriptstyle t}}\left(U+\boldsymbol{A}\right)+\boldsymbol{\nabla}\cdot\boldsymbol{A}-\boldsymbol{\nabla}U\\
= & -\dfrac{\partial U}{\partial t}+\boldsymbol{\nabla}\cdot\boldsymbol{A}-\boldsymbol{\nabla}U-\dfrac{\partial\boldsymbol{A}}{\partial t}\\
= & E_{{\scriptscriptstyle 0}}+\boldsymbol{E},\begin{cases}
E_{{\scriptscriptstyle 0}}=-\dfrac{\partial U}{\partial t}+\boldsymbol{\nabla}\cdot\boldsymbol{A}\\
\boldsymbol{E}=-\boldsymbol{\nabla}U-\dfrac{\partial\boldsymbol{A}}{\partial t} & \text{electric field 3-vector}
\end{cases}
\end{aligned}
$$

magnetic field

$$
B=\left[\mathrm{D},A\right]=\boldsymbol{\nabla}\times\boldsymbol{A}=\begin{vmatrix}\boldsymbol{i} & \boldsymbol{j} & \boldsymbol{k}\\
\dfrac{\partial}{\partial x} & \dfrac{\partial}{\partial y} & \dfrac{\partial}{\partial z}\\
A_{{\scriptscriptstyle 1}} & A_{{\scriptscriptstyle 2}} & A_{{\scriptscriptstyle 3}}
\end{vmatrix}=\boldsymbol{B}
$$

Work on time? Yes.

$$
qE=qE_{{\scriptscriptstyle 0}}+q\boldsymbol{E}=qE_{{\scriptscriptstyle 0}}+\boldsymbol{F}_{E}
$$

force equivalent on time

$$
qE_{{\scriptscriptstyle 0}}
$$

***

$$
\begin{cases}
E=-\left\{ \mathrm{D},A\right\} =E_{{\scriptscriptstyle 0}}+\boldsymbol{E}\\
B=+\left[\mathrm{D},A\right]=B_{{\scriptscriptstyle 0}}+\boldsymbol{B}=0+\boldsymbol{B}=\boldsymbol{B} & B_{{\scriptscriptstyle 0}}=0
\end{cases}
$$

***

for any quaternion commutating and anticommutating with differential operator

$$
\begin{aligned}
\left[\mathrm{D},Q\right]= & \boldsymbol{\nabla}\times\boldsymbol{Q}\\
\left\{ \mathrm{D},Q\right\} = & \partial_{{\scriptscriptstyle 0}}Q-\boldsymbol{\nabla}\cdot\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\\
= & \partial_{{\scriptscriptstyle 0}}\left(Q_{{\scriptscriptstyle 0}}+\boldsymbol{Q}\right)-\boldsymbol{\nabla}\cdot\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\\
= & \left(\partial_{{\scriptscriptstyle 0}}Q_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\cdot\boldsymbol{Q}\right)+\left(\partial_{{\scriptscriptstyle 0}}\boldsymbol{Q}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\right)\\
= & \left(\dfrac{\partial Q_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{Q}\right)+\left(\dfrac{\partial\boldsymbol{Q}}{\partial t}+\boldsymbol{\nabla}Q_{{\scriptscriptstyle 0}}\right)
\end{aligned}
$$

***

$$
\begin{cases}
\left[\mathrm{D},E\right]=\boldsymbol{\nabla}\times\boldsymbol{E}=\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{E}\right)\\
\left\{ \mathrm{D},E\right\} =\left(\partial_{{\scriptscriptstyle 0}}E_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\partial_{{\scriptscriptstyle 0}}\boldsymbol{E}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)=\left(\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)\\
\left[\mathrm{D},B\right]=\boldsymbol{\nabla}\times\boldsymbol{B}=\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{B}\right)\\
\left\{ \mathrm{D},B\right\} =\left(\partial_{{\scriptscriptstyle 0}}B_{{\scriptscriptstyle 0}}-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\partial_{{\scriptscriptstyle 0}}\boldsymbol{B}+\boldsymbol{\nabla}B_{{\scriptscriptstyle 0}}\right)\overset{B_{{\scriptscriptstyle 0}}=0}{=}-\boldsymbol{\nabla}\cdot\boldsymbol{B}+\partial_{{\scriptscriptstyle 0}}\boldsymbol{B}=\left(-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\dfrac{\partial\boldsymbol{B}}{\partial t}\right)
\end{cases}
$$

by comparing ( real & imaginary ) / ( scalar & vector ) parts,

Maxwell equations without source terms

::: {show-in="html"}
$$
\begin{aligned}
 & \begin{cases}
\left[\mathrm{D},B\right]=+\left\{ \mathrm{D},E\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{B}\right)=\left(\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)\\
\left[\mathrm{D},E\right]=-\left\{ \mathrm{D},B\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{E}\right)=\left(-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\dfrac{\partial\boldsymbol{B}}{\partial t}\right)
\end{cases}\\
\Leftrightarrow & \begin{cases}
\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{E}=\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}\\
\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}=\boldsymbol{\nabla}\times\boldsymbol{B} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{B}=\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\text{ 動電生磁}\\
-\boldsymbol{\nabla}\cdot\boldsymbol{B}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{B}=0\\
\dfrac{\partial\boldsymbol{B}}{\partial t}=\boldsymbol{\nabla}\times\boldsymbol{E} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{E}=\dfrac{\partial\boldsymbol{B}}{\partial t}\text{ 動磁生電}
\end{cases}
\end{aligned}
$$
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
$$
\begin{aligned}
 & \begin{cases}
\left[\mathrm{D},B\right]=+\left\{ \mathrm{D},E\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{B}\right)=\left(\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)\\
\left[\mathrm{D},E\right]=-\left\{ \mathrm{D},B\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{E}\right)=\left(-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\dfrac{\partial\boldsymbol{B}}{\partial t}\right)
\end{cases}\\
\Leftrightarrow & \begin{cases}
\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{E}=\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}\\
\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}=\boldsymbol{\nabla}\times\boldsymbol{B} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{B}=\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\text{ 動電生磁}\\
-\boldsymbol{\nabla}\cdot\boldsymbol{B}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{B}=0\\
\dfrac{\partial\boldsymbol{B}}{\partial t}=\boldsymbol{\nabla}\times\boldsymbol{E} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{E}=\dfrac{\partial\boldsymbol{B}}{\partial t}\text{ 動磁生電}
\end{cases}
\end{aligned}
$$
\end{CJK}
:::

#### Joule heat vs. Thomson heat (Kelvin heat?)

The Lord Kelvin = William Thomson

##### thermoelectric effect

thermoelectric effect = Seebeck effect = Peltier effect = Thomson effect

#### source term

$$
J=J_{{\scriptscriptstyle 0}}+\mathrm{i}J_{{\scriptscriptstyle 1}}+\mathrm{j}J_{{\scriptscriptstyle 2}}+\mathrm{k}J_{{\scriptscriptstyle 3}}=J_{{\scriptscriptstyle 0}}+\boldsymbol{J}=\rho+\boldsymbol{J}
$$

Maxwell equations with source terms

::: {show-in="html"}
$$
\begin{aligned}
 & \begin{cases}
\left[\mathrm{D},B\right]=J+\left\{ \mathrm{D},E\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{B}\right)=\left(\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)\\
\left[\mathrm{D},E\right]=0-\left\{ \mathrm{D},B\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{E}\right)=\left(-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\dfrac{\partial\boldsymbol{B}}{\partial t}\right)
\end{cases}\\
\Leftrightarrow & \begin{cases}
\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{E}=\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}\\
\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}=\boldsymbol{\nabla}\times\boldsymbol{B} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{B}=\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\text{ 動電生磁}\\
-\boldsymbol{\nabla}\cdot\boldsymbol{B}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{B}=0\\
\dfrac{\partial\boldsymbol{B}}{\partial t}=\boldsymbol{\nabla}\times\boldsymbol{E} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{E}=\dfrac{\partial\boldsymbol{B}}{\partial t}\text{ 動磁生電}
\end{cases}
\end{aligned}
$$
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
$$
\begin{aligned}
 & \begin{cases}
\left[\mathrm{D},B\right]=J+\left\{ \mathrm{D},E\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{B}\right)=\left(\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}\right)+\left(\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\right)\\
\left[\mathrm{D},E\right]=0-\left\{ \mathrm{D},B\right\}  & \Leftrightarrow\left(0\right)+\left(\boldsymbol{\nabla}\times\boldsymbol{E}\right)=\left(-\boldsymbol{\nabla}\cdot\boldsymbol{B}\right)+\left(\dfrac{\partial\boldsymbol{B}}{\partial t}\right)
\end{cases}\\
\Leftrightarrow & \begin{cases}
\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}-\boldsymbol{\nabla}\cdot\boldsymbol{E}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{E}=\rho+\dfrac{\partial E_{{\scriptscriptstyle 0}}}{\partial t}\\
\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}=\boldsymbol{\nabla}\times\boldsymbol{B} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{B}=\boldsymbol{J}+\dfrac{\partial\boldsymbol{E}}{\partial t}+\boldsymbol{\nabla}E_{{\scriptscriptstyle 0}}\text{ 動電生磁}\\
-\boldsymbol{\nabla}\cdot\boldsymbol{B}=0 & \Leftrightarrow\boldsymbol{\nabla}\cdot\boldsymbol{B}=0\\
\dfrac{\partial\boldsymbol{B}}{\partial t}=\boldsymbol{\nabla}\times\boldsymbol{E} & \Leftrightarrow\boldsymbol{\nabla}\times\boldsymbol{E}=\dfrac{\partial\boldsymbol{B}}{\partial t}\text{ 動磁生電}
\end{cases}
\end{aligned}
$$
\end{CJK}
:::

### quaternion group

https://en.wikipedia.org/wiki/Quaternion_group

[group theory]^[\@ref(group-theory)]^

or please first see [quaternion group representation]^[\@ref(quaternion-group-representation)]^.

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1rj41117VW&autoplay=0"></iframe>
:::



### octonion

## 3Blue1Brown

### Ben Eater

https://eater.net/quaternions/video/intro

https://eater.net/quaternions

### 3B1B

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://www.youtube.com/embed/zjMuIxRvygQ"></iframe>

<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://www.youtube.com/embed/d4EgbgTm0Bg"></iframe>
:::

### Sutrabla

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://www.youtube.com/embed/syQnn_xuB8U"></iframe>

<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://www.youtube.com/embed/4mXL751ko0w"></iframe>
:::

https://www.newscientist.com/article/mg20427391-600-alices-adventures-in-algebra-wonderland-solved/

https://threejs.org/

## CCJou: LA Revelation


::: {show-in="html"}
https://ccjou.wordpress.com/2014/04/21/四元數/
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
https://ccjou.wordpress.com/2014/04/21/四元數/
\end{CJK}
:::



