# test2

```{r}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

https://bookdown.org/yihui/rmarkdown-cookbook/html-scroll.html

```{css, echo=FALSE}
pre {
  max-width: 400px;
  overflow-x: auto;
  max-height: 600px;
  overflow-y: auto;
}

/*
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
*/
```

## commutative diagram

cd = CD = commutative diagram

tikz-cd = TikZ-CD

`\usetikzlibrary{cd}` vs. `usepackage{tikz-cd}`

https://tex.stackexchange.com/questions/546392/usepackagetikz-cd-vs-usetikzlibrarycd

> No (practical) difference. The code (excluding copyright comment lines) in `tikz-cd.sty` is
> ```
> \ProvidesPackage{tikz-cd}[2014/10/30 v0.9e Commutative diagrams with tikz]
> \RequirePackage{tikz}[2013/12/13] % pgf version 3.0.0 required
> \usetikzlibrary{cd}
>
> \endinput
> ```

Never put `\begin{tikzcd}` inside `\begin{tikzpicture}`

https://tex.stackexchange.com/questions/425296/half-of-tikzcd-diagram-is-missing

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
00 & 01 & 02 & 03 \\
10 & 11 & 12 & 13 \\
20 & 21 & 22 & 23 \\
30 & 31 & 32 & 33
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
00 & 01 & 02 & 03 \\
10 & 11 & 12 & 13 \\
20 & 21 & 22 & 23 \\
30 & 31 & 32 & 33
\end{tikzcd}
```
:::
::::::

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
00
&
01
&
02
&
03
\\
10
&
11
&
12
&
13
\\
20
&
21
&
22
&
23
\\
30
&
31
&
32
&
33
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
00
&
01
&
02
&
03
\\
10
&
11
&
12
&
13
\\
20
&
21
&
22
&
23
\\
30
&
31
&
32
&
33
\end{tikzcd}
```
:::
::::::

***

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm]%small,large,huge
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm]%small,large,huge
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm]%small,large,huge
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```
:::
::::::

***

https://tex.stackexchange.com/questions/515267/how-can-i-make-arrows-touch-boxed-expressions-in-tikzcd

```
cells={nodes={draw}}

in

\begin{tikzcd}[column sep=2.75cm, %small,large,huge
			         cells={nodes={draw}}
			        ]
```

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm, %small,large,huge
			         cells={nodes={draw}}
			        ]
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm, %small,large,huge
			         cells={nodes={draw}}
			        ]
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='learn TikZ-CD or tikz-cd in one picture'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[column sep=2.75cm, %small,large,huge
			         cells={nodes={draw}}
			        ]
00
\ar[r,"\backslash \text{ar[r]}"]
\ar[d,"\backslash \text{ar[d]}"]
&
01
\ar[r,"\text{[,"swap"']}"']
&
02
\ar[r,"\backslash \text{ar[r]}","\text{[,"swap"']}"']
&
03
\\
10
\ar[d,"\text{[,"swap"']}"']
&
11
\ar[u,"\backslash \text{ar[u]}"]
\ar[l,"\backslash \text{ar[l]}"]
\ar[r,-stealth,"\text{[,-}\text{stealth}\text{]}"]
\ar[d,-{Stealth[reversed]},"\text{[,-}\{\text{Stealth[reversed]}\}\text{]}"]
&
12
\ar[r,-{Stealth[open]},"\text{[,-}\{\text{Stealth[open]}\}\text{]}"]
&
13
\\
20
\ar[r,"\text{[,"r" description]}" description]
\ar[d,"\backslash \text{ar[d]}","\text{[,"swap"']}"']
&
21
\ar[r,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
&
22
\ar[u,shift right=1.75pt,"\text{[,shift right=1.75pt]}"']
\ar[lld,-Stealth,"\backslash \text{ar[lld]}" description]
\ar[r,latex-latex,"\text{[,latex-latex]}"]
\ar[d,shift right=1.75pt,"\text{[,shift right=1.75pt]}"]
&
23
\\
30
\ar[ru,"\backslash \text{ar[ru]}" description]
\ar[r,bend right,-stealth,"\text{bend right}"]
\ar[r,bend right=42,-stealth,"\text{bend right=42}"']
\ar[r,bend right=100,-stealth,"\text{bend right=100}"']
\ar[dd,bend right,"\text{[,bend right]}"']
&
31
\ar[r,bend left,stealth-stealth,"\text{bend left}"']
\ar[ddr]
&
32
\ar[l,-{Stealth[harpoon]},"\text{[,-}\{\text{Stealth[harpoon]}\}\text{]}"]
\ar[r,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\ar[ddl,crossing over,"\text{[,crossing over]; rounded corneres, to path}"]
\ar[ddr,
	rounded corners,
	to path={--([yshift=-2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+2ex]\tikztostart.south)
			  --([yshift=-2ex,xshift=+8ex]\tikztostart.south)
			  --([xshift=-12ex]\tikztotarget.west)
			  --(\tikztotarget)
			 },
	]
&
33
\ar[l,-{Stealth[harpoon]},shift left=.75pt,"\text{[,shift left=.75pt]}"]
\\
&
&
&
\\
50
&
51
&
52
&
53
\end{tikzcd}
```
:::
::::::

https://tex.stackexchange.com/questions/484743/format-single-node-in-tikzcd

https://tex.stackexchange.com/questions/586160/drawing-commutative-diagram-with-encircled-nodes-in-tikz-cd-package

[box in TikZ-CD]^[\@ref(box-in-tikz-cd)]^

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
A \ar[r,-stealth,"f","g"'] \ar[d,-{Stealth[harpoon]},shift left=0.5pt,"\varphi"] \ar[rr,-stealth,bend left] & B & E \\
C \ar[r,-latex,"g"'] \ar[u,-{Stealth[harpoon]},shift left=0.5pt,"\varphi^{{\scriptscriptstyle -1}}"] \ar[d,harpoon,shift left=0.5pt,"\varphi"] & D & F \\
G \ar[r,"f"] \ar[u,harpoon,shift left=0.5pt,"\varphi^{-1}"] & H \ar[uur,-stealth,"\phi" description]
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='TikZ-CD or tikz-cd + arrows.meta'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
A \ar[r,-stealth,"f","g"'] \ar[d,-{Stealth[harpoon]},shift left=0.5pt,"\varphi"] \ar[rr,-stealth,bend left] & B & E \\
C \ar[r,-latex,"g"'] \ar[u,-{Stealth[harpoon]},shift left=0.5pt,"\varphi^{{\scriptscriptstyle -1}}"] \ar[d,harpoon,shift left=0.5pt,"\varphi"] & D & F \\
G \ar[r,"f"] \ar[u,harpoon,shift left=0.5pt,"\varphi^{-1}"] & H \ar[uur,-stealth,"\phi" description]
\end{tikzcd}
```
:::
::::::

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
A \ar[r,-{Stealth[harpoon,swap,open]}] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='TikZ-CD or tikz-cd + arrows.meta'}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}
A \ar[r,-{Stealth[harpoon,swap,open]}] & B
\end{tikzcd}
```
:::
::::::

TikZcd editor

https://github.com/yishn/tikzcd-editor

https://tikzcd.yichuanshen.de/

https://darknmt.github.io/res/tikzcd-editor/

tikz-cd @ CTAN(Comprehensive TEX Archive Network)

https://ctan.org/tex-archive/graphics/pgf/contrib/tikz-cd

tikzcd manual

https://ctan.mirror.twds.com.tw/tex-archive/graphics/pgf/contrib/tikz-cd/tikz-cd-doc.pdf

### arrows

https://latexdraw.com/exploring-tikz-arrows/

#### predefined arrow tip in TikZ

##### stealth arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]        (0, 0) -- (1, 0);
	\draw [stealth-]        (0,-1) -- (1,-1);
	\draw [stealth-stealth] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='stealth arrow tip'}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]        (0, 0) -- (1, 0);
	\draw [stealth-]        (0,-1) -- (1,-1);
	\draw [stealth-stealth] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### reversed stealth arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth reversed]                 (0, 0) -- (1, 0);
	\draw [stealth reversed-]                 (0,-1) -- (1,-1);
	\draw [stealth reversed-stealth reversed] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '50%', fig.cap='reversed stealth arrow tip'}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth reversed]                 (0, 0) -- (1, 0);
	\draw [stealth reversed-]                 (0,-1) -- (1,-1);
	\draw [stealth reversed-stealth reversed] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### to arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-to]   (0, 0) -- (1, 0);
	\draw [to-]   (0,-1) -- (1,-1);
	\draw [to-to] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='to arrow tip'}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-to]   (0, 0) -- (1, 0);
	\draw [to-]   (0,-1) -- (1,-1);
	\draw [to-to] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### latex arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-latex]      (0, 0) -- (1, 0);
	\draw [latex-]      (0,-1) -- (1,-1);
	\draw [latex-latex] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='to arrow tip'}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-latex]      (0, 0) -- (1, 0);
	\draw [latex-]      (0,-1) -- (1,-1);
	\draw [latex-latex] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### | arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-|]  (0, 0) -- (1, 0);
	\draw [|-]  (0,-1) -- (1,-1);
	\draw [|-|] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='to arrow tip'}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-|]  (0, 0) -- (1, 0);
	\draw [|-]  (0,-1) -- (1,-1);
	\draw [|-|] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

#### arrows.meta

customizing arrow heads

##### reversing, halving, swapping, opening

https://tikz.dev/tikz-arrows#sec-16.3.5

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-{Stealth[reversed]}]      (0, 0) -- (1, 0);
	\draw [-{Stealth[harpoon]}]       (0,-1) -- (1,-1);
	\draw [-{Stealth[harpoon,swap]}]  (0,-2) -- (1,-2);
  \draw [-{Stealth[open]}]          (0,-3) -- (1,-3);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '25%', fig.cap='reversing, halving, swapping, opening'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-{Stealth[reversed]}]      (0, 0) -- (1, 0);
	\draw [-{Stealth[harpoon]}]       (0,-1) -- (1,-1);
	\draw [-{Stealth[harpoon,swap]}]  (0,-2) -- (1,-2);
  \draw [-{Stealth[open]}]          (0,-3) -- (1,-3);
  \node at (0,-3) {};
\end{tikzpicture}
```
:::
::::::

##### mathematical arrows

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-{Stealth[reversed]}]      (0, 0) -- (1, 0);
	\draw [-{Stealth[harpoon]}]       (0,-1) -- (1,-1);
	\draw [-{Stealth[harpoon,swap]}]  (0,-2) -- (1,-2);
  \draw [-{Stealth[open]}]          (0,-3) -- (1,-3);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '50%', fig.cap='mathematical arrows'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-{Classical TikZ Rightarrow}]  (0, 0) -- (1, 0);
	\draw [-{Computer Modern Rightarrow}] (0,-1) -- (1,-1);
	\draw [-{Implies},double]             (0,-2) -- (1,-2);
  \draw [-To]                           (0,-3) -- (1,-3);
  \node at (0,-3) {};
\end{tikzpicture}
```
:::
::::::

##### ray arrow tip

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-Rays]         (0, 0) -- (1, 0);
	\draw [-{Rays[n=6]}]  (0,-1) -- (1,-1);
	\draw [-{Rays[n=8]}]  (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='Ray arrow tip'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-Rays]         (0, 0) -- (1, 0);
	\draw [-{Rays[n=6]}]  (0,-1) -- (1,-1);
	\draw [-{Rays[n=8]}]  (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### arrow head scale

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]            (0, 0) -- (1, 0);
	\draw [-{Stealth[scale=2]}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='arrow head scale'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]            (0, 0) -- (1, 0);
	\draw [-{Stealth[scale=2]}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::
::::::

##### arrow head size

https://tikz.dev/tikz-arrows#sec-16.3.1

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]                          (0, 0) -- (1, 0);
	\draw [-{Stealth[length=3mm, width=2mm]}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='arrow head size'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]                          (0, 0) -- (1, 0);
	\draw [-{Stealth[length=3mm, width=2mm]}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::
::::::

##### arrow head slant

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-Stealth]               (0, 0) -- (1, 0);
  \draw [-{Stealth[slant=-0.5]}] (0,-1) -- (1,-1);
  \draw [-{Stealth[slant=   1]}] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='arrow head slant'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]               (0, 0) -- (1, 0);
  \draw [-{Stealth[slant=-0.5]}] (0,-1) -- (1,-1);
  \draw [-{Stealth[slant=   1]}] (0,-2) -- (1,-2);
  \node at (0,-2) {};
\end{tikzpicture}
```
:::
::::::

##### arrow head color

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]               (0, 0) -- (1, 0);
	\draw [-{Stealth[color=cyan]}] (0,-1) -- (1,-1);
  \draw [-{Stealth[fill=red]}  ] (0,-2) -- (1,-2);
  \draw [red,-{Stealth}        ] (0,-3) -- (1,-3);
  \node at (0,-3) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '50%', fig.cap='arrow and arrow head color'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]               (0, 0) -- (1, 0);
  \draw [-{Stealth[color=cyan]}] (0,-1) -- (1,-1);
  \draw [-{Stealth[fill=red]}  ] (0,-2) -- (1,-2);
  \draw [red,-{Stealth}        ] (0,-3) -- (1,-3);
  \node at (0,-3) {};
\end{tikzpicture}
```
:::
::::::

##### multiple arrow head

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]                     (0, 0) -- (1, 0);
	\draw [-{Stealth}{Stealth}{Stealth}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='multiple arrow head'}
\usetikzlibrary{arrows.meta}
\begin{tikzpicture}
  \node at (0, 0) {};
	\draw [-stealth]                     (0, 0) -- (1, 0);
	\draw [-{Stealth}{Stealth}{Stealth}] (0,-1) -- (1,-1);
  \node at (0,-1) {};
\end{tikzpicture}
```
:::
::::::

#### default stealth arrow

https://tex.stackexchange.com/questions/113437/stealth-arrow-in-math

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows}
\tikzset{
  commutative diagrams/.cd, 
  arrow style=tikz, 
  diagrams={>=stealth}
}
\begin{tikzcd}
A \ar[r,"f"] \ar[d,harpoon,shift left=0.5pt,"\varphi"] \ar[rr,bend left] & B & E \\
C \ar[r] \ar[u,harpoon,shift left=0.5pt,"\varphi^{-1}"] & D & F \\
G & H \ar[uur,-stealth]
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='default stealth arrow'}
\usetikzlibrary{cd,arrows}
\tikzset{
  commutative diagrams/.cd, 
  arrow style=tikz, 
  diagrams={>=stealth}
}
\begin{tikzcd}
A \ar[r,"f"] \ar[d,harpoon,shift left=0.5pt,"\varphi"] \ar[rr,bend left] & B & E \\
C \ar[r] \ar[u,harpoon,shift left=0.5pt,"\varphi^{-1}"] & D & F \\
G & H \ar[uur,-stealth]
\end{tikzcd}
```
:::
::::::

#### harpoon

harpoon ~ half arrow

https://tex.stackexchange.com/questions/715462/rightleftharpoons-in-commutative-diagram

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
cc \ar[r,shift left=1pt,harpoon,"a"] \ar[d,shift left=1pt,harpoon,"r"] &
ct \ar[l,shift left=1pt,harpoon,"b"] \ar[d,shift left=1pt,harpoon,"r"]
\\
tc \ar[r,shift left=1pt,harpoon,"a"] \ar[u,shift left=1pt,harpoon,"l"] &
tt \ar[l,shift left=1pt,harpoon,"b"] \ar[u,shift left=1pt,harpoon,"l"]
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='harpoon in cd'}
\usetikzlibrary{cd}
\begin{tikzcd}
cc \ar[r,shift left=1pt,harpoon,"a"] \ar[d,shift left=1pt,harpoon,"r"] &
ct \ar[l,shift left=1pt,harpoon,"b"] \ar[d,shift left=1pt,harpoon,"r"]
\\
tc \ar[r,shift left=1pt,harpoon,"a"] \ar[u,shift left=1pt,harpoon,"l"] &
tt \ar[l,shift left=1pt,harpoon,"b"] \ar[u,shift left=1pt,harpoon,"l"]
\end{tikzcd}
```
:::
::::::

#### bend

https://tex.stackexchange.com/questions/354480/very-curved-arrow-with-tikzcd

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A
\ar[r]
\ar[r, bend right]
\ar[r, bend left, blue]
\ar[r, bend left=10]
\ar[r, bend left=20]
\ar[r, bend left=50]
\ar[r, bend left=100]
& B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='arrow bend'}
\usetikzlibrary{cd}
\begin{tikzcd}
A
\ar[r]
\ar[r, bend right]
\ar[r, bend left, blue]
\ar[r, bend left=10]
\ar[r, bend left=20]
\ar[r, bend left=50]
\ar[r, bend left=100]
& B
\end{tikzcd}
```
:::
::::::

#### rounded corners

https://tex.stackexchange.com/questions/295428/rounded-arrow-in-tikzcd-with-text-on-it

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,positioning}
\begin{tikzcd}[sep=large, execute at end picture={\node[below = 1mm of tikz@f@1-1-2] {$\scriptstyle  g' \circ g$};}]
    B \ar{r}[swap]{g\vphantom{'}}
    \ar[to path={--([yshift=-4ex]\tikztostart.south)
                 -|(\tikztotarget)},rounded corners=12pt]{rr}
    & B' \ar{r}[swap]{g'} & B''
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='rounded corners'}
\usetikzlibrary{cd,positioning}
\begin{tikzcd}[sep=large, execute at end picture={\node[below = 1mm of tikz@f@1-1-2] {$\scriptstyle  g' \circ g$};}]
    B \ar{r}[swap]{g\vphantom{'}}
    \ar[to path={--([yshift=-4ex]\tikztostart.south)
                 -|(\tikztotarget)},rounded corners=12pt]{rr}
    & B' \ar{r}[swap]{g'} & B''
\end{tikzcd}
```
:::
::::::

#### crossing over

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
 A \arrow[dr] & B \arrow[dl, crossing over] \\
 C
 & D
 \end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='crossing over'}
\usetikzlibrary{cd}
\begin{tikzcd}
 A \arrow[dr] & B \arrow[dl, crossing over] \\
 C
 & D
 \end{tikzcd}
```
:::
::::::

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}[row sep=scriptsize, column sep=scriptsize]
 & f^* E_V \arrow[dl] \arrow[rr] \arrow[dd] & & E_V \arrow[dl] \arrow[dd] \\
 f^* E \arrow[rr, crossing over] \arrow[dd] & & E \\
 & U \arrow[dl] \arrow[rr] & & V \arrow[dl] \\
 M \arrow[rr] & & N \arrow[from=uu, crossing over]\\
 \end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='pseudo 3D diagram'}
\usetikzlibrary{cd}
\begin{tikzcd}[row sep=scriptsize, column sep=scriptsize]
 & f^* E_V \arrow[dl] \arrow[rr] \arrow[dd] & & E_V \arrow[dl] \arrow[dd] \\
 f^* E \arrow[rr, crossing over] \arrow[dd] & & E \\
 & U \arrow[dl] \arrow[rr] & & V \arrow[dl] \\
 M \arrow[rr] & & N \arrow[from=uu, crossing over]\\
 \end{tikzcd}
```
:::
::::::

#### tweaking to paths

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[dr, controls={+(1.5,0.5) and +(-1,0.8)}]
    \arrow[dr, dashed, to path=|- (\tikztotarget)]
    & \\
    & B \arrow[loop right]
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='crossing over'}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[dr, controls={+(1.5,0.5) and +(-1,0.8)}]
    \arrow[dr, dashed, to path=|- (\tikztotarget)]
    & \\
    & B \arrow[loop right]
\end{tikzcd}
```
:::
::::::

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[r]
    & B \arrow[r]
        \arrow[d, phantom,""{coordinate,name=Z}]
    & C \arrow[dll,
               "\delta",
               rounded corners,
               to path={--([xshift=2ex]\tikztostart.east)
                        |-(Z)[near end]\tikztonodes
                        -|([xshift=-2ex]\tikztotarget.west)
                        --(\tikztotarget)}] \\
  D \arrow[r]
    & E \arrow[r]
    & F
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='crossing over'}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[r]
    & B \arrow[r]
        \arrow[d, phantom,""{coordinate,name=Z}]
    & C \arrow[dll,
               "\delta",
               rounded corners,
               to path={--([xshift=2ex]\tikztostart.east)
                        |-(Z)[near end]\tikztonodes
                        -|([xshift=-2ex]\tikztotarget.west)
                        --(\tikztotarget)}] \\
  D \arrow[r]
    & E \arrow[r]
    & F
\end{tikzcd}
```
:::
::::::

***

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[r]
    & B \arrow[r]
    & C \arrow[dll,
               "\delta",
               rounded corners,
               to path={--([xshift=2ex]\tikztostart.east)
                        --([xshift=-2ex]\tikztotarget.west)
                        --(\tikztotarget)}] \\
  D \arrow[r]
    & E \arrow[r]
    & F
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='crossing over'}
\usetikzlibrary{cd}
\begin{tikzcd}
  A \arrow[r]
    & B \arrow[r]
    & C \arrow[dll,
               "\delta",
               rounded corners,
               to path={--([xshift=2ex]\tikztostart.east)
                        --([xshift=-2ex]\tikztotarget.west)
                        --(\tikztotarget)}] \\
  D \arrow[r]
    & E \arrow[r]
    & F
\end{tikzcd}
```
:::
::::::

### label

https://tex.stackexchange.com/questions/477733/two-labels-up-and-down-for-same-arrow

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}%              V
A \ar[r, "\alpha", "\beta"'] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='labels over and under the arrow'}
\usetikzlibrary{cd}
\begin{tikzcd}%              V
A \ar[r, "\alpha", "\beta"'] & B
\end{tikzcd}
```
:::
::::::

#### single quotation mark to swap

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='label over the arrow without single quotation'}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"] & B
\end{tikzcd}
```
:::
::::::

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"'] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='label under the arrow with single quotation'}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"'] & B
\end{tikzcd}
```
:::
::::::

#### swap

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='label over the arrow or no swap'}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha"] & B
\end{tikzcd}
```
:::
::::::

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha", swap] & B
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '75%', fig.cap='label swap under the arrow'}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar[r, "\alpha", swap] & B
\end{tikzcd}
```
:::
::::::

#### every arrow swap

https://tex.stackexchange.com/questions/515697/how-can-we-change-the-positions-of-the-labels-on-a-tikz-cd-commutative-diagram

`every arrow/.append style={swap}`

### box in TikZ-CD

https://stackoverflow.com/questions/50392582/box-in-commutative-diagram

https://tex.stackexchange.com/questions/360083/filling-of-diagrams-using-tikzcd/360152#360152

### examples

https://tex.stackexchange.com/questions/218274/how-can-i-draw-commutative-diagrams-in-latex

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar{r}{\varphi} \ar[swap]{d}{\varrho_f} & B \ar{d}{\varrho_g} \\%
A_f \ar{r}{\varphi_f}& B_g
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 0'}
\usetikzlibrary{cd}
\begin{tikzcd}
A \ar{r}{\varphi} \ar[swap]{d}{\varrho_f} & B \ar{d}{\varrho_g} \\%
A_f \ar{r}{\varphi_f}& B_g
\end{tikzcd}
```
:::
::::::

https://tex.stackexchange.com/questions/402572/how-to-make-a-commutative-diagram-using-tikz-cd

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}[row sep=huge]
A \ar[r,"\phi_n"] \ar[d,swap,"\pi"] &
A_n \ar[r,"\epsilon_{i,n}"] \ar[dl,swap,"\psi_n"] \ar[dr,"\pi_n"] &
M_{n_k}(\mathbb{C}) \ar[d,"\eta_{i,n}"]
\\
B(H) & & B(H_n) \ar[ll,dashed]
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 1',engine.opts=list(extra.preamble=c("\\usepackage{amssymb}"))}
\usetikzlibrary{cd}
\begin{tikzcd}[row sep=huge]
A \ar[r,"\phi_n"] \ar[d,swap,"\pi"] &
A_n \ar[r,"\epsilon_{i,n}"] \ar[dl,swap,"\psi_n"] \ar[dr,"\pi_n"] &
M_{n_k}(\mathbb{C}) \ar[d,"\eta_{i,n}"]
\\
B(H) & & B(H_n) \ar[ll,dashed]
\end{tikzcd}
```
:::
::::::

https://tex.stackexchange.com/questions/115783/how-to-draw-commutative-diagrams

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd,arrows}
\begin{tikzcd}
A \ar{r}{f} \ar[swap]{dr}{g\circ f} & B \ar{d}{g} \\
& C
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 2'}
\usetikzlibrary{cd,arrows}
\begin{tikzcd}
A \ar{r}{f} \ar[swap]{dr}{g\circ f} & B \ar{d}{g} \\
& C
\end{tikzcd}
```
:::
::::::

https://tex.stackexchange.com/questions/414292/tikz-commutative-diagrams-compiling-and-best-practice

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
C \ar[d, "\langle f_i \rangle_{i \in I}"', dotted] \ar[rd, "f_i"] &  \\
\prod_{i \in I} A_i \ar[r, "\pi_i"'] & A_i
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 3'}
\usetikzlibrary{cd}
\begin{tikzcd}
C \ar[d, "\langle f_i \rangle_{i \in I}"', dotted] \ar[rd, "f_i"] &  \\
\prod_{i \in I} A_i \ar[r, "\pi_i"'] & A_i
\end{tikzcd}
```
:::
::::::

https://tex.stackexchange.com/questions/483801/mapping-arrows-in-commutative-diagrams

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
a \ar[r, maps to] & \phi(a) &[-2em] \\[-4ex]
R\ar[r, hook, "\phi"] \ar[dr, hook, "\iota"] & K & \phi(a) \\
& Q(R) \ar[u, dotted, "\exists ! \Phi"'] & \frac{a}{1} \ar[u, maps to] \\[-10ex]
a \ar[dr, maps to] \\
& \frac{a}{1} & 
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 4'}
\usetikzlibrary{cd}
\begin{tikzcd}
a \ar[r, maps to] & \phi(a) &[-2em] \\[-4ex]
R\ar[r, hook, "\phi"] \ar[dr, hook, "\iota"] & K & \phi(a) \\
& Q(R) \ar[u, dotted, "\exists ! \Phi"'] & \frac{a}{1} \ar[u, maps to] \\[-10ex]
a \ar[dr, maps to] \\
& \frac{a}{1} & 
\end{tikzcd}
```
:::
::::::

if not using `tikz-cd`,

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{arrows,positioning}
\begin{tikzpicture}
\node (r) at (0,0) {$R$};
\node (k) at (3,0) {$K$};
\node (q) at (3,-3) {$Q(R)$};
\node (ra) at (0,.8) {$a$};
\node (ka) at (3,.8) {$\phi(a)$};
\node (kb) at (4,0) {$\phi(k)$};
\node (qb) at (4,-3) {$\frac a1$};
\path (q) node[below left=1em and 1em] (qc) {$\frac a1$};
\path (r) node[below left=1em and 1em] (rc) {$a$};
\draw[right hook->] (r)--(k) node[midway,above] {$\scriptstyle\phi$};
\draw[right hook->] (r)--(q) node[midway,above right] {$\scriptstyle\iota$};
\draw[dotted,->] (q)--(k) node[midway,right] {$\scriptstyle\exists!\Phi$};
\draw[|->] (ra) edge (ka) (qb) edge (kb) (rc) edge (qc);
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 4 not using tikz-cd'}
\usetikzlibrary{arrows,positioning}
\begin{tikzpicture}
\node (r) at (0,0) {$R$};
\node (k) at (3,0) {$K$};
\node (q) at (3,-3) {$Q(R)$};
\node (ra) at (0,.8) {$a$};
\node (ka) at (3,.8) {$\phi(a)$};
\node (kb) at (4,0) {$\phi(k)$};
\node (qb) at (4,-3) {$\frac a1$};
\path (q) node[below left=1em and 1em] (qc) {$\frac a1$};
\path (r) node[below left=1em and 1em] (rc) {$a$};
\draw[right hook->] (r)--(k) node[midway,above] {$\scriptstyle\phi$};
\draw[right hook->] (r)--(q) node[midway,above right] {$\scriptstyle\iota$};
\draw[dotted,->] (q)--(k) node[midway,right] {$\scriptstyle\exists!\Phi$};
\draw[|->] (ra) edge (ka) (qb) edge (kb) (rc) edge (qc);
\end{tikzpicture}
```
:::
::::::

separated diagrams with `tikz-cd`,

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.45\textwidth}"}
```{tex}
\usetikzlibrary{cd}
\begin{tikzcd}
R\ar[r, hook, "\phi"] \ar[dr, hook, "\iota"] & K &
a \ar[r, maps to,"\phi"] \ar[dr, maps to,"\iota"] & \phi(a) \\
& Q(R) \ar[u, dotted, "\exists ! \Phi"'] &
& \frac{a}{1} \ar[u, maps to,"\Phi"']
\end{tikzcd}
```
:::

::: {.col data-latex="{0.10\textwidth}"}
\ 
:::

::: {.col data-latex="{0.45\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='commutative diagram example 4'}
\usetikzlibrary{cd}
\begin{tikzcd}
R\ar[r, hook, "\phi"] \ar[dr, hook, "\iota"] & K &
a \ar[r, maps to,"\phi"] \ar[dr, maps to,"\iota"] & \phi(a) \\
& Q(R) \ar[u, dotted, "\exists ! \Phi"'] &
& \frac{a}{1} \ar[u, maps to,"\Phi"']
\end{tikzcd}
```
:::
::::::

register map

https://tex.stackexchange.com/questions/510088/tikz-register-map

### put a commutative diagram in a nodepart in a tikzpicture

https://tex.stackexchange.com/questions/524955/put-a-commutative-diagram-in-a-nodepart-in-a-tikzpicture/525046#525046