# linear algebra

## The Art of Linear Algeba

https://github.com/kenjihiranabe/The-Art-of-Linear-Algebra

https://stackoverflow.com/questions/39173714/r-markdown-can-i-insert-a-pdf-to-the-r-markdown-file-as-an-image

https://stackoverflow.com/a/39177166

::: {show-in="html"}
![ArtLA](The-Art-of-Linear-Algebra-zh-TW.pdf){width=100% height=1000}
:::

::: {show-in="pdf"}
\begin{center} <br>
\includegraphics[width=8in]{The-Art-of-Linear-Algebra-zh-TW.pdf} <br>
\end{center}
:::

## CCJou

https://www.youtube.com/playlist?list=PLP-JUp2VR1LsFtHT-i_vZ3oNFIAc3t_Ju

### coordinate

https://www.youtube.com/watch?v=eMUFexQsKXA&list=PLP-JUp2VR1LsFtHT-i_vZ3oNFIAc3t_Ju&index=20

$$
\begin{aligned}
\boldsymbol{v}= & v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}+v^{{\scriptscriptstyle 2}}\boldsymbol{v}_{{\scriptscriptstyle 2}}+\cdots+v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}\\
= & v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}+\cdots+v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}\\
= & c^{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+\cdots+c^{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=c_{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+\cdots+c_{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=B\boldsymbol{c}=B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
= & e^{{\scriptscriptstyle 1}}\hat{\boldsymbol{\beta}}_{{\scriptscriptstyle 1}}+\cdots+e^{{\scriptscriptstyle n}}\hat{\boldsymbol{\beta}}_{{\scriptscriptstyle n}}=e_{{\scriptscriptstyle 1}}\hat{\boldsymbol{\beta}}_{{\scriptscriptstyle 1}}+\cdots+e_{{\scriptscriptstyle n}}\hat{\boldsymbol{\beta}}_{{\scriptscriptstyle n}}\\
= & e^{{\scriptscriptstyle 1}}\hat{\boldsymbol{e}}_{{\scriptscriptstyle 1}}+\cdots+e^{{\scriptscriptstyle n}}\hat{\boldsymbol{e}}_{{\scriptscriptstyle n}}=e_{{\scriptscriptstyle 1}}\hat{\boldsymbol{e}}_{{\scriptscriptstyle 1}}+\cdots+e_{{\scriptscriptstyle n}}\hat{\boldsymbol{e}}_{{\scriptscriptstyle n}}\\
\overset{\text{e.g.}}{=} & e^{{\scriptscriptstyle 1}}\begin{pmatrix}1\\
\vdots\\
0\\
\vdots\\
0
\end{pmatrix}_{{\scriptscriptstyle n\times1}}+\cdots+e^{{\scriptscriptstyle n}}\begin{pmatrix}0\\
\vdots\\
0\\
\vdots\\
1
\end{pmatrix}_{{\scriptscriptstyle n\times1}}=e_{{\scriptscriptstyle 1}}\begin{pmatrix}1\\
\vdots\\
0\\
\vdots\\
0
\end{pmatrix}+\cdots+e_{{\scriptscriptstyle n}}\begin{pmatrix}0\\
\vdots\\
0\\
\vdots\\
1
\end{pmatrix}\\
= & \begin{pmatrix}1 & \cdots & 0 & \cdots & 0\\
\vdots & \ddots & \vdots & \ddots & \vdots\\
0 & \cdots & 1 & \cdots & 0\\
\vdots & \ddots & \vdots & \ddots & \vdots\\
0 & \cdots & 0 & \cdots & 1
\end{pmatrix}_{{\scriptscriptstyle n\times n}}\begin{pmatrix}e^{{\scriptscriptstyle 1}}\\
\vdots\\
e^{{\scriptscriptstyle j}}\\
\vdots\\
e^{{\scriptscriptstyle n}}
\end{pmatrix}_{{\scriptscriptstyle n\times1}}=I_{{\scriptscriptstyle n\times n}}\boldsymbol{e}=I\boldsymbol{e}
\end{aligned}
$$

$$
\begin{aligned}
B\boldsymbol{c}= & c^{{\scriptscriptstyle 1}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{pmatrix}+\cdots+c^{{\scriptscriptstyle j}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle j}}\\
|
\end{pmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{pmatrix}=\begin{pmatrix}| &  & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle j}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & | &  & |
\end{pmatrix}\begin{pmatrix}c^{{\scriptscriptstyle 1}}\\
\vdots\\
c^{{\scriptscriptstyle j}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{pmatrix}\\
= & c^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+c^{{\scriptscriptstyle j}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle j}}\\
|
\end{bmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle j}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & | &  & |
\end{bmatrix}\begin{bmatrix}c^{{\scriptscriptstyle 1}}\\
\vdots\\
c^{{\scriptscriptstyle j}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{bmatrix}\\
= & c^{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+c^{{\scriptscriptstyle 2}}\boldsymbol{\beta}_{{\scriptscriptstyle 2}}+\cdots+c^{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=c^{{\scriptscriptstyle 1}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{pmatrix}+c^{{\scriptscriptstyle 2}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 2}}\\
|
\end{pmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{pmatrix}=\begin{pmatrix}| & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \boldsymbol{\beta}_{{\scriptscriptstyle 2}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| & | &  & |
\end{pmatrix}\begin{pmatrix}c^{{\scriptscriptstyle 1}}\\
c^{{\scriptscriptstyle 2}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{pmatrix}\\
= & c^{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+c^{{\scriptscriptstyle 2}}\boldsymbol{\beta}_{{\scriptscriptstyle 2}}+\cdots+c^{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=c^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+c^{{\scriptscriptstyle 2}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 2}}\\
|
\end{bmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{bmatrix}=\begin{bmatrix}| & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \boldsymbol{\beta}_{{\scriptscriptstyle 2}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| & | &  & |
\end{bmatrix}\begin{bmatrix}c^{{\scriptscriptstyle 1}}\\
c^{{\scriptscriptstyle 2}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{bmatrix}\\
= & c^{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+\cdots+c^{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=c^{{\scriptscriptstyle 1}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{pmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{pmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{pmatrix}=\begin{pmatrix}| &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & |
\end{pmatrix}\begin{pmatrix}c^{{\scriptscriptstyle 1}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{pmatrix}=B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
= & c^{{\scriptscriptstyle 1}}\boldsymbol{\beta}_{{\scriptscriptstyle 1}}+\cdots+c^{{\scriptscriptstyle n}}\boldsymbol{\beta}_{{\scriptscriptstyle n}}=c^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+c^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}c^{{\scriptscriptstyle 1}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}\\
B= & \begin{pmatrix}| &  & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle j}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & | &  & |
\end{pmatrix}=\begin{bmatrix}| &  & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle j}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & | &  & |
\end{bmatrix}=\begin{bmatrix}| & | &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \boldsymbol{\beta}_{{\scriptscriptstyle 2}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| & | &  & |
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{\beta}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{\beta}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & \boldsymbol{c}=\begin{bmatrix}c^{{\scriptscriptstyle 1}}\\
\vdots\\
c^{{\scriptscriptstyle n}}
\end{bmatrix}=\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}\in F^{n} = \mathbb{F}^n \in\left\{ \mathbb{R}^{n},\mathbb{C}^{n},\cdots\right\} 
\end{aligned}
$$

***

$$
\boldsymbol{v}=B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}
$$

***

$$
\begin{aligned}
\boldsymbol{v}= & B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
B^{-1}\boldsymbol{v}= & B^{-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=I\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{-1}\boldsymbol{v}
\end{aligned}
$$

***

$$
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{-1}\boldsymbol{v}
$$

$$
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{-1}\boldsymbol{v}\in\in F^{n}\in\left\{ \mathbb{R}^{n},\mathbb{C}^{n},\cdots\right\} 
$$

***

$$
\boldsymbol{v}=B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}
$$

***

$$
\begin{aligned}
B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=I\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}=I\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}=B^{\prime-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B^{\prime-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B^{\prime-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}
\end{aligned}
$$

### linear transformation

$$
\begin{aligned}
\boldsymbol{v}= & v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}+\cdots+v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}=v^{{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}=\sum_{j=1}^{n}v^{{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}\\
= & v^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+v^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}=V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
, & \begin{cases}
V=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix},\mathfrak{V}=\left\{ \boldsymbol{v}_{{\scriptscriptstyle j}}\right\} _{{\scriptscriptstyle j}=1}^{{\scriptscriptstyle n}}=\left\{ \boldsymbol{v}_{{\scriptscriptstyle 1}},\cdots,\boldsymbol{v}_{{\scriptscriptstyle n}}\right\} \\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}\in F^{n}\in\left\{ \mathbb{R}^{n},\mathbb{C}^{n},\cdots\right\} 
\end{cases}\overset{\text{if }V\text{ invertible}}{\implies}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=V^{-1}\boldsymbol{v}
\end{aligned}
$$

$$
\backslash\text{mathfrak}V=\mathfrak{V}\text{ vs.}\backslash\text{mathfrak}B=\mathfrak{B}
$$

$$
\begin{aligned}
 & \boldsymbol{v}\in\mathcal{V}\overset{T}{\rightarrow}\mathcal{W}\ni\boldsymbol{w}=T\left(\boldsymbol{v}\right)\\
 & ,T:\mathcal{V}\rightarrow\mathcal{W}\begin{cases}
T\left(\boldsymbol{u}+\boldsymbol{v}\right)=T\left(\boldsymbol{u}\right)+T\left(\boldsymbol{v}\right) & \left(a\right)\text{additivity}\\
T\left(\lambda\boldsymbol{v}\right)=\lambda T\left(\boldsymbol{v}\right) & \left(h\right)\text{homogeneity}
\end{cases}
\end{aligned}
$$

$$
\begin{aligned}
\boldsymbol{w}=T\left(\boldsymbol{v}\right)= & T\left(v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}+\cdots+v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\overset{\left(a\right)}{=}T\left(v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)+\cdots+T\left(v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\overset{\left(h\right)}{=}v^{{\scriptscriptstyle 1}}T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)+\cdots+v^{{\scriptscriptstyle n}}T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
= & v^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\\
|
\end{bmatrix}+\cdots+v^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
\end{aligned}
$$

$$
\boldsymbol{w}=T\left(\boldsymbol{v}\right)=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=T\left(V\right)\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}},T\left(V\right)=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}
$$

***

$$
\boldsymbol{w}=T\left(\boldsymbol{v}\right)\overset{\boldsymbol{v}=V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}}{=}T\left(V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\right)\overset{T\text{ linear}}{=}T\left(V\right)\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
$$

***

$$
\begin{aligned}
T\left(\boldsymbol{v}\right)=\boldsymbol{w}= & w^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+w^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=w^{{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}=\sum_{j=1}^{n}w^{{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}\\
= & w^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+w^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}w^{{\scriptscriptstyle 1}}\\
\vdots\\
w^{{\scriptscriptstyle m}}
\end{bmatrix}=W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}\overset{\boldsymbol{w}=T\left(\boldsymbol{v}\right)}{=}W\left[T\left(\boldsymbol{v}\right)\right]_{{\scriptscriptstyle W}}\\
, & \begin{cases}
W=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix},\mathfrak{W}=\left\{ \boldsymbol{w}_{{\scriptscriptstyle j}}\right\} _{{\scriptscriptstyle j}=1}^{{\scriptscriptstyle m}}=\left\{ \boldsymbol{w}_{{\scriptscriptstyle 1}},\cdots,\boldsymbol{w}_{{\scriptscriptstyle m}}\right\} \\
\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=\begin{bmatrix}w^{{\scriptscriptstyle 1}}\\
\vdots\\
w^{{\scriptscriptstyle m}}
\end{bmatrix}\in F^{m}\in\left\{ \mathbb{R}^{m},\mathbb{C}^{m},\cdots\right\} 
\end{cases}\overset{\text{if }W\text{ invertible}}{\implies}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=W^{-1}\boldsymbol{w}
\end{aligned}
$$

$$
\backslash\text{mathfrak}W=\mathfrak{W}\text{ vs.}\backslash\text{mathfrak}V=\mathfrak{V}
$$

$$
\begin{aligned}
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)= & t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}|\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}}\\
|
\end{bmatrix}=W\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}}\\
\vdots\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)= & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}|\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
|
\end{bmatrix}=W\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}
\end{aligned}
$$

$$
\begin{aligned}
T\left(V\right)=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}}
\end{bmatrix} & \cdots & \begin{bmatrix}t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}=WT\\
, & \begin{cases}
W=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\\
T=\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}
\end{cases}
\end{aligned}
$$

$$
T\left(V\right)=\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}=WT,\begin{cases}
W=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\\
T=\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}
\end{cases}
$$

$$
T=\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}\overset{V=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}}{=}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}
$$

***

$$
\begin{aligned}
T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}= & W^{{\scriptscriptstyle -1}}T\left(V\right)=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}^{{\scriptscriptstyle -1}}T\left(V\right),\text{if }W\text{ invertible}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}^{{\scriptscriptstyle -1}}T\left(\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\right)\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}^{{\scriptscriptstyle -1}}\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}\\
= & W^{{\scriptscriptstyle -1}}\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}=\begin{bmatrix}| &  & |\\
W^{{\scriptscriptstyle -1}}T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & W^{{\scriptscriptstyle -1}}T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}
\end{aligned}
$$

***

$$
T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}^{{\scriptscriptstyle -1}}\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}=W^{{\scriptscriptstyle -1}}T\left(V\right)
$$

***

$$
\begin{aligned}
T= & \begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{t}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{t}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix},\boldsymbol{t}_{{\scriptscriptstyle j}}=\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle j}}\right)\right]_{{\scriptscriptstyle W}}
\end{aligned}
$$

$$
\begin{aligned}
W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=\boldsymbol{w}=T\left(\boldsymbol{v}\right)= & T\left(V\right)\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{T\left(V\right)=WT}{=}WT\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=W\left(T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\right)\\
W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}= & W\left(T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\right)\\
\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=I\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=W^{{\scriptscriptstyle -1}}W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}= & W^{{\scriptscriptstyle -1}}W\left(T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\right)=IT\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}},\text{if }W\text{ invertible}\\
\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}= & T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\left[T\left(\boldsymbol{v}\right)\right]_{{\scriptscriptstyle W}}\overset{\boldsymbol{w}=T\left(\boldsymbol{v}\right)}{=}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}= & T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\left[T\left(\boldsymbol{v}\right)\right]_{{\scriptscriptstyle W}}= & T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}}{=}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
\end{aligned}
$$

***

$$
\begin{aligned}
\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}= & T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
=\left[T\left(\boldsymbol{v}\right)\right]_{{\scriptscriptstyle W}}= & \left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
\end{aligned}
$$

***

$$
\begin{aligned}
T=\begin{bmatrix}| &  & |\\
\boldsymbol{t}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{t}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}= & \begin{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}}
\end{bmatrix} & \cdots & \begin{bmatrix}t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}\end{bmatrix}=\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix},\begin{array}{c}
\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=T\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\Updownarrow\\
w^{{\scriptscriptstyle i}}=t_{{\scriptscriptstyle j}}^{{\scriptscriptstyle i}}v^{{\scriptscriptstyle j}}=t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}v^{{\scriptscriptstyle j}}
\end{array}\\
= & \begin{bmatrix}\begin{bmatrix}t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle 1}}\\
\vdots\\
t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle 1}}
\end{bmatrix} & \cdots & \begin{bmatrix}t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle n}}\\
\vdots\\
t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle n}}
\end{bmatrix}\end{bmatrix}=\begin{bmatrix}t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle 1}} & \cdots & t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle n}}\\
\vdots & \ddots & \vdots\\
t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle 1}} & \cdots & t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle n}}
\end{bmatrix}=\left[t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\right]_{{\scriptscriptstyle m\times n}}=\left[t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\right]=t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\\
= & \begin{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 11}}\\
\vdots\\
t_{{\scriptscriptstyle m1}}
\end{bmatrix} & \cdots & \begin{bmatrix}t_{{\scriptscriptstyle 1n}}\\
\vdots\\
t_{{\scriptscriptstyle mn}}
\end{bmatrix}\end{bmatrix}=\begin{bmatrix}t_{{\scriptscriptstyle 11}} & \cdots & t_{{\scriptscriptstyle 1n}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle m1}} & \cdots & t_{{\scriptscriptstyle mn}}
\end{bmatrix}=\left[t_{{\scriptscriptstyle ij}}\right]_{{\scriptscriptstyle m\times n}}=\left[t_{{\scriptscriptstyle ij}}\right]=t_{{\scriptscriptstyle ij}}
\end{aligned}
$$

***

$$
T=\begin{bmatrix}| &  & |\\
\boldsymbol{t}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{t}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}=\begin{bmatrix}t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle 1}} & \cdots & t^{{\scriptscriptstyle 1}}{}_{{\scriptscriptstyle n}}\\
\vdots & \ddots & \vdots\\
t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle 1}} & \cdots & t^{{\scriptscriptstyle m}}{}_{{\scriptscriptstyle n}}
\end{bmatrix}=\left[t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\right]_{{\scriptscriptstyle m\times n}}=t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}
$$

***

$$
\begin{aligned}
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)= & t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}}
\end{bmatrix}\\
= & W\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}}\\
\vdots\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)= & t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}
\end{bmatrix}\\
= & W\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
\Downarrow\\
\begin{bmatrix}| &  & |\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right) & \cdots & T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\\
| &  & |
\end{bmatrix}= & \begin{array}{c}
t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}}\\
\vdots\\
t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}
\end{array}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}\begin{array}{c}
+\\
\\
+
\end{array}\begin{array}{c}
\cdots\\
\curvearrowright\\
\cdots
\end{array}\begin{array}{c}
+\\
\\
+
\end{array}\begin{array}{c}
t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}}\\
\vdots\\
t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}
\end{array}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}| &  & |\\
\boldsymbol{t}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{t}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}=WT\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}| &  & |\\
\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle 1}}\right)\right]_{{\scriptscriptstyle W}} & \cdots & \left[T\left(\boldsymbol{v}_{{\scriptscriptstyle n}}\right)\right]_{{\scriptscriptstyle W}}\\
| &  & |
\end{bmatrix}=W\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\\
T\left(V\right)= & WT=W\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}
\end{aligned}
$$

***

$$
\begin{aligned}
WT= & T\left(V\right)\\
\Downarrow\\
T= & W^{{\scriptscriptstyle -1}}T\left(V\right)
\end{aligned}
$$

***

$$
T=\begin{bmatrix}t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}{}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}{}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle 1}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
t_{{\scriptscriptstyle 1}}^{{\scriptscriptstyle m}} & \cdots & t_{{\scriptscriptstyle n}}^{{\scriptscriptstyle m}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{t}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{t}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}=\left[t_{{\scriptscriptstyle j}}^{{\scriptscriptstyle i}}\right]_{{\scriptscriptstyle m\times n}}=\left[t_{{\scriptscriptstyle j}}{}^{{\scriptscriptstyle i}}\right]_{{\scriptscriptstyle m\times n}}=t_{{\scriptscriptstyle j}}{}^{{\scriptscriptstyle i}}
$$

***

$$
\left[t_{{\scriptscriptstyle j}}{}^{{\scriptscriptstyle i}}\right]_{{\scriptscriptstyle m\times n}}=\left[t_{{\scriptscriptstyle j}}^{{\scriptscriptstyle i}}\right]_{{\scriptscriptstyle m\times n}}=\left[t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\right]_{{\scriptscriptstyle m\times n}},\begin{cases}
w^{{\scriptscriptstyle i}}=t_{{\scriptscriptstyle j}}^{{\scriptscriptstyle i}}v^{{\scriptscriptstyle j}}=t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}v^{{\scriptscriptstyle j}}\\
T\left(\boldsymbol{v}_{{\scriptscriptstyle j}}\right)=t_{{\scriptscriptstyle j}}^{{\scriptscriptstyle i}}\boldsymbol{w}_{{\scriptscriptstyle i}}=t_{{\scriptscriptstyle j}}{}^{{\scriptscriptstyle i}}\boldsymbol{w}_{{\scriptscriptstyle i}}
\end{cases}
$$

***

$$
T=\left[t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\right]_{{\scriptscriptstyle m\times n}}=t^{{\scriptscriptstyle i}}{}_{{\scriptscriptstyle j}}\text{ is the matrix representation of the linear transformation }T\left(\cdot\right):\mathcal{V}\rightarrow\mathcal{W}
$$

***

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='coordinate under linear transformation',engine.opts=list(extra.preamble=c("\\usepackage{amssymb}"))}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[
	row sep=30,
	column sep=1,
	every label/.append style={font=\small} % \tiny \scriptsize \footnotesize \small \normalsize \large \Large \LARGE \huge \Huge
]
%11
||
&
%12
\mathcal{V}
\ar[l,white,text=black,"\left\{ \boldsymbol{v}_{{\scriptscriptstyle j}}\right\} _{{\scriptscriptstyle j}=1}^{{\scriptscriptstyle n}}=\mathfrak{V}\subseteq" description]
\ar[r,-stealth,"T(\cdot)"]
&
%13
\mathcal{W}
\ar[r,white,text=black,"\supseteq\mathfrak{W}=\left\{ \boldsymbol{w}_{{\scriptscriptstyle j}}\right\} _{{\scriptscriptstyle j}=1}^{{\scriptscriptstyle m}}" description]
&
%14
||
\\
%21
V\left[\boldsymbol{v}\right]_{\scriptscriptstyle V}=
\ar[d,white,text=black,"VT^{\scriptscriptstyle -1}W^{\scriptscriptstyle -1}\boldsymbol{w}=" description]
&
%22
\boldsymbol{v}
\ar[l,white,text=black,"v^{{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}=" description]
\ar[u,white,text=black,"\in" rotate=90,near end]
\ar[r,-{Stealth[harpoon,swap]},shift right=0.75,"T(\cdot)",swap]
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"V^{\scriptscriptstyle -1}",swap]
&
%23
\boldsymbol{w}
\ar[u,white,text=black,"\in" rotate=90,near end]
\ar[l,-{Stealth[harpoon,swap]},shift right=0.75,"T^{\scriptscriptstyle -1}(\cdot)",swap]
\ar[r,white,text=black,"=w^{{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}" description]
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"W^{\scriptscriptstyle-1}",swap]
&
%24
=W\left[\boldsymbol{w}\right]_{\scriptscriptstyle W}=T(\boldsymbol{v})=T\left(V\right)\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
\ar[d,white,text=black,"=WTV^{\scriptscriptstyle -1}\boldsymbol{v}" description]
\\
%31
||
&
%32
\left[\boldsymbol{v}\right]_{\scriptscriptstyle V}
\ar[l,white,text=black,"V^{\scriptscriptstyle -1}\boldsymbol{v}=" description,near start]
\ar[r,-{Stealth[harpoon,swap]},shift right=0.75,"T",swap]
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"V",swap]
\ar[d,white,text=black,"\in" rotate=-90,near start]
\ar[drr,-,xshift=22,yshift=-5]
\ar[r,-{Stealth[harpoon,swap]},bend left=50]
&
%33
\left[\boldsymbol{w}\right]_{\scriptscriptstyle W}
\ar[r,white,text=black,"=W^{\scriptscriptstyle -1}\boldsymbol{w}" description]
\ar[l,-{Stealth[harpoon,swap]},shift right=0.75,"T^{\scriptscriptstyle -1}",swap]
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"W",swap]
\ar[d,white,text=black,"\in" rotate=-90,near start]
&
%34
=\left[T(\boldsymbol{v})\right]_{\scriptscriptstyle W}=T\left[\boldsymbol{v}\right]_{\scriptscriptstyle V}
\ar[d,-,xshift=20,yshift=0]
\\
%41
\boldsymbol{t}_{{\scriptscriptstyle j}}=\left[T\left(\boldsymbol{v}_{{\scriptscriptstyle j}}\right)\right]_{{\scriptscriptstyle W}}
&
%42
\mathbb{F}^n
\ar[r,-stealth,"T"]
\ar[d,white,text=black,"\in" rotate=-90,near start]
&
%43
\mathbb{F}^m
\ar[d,white,text=black,"\in" rotate=-90,near start]
&
%44
T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}
\\
%51
||
&
%52
\left\{\mathbb{R}^n,\mathbb{C}^m,\cdots\right\}
&
%53
\left\{\mathbb{R}^m,\mathbb{C}^m,\cdots\right\}
\end{tikzcd}
```

### change of basis

https://www.youtube.com/watch?v=WAtLPk55ljM&list=PLP-JUp2VR1LsFtHT-i_vZ3oNFIAc3t_Ju&index=22

$$
\boldsymbol{v}=B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}
$$

***

$$
\begin{aligned}
B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=I\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}=B^{-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}= & B^{-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}\\
B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}=I\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}=B^{\prime-1}B^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B^{\prime-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B^{\prime}}}= & B^{\prime-1}B\left[\boldsymbol{v}\right]_{{\scriptscriptstyle B}}
\end{aligned}
$$

***

$$
\begin{aligned}
\boldsymbol{v}= & v^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}+\cdots+v^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}=v^{{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}=\sum_{j=1}^{n}v^{{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}=v^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+v^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v^{{\scriptscriptstyle 1}}\\
\vdots\\
v^{{\scriptscriptstyle n}}
\end{bmatrix}=V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
= & v^{\prime{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}+\cdots+v^{\prime{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}=v^{\prime{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}^{\prime}=\sum_{j=1}^{n}v^{\prime{\scriptscriptstyle j}}\boldsymbol{v}_{{\scriptscriptstyle j}}^{\prime}=v^{\prime{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}\\
|
\end{bmatrix}+\cdots+v^{\prime{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v^{\prime{\scriptscriptstyle 1}}\\
\vdots\\
v^{\prime{\scriptscriptstyle n}}
\end{bmatrix}=V^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\\
\boldsymbol{v}= & V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=V^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}
\end{aligned}
$$

$$
\begin{aligned}
\boldsymbol{w}= & w^{{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}+\cdots+w^{{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}=w^{{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}=\sum_{j=1}^{m}w^{{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}=w^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}\\
|
\end{bmatrix}+\cdots+w^{{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}\\
| &  & |
\end{bmatrix}\begin{bmatrix}w^{{\scriptscriptstyle 1}}\\
\vdots\\
w^{{\scriptscriptstyle m}}
\end{bmatrix}=W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}\\
= & w^{\prime{\scriptscriptstyle 1}}\boldsymbol{w}_{{\scriptscriptstyle 1}}^{\prime}+\cdots+w^{\prime{\scriptscriptstyle m}}\boldsymbol{w}_{{\scriptscriptstyle m}}^{\prime}=w^{\prime{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}^{\prime}=\sum_{j=1}^{m}w^{\prime{\scriptscriptstyle j}}\boldsymbol{w}_{{\scriptscriptstyle j}}^{\prime}=w^{\prime{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}^{\prime}\\
|
\end{bmatrix}+\cdots+w^{\prime{\scriptscriptstyle m}}\begin{bmatrix}|\\
\boldsymbol{w}_{{\scriptscriptstyle m}}^{\prime}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{w}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{w}_{{\scriptscriptstyle m}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}w^{\prime{\scriptscriptstyle 1}}\\
\vdots\\
w^{\prime{\scriptscriptstyle m}}
\end{bmatrix}=W^{\prime}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W^{\prime}}}\\
\boldsymbol{w}= & W\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}=W^{\prime}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W^{\prime}}}
\end{aligned}
$$

$$
\begin{aligned}
\boldsymbol{v}_{{\scriptscriptstyle 1}}= & v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}+\cdots+v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}=v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}\\
|
\end{bmatrix}+\cdots+v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle 1}}\\
\vdots\\
v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}}
\end{bmatrix}=V^{\prime}\left[\boldsymbol{v}_{{\scriptscriptstyle 1}}\right]_{{\scriptscriptstyle V^{\prime}}}\\
\vdots\\
\boldsymbol{v}_{{\scriptscriptstyle n}}= & v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle 1}}\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}+\cdots+v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle n}}\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}=v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle 1}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}\\
|
\end{bmatrix}+\cdots+v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle n}}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
|
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}}\\
\vdots\\
v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle n}}
\end{bmatrix}=V^{\prime}\left[\boldsymbol{v}_{{\scriptscriptstyle n}}\right]_{{\scriptscriptstyle V^{\prime}}}\\
\Downarrow\\
\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}\\
| &  & |
\end{bmatrix}= & \begin{array}{c}
v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle 1}}\\
\vdots\\
v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle 1}}
\end{array}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime}\\
|
\end{bmatrix}\begin{array}{c}
+\\
\\
+
\end{array}\begin{array}{c}
\cdots\\
\curvearrowright\\
\cdots
\end{array}\begin{array}{c}
+\\
\\
+
\end{array}\begin{array}{c}
v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}}\\
\vdots\\
v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle n}}
\end{array}\begin{bmatrix}|\\
\boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
|
\end{bmatrix}\\
= & \begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle 1}} & \cdots & v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle 1}}\\
\vdots & \ddots & \vdots\\
v_{{\scriptscriptstyle 1}}^{\prime}{}^{{\scriptscriptstyle n}} & \cdots & v_{{\scriptscriptstyle n}}^{\prime}{}^{{\scriptscriptstyle n}}
\end{bmatrix}=\begin{bmatrix}| &  & |\\
\boldsymbol{v}_{{\scriptscriptstyle 1}}^{\prime} & \cdots & \boldsymbol{v}_{{\scriptscriptstyle n}}^{\prime}\\
| &  & |
\end{bmatrix}\begin{bmatrix}| &  & |\\
\left[\boldsymbol{v}_{{\scriptscriptstyle 1}}\right]_{{\scriptscriptstyle V^{\prime}}} & \cdots & \left[\boldsymbol{v}_{{\scriptscriptstyle n}}\right]_{{\scriptscriptstyle V^{\prime}}}\\
| &  & |
\end{bmatrix}=V^{\prime}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\\
V= & V^{\prime}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\overset{\text{if }V^{\prime}\text{ invertible}}{\Longleftrightarrow}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}=V^{\prime{\scriptscriptstyle -1}}V
\end{aligned}
$$

***

$$
\left[V\right]_{{\scriptscriptstyle V^{\prime}}}=V^{\prime{\scriptscriptstyle -1}}V:\left[V^{\prime}\middle|V\right]\overset{\text{Gauss-Jordan delimination}}{\longrightarrow}\left[I\middle|V^{\prime{\scriptscriptstyle -1}}V\right]=\left[I\middle|\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\right]
$$

$$
\left[V^{\prime}\right]_{{\scriptscriptstyle V}}=V^{{\scriptscriptstyle -1}}V^{\prime}:\left[V\middle|V^{\prime}\right]\overset{\text{Gauss-Jordan delimination}}{\longrightarrow}\left[I\middle|V^{{\scriptscriptstyle -1}}V^{\prime}\right]=\left[I\middle|\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\right]
$$

***

$$
\begin{cases}
\left[V\right]_{{\scriptscriptstyle V^{\prime}}}=V^{\prime{\scriptscriptstyle -1}}V: & \left[V^{\prime}\middle|V\right]\overset{\text{Gauss-Jordan delimination}}{\longrightarrow}\left[I\middle|V^{\prime{\scriptscriptstyle -1}}V\right]=\left[I\middle|\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\right]\\
\left[V^{\prime}\right]_{{\scriptscriptstyle V}}=V^{{\scriptscriptstyle -1}}V^{\prime}: & \left[V\middle|V^{\prime}\right]\overset{\text{Gauss-Jordan delimination}}{\longrightarrow}\left[I\middle|V^{{\scriptscriptstyle -1}}V^{\prime}\right]=\left[I\middle|\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\right]
\end{cases}
$$

***

$$
\begin{aligned}
\boldsymbol{v}= & V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=V^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\\
= & V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}},\wedge V=V^{\prime}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\\
= & V^{\prime}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
V^{\prime}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & V^{\prime}\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & \left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
\end{aligned}
$$

***

$$
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}=\left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}
$$

***

symmetrically,

$$
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}=\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}
$$

***

```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '100%', fig.cap='change of coordinate basis under linear transformation',engine.opts=list(extra.preamble=c("\\usepackage{amssymb}"))}
\usetikzlibrary{cd,arrows.meta}
\begin{tikzcd}[
	row sep=50,
	column sep=60,
	every label/.append style={font=\small} % \tiny \scriptsize \footnotesize \small \normalsize \large \Large \LARGE \huge \Huge
]
%11
V[\boldsymbol{v}]_{\scriptscriptstyle V} = \boldsymbol{v} \in \mathcal{V} \supseteq \begin{cases}
\mathfrak{V}=\left\{ \boldsymbol{v}_{{\scriptscriptstyle j}}\right\} \\
\mathfrak{V}^{\prime}=\left\{ \boldsymbol{v}_{{\scriptscriptstyle j}}^{\prime}\right\} 
\end{cases}
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"V^{\scriptscriptstyle -1}",swap]
\ar[dd,-{Stealth[harpoon,swap]},bend right=60,"V^{\prime \scriptscriptstyle -1}",swap,near start]
\ar[r,-{Stealth[harpoon,swap]},shift right=0.75,"T(\cdot)",swap]
&
%12
T(\boldsymbol{v}) = \boldsymbol{w} \in \mathcal{W} \supseteq \begin{cases}
\mathfrak{W}=\left\{ \boldsymbol{w}_{{\scriptscriptstyle j}}\right\} \\
\mathfrak{W}^{\prime}=\left\{ \boldsymbol{w}_{{\scriptscriptstyle j}}^{\prime}\right\} 
\end{cases}
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"W^{\scriptscriptstyle -1}",swap]
\ar[dd,-{Stealth[harpoon,swap]},bend left=55,"W^{\prime \scriptscriptstyle -1}",swap,near start]
\ar[l,-{Stealth[harpoon,swap]},shift right=0.75,"T^{\scriptscriptstyle -1}(\cdot)",swap]
\\
%21
V^{\scriptscriptstyle -1}\boldsymbol{v} = \left[\boldsymbol{v}\right]_{\scriptscriptstyle V} \in \mathbb{F}^n
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"\text{[}V\text{]}_{\scriptscriptstyle V^\prime}",swap]
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"V",swap]
\ar[r,-{Stealth[harpoon,swap]},shift right=0.75,"T=\text{[}T\left(V\right)\text{]}_{{\scriptscriptstyle W}}",swap]
\ar[r,-{Stealth[harpoon,swap]},bend left=45]
&
%22
\left[T(\boldsymbol{v})\right]_{\scriptscriptstyle W} = \left[\boldsymbol{w}\right]_{\scriptscriptstyle W} \in \mathbb{F}^m
\ar[d,-{Stealth[harpoon,swap]},shift right=0.75,"\text{[}W\text{]}_{\scriptscriptstyle W^\prime}",swap]
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"W",swap]
\ar[l,-{Stealth[harpoon,swap]},shift right=0.75,"T^{\scriptscriptstyle -1}",swap]
\\
%31
V^{\prime \scriptscriptstyle -1}\boldsymbol{v} = \left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}} = \left[\boldsymbol{v}\right]_{\scriptscriptstyle V^\prime} \in \mathbb{F}^n
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"\text{[}V^\prime\text{]}_{\scriptscriptstyle V}",swap]
\ar[uu,-{Stealth[harpoon,swap]},bend left=55,"V^\prime",swap,near end]
\ar[r,-{Stealth[harpoon,swap]},shift right=0.75,"T^\prime=\text{[}T\left(V^\prime\right)\text{]}_{{\scriptscriptstyle W^\prime}}",swap]
\ar[r,-{Stealth[harpoon,swap]},bend left=45]
&
%32
\left[T(\boldsymbol{v})\right]_{\scriptscriptstyle W^\prime} = \left[\boldsymbol{w}\right]_{\scriptscriptstyle W^\prime} \in \mathbb{F}^m
\ar[u,-{Stealth[harpoon,swap]},shift right=0.75,"\text{[}W^\prime\text{]}_{\scriptscriptstyle W}",swap]
\ar[uu,-{Stealth[harpoon,swap]},bend right=60,"W^\prime",swap,near end]
\ar[l,-{Stealth[harpoon,swap]},shift right=0.75,"T^{\prime \scriptscriptstyle -1}",swap]
\end{tikzcd}
```

$\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\rightleftharpoons\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}$

$$
\begin{aligned}
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{\left[V\right]_{{\scriptscriptstyle V^{\prime}}}}{\rightarrow}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & \left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{\left[V^{\prime}\right]_{{\scriptscriptstyle V}}}{\leftarrow}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & \left[V^{\prime}\right]_{{\scriptscriptstyle V}}^{{\scriptscriptstyle -1}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\Downarrow\\
\left[V\right]_{{\scriptscriptstyle V^{\prime}}}= & \left[V^{\prime}\right]_{{\scriptscriptstyle V}}^{{\scriptscriptstyle -1}}\Leftrightarrow\left[V^{\prime}\right]_{{\scriptscriptstyle V}}=\left[V\right]_{{\scriptscriptstyle V^{\prime}}}^{{\scriptscriptstyle -1}}
\end{aligned}
$$

$\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\rightarrow\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}$

$$
\begin{aligned}
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{\left[V\right]_{{\scriptscriptstyle V^{\prime}}}}{\rightarrow}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & \left[V\right]_{{\scriptscriptstyle V^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{V}{\rightarrow}\boldsymbol{v}\overset{V^{\prime{\scriptscriptstyle -1}}}{\rightarrow}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}= & V^{\prime{\scriptscriptstyle -1}}V\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\\
\Downarrow\\
\left[V\right]_{{\scriptscriptstyle V^{\prime}}}= & V^{\prime{\scriptscriptstyle -1}}V\overset{\left[V^{\prime}\right]_{{\scriptscriptstyle V}}=\left[V\right]_{{\scriptscriptstyle V^{\prime}}}^{{\scriptscriptstyle -1}}}{\Longleftrightarrow}\left[V^{\prime}\right]_{{\scriptscriptstyle V}}=\left(V^{\prime{\scriptscriptstyle -1}}V\right)^{{\scriptscriptstyle -1}}=V^{{\scriptscriptstyle -1}}V^{\prime}
\end{aligned}
$$

$\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\rightarrow\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W^{\prime}}}$

$$
\begin{aligned}
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\overset{T^{\prime}=\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}}{\rightarrow}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W^{\prime}}}= & \left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\\
\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\overset{\left[V^{\prime}\right]_{{\scriptscriptstyle V}}}{\rightarrow}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V}}\overset{T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}}{\rightarrow}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W}}\overset{\left[W\right]_{{\scriptscriptstyle W^{\prime}}}}{\rightarrow}\left[\boldsymbol{w}\right]_{{\scriptscriptstyle W^{\prime}}}= & \left[W\right]_{{\scriptscriptstyle W^{\prime}}}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\left[\boldsymbol{v}\right]_{{\scriptscriptstyle V^{\prime}}}\\
\Downarrow\\
\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}= & \left[W\right]_{{\scriptscriptstyle W^{\prime}}}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\\
= & W^{\prime{\scriptscriptstyle -1}}W\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}V^{{\scriptscriptstyle -1}}V^{\prime}\\
= & W^{\prime{\scriptscriptstyle -1}}WW^{{\scriptscriptstyle -1}}T\left(V\right)V^{{\scriptscriptstyle -1}}V^{\prime}\\
= & W^{\prime{\scriptscriptstyle -1}}IT\left(V\right)V^{{\scriptscriptstyle -1}}V^{\prime}\\
= & W^{\prime{\scriptscriptstyle -1}}T\left(V\right)V^{{\scriptscriptstyle -1}}V^{\prime}
\end{aligned}
$$

***

$\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}\rightleftharpoons\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}$

$$
\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}=\left[W\right]_{{\scriptscriptstyle W^{\prime}}}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[V^{\prime}\right]_{{\scriptscriptstyle V}}
$$

***

$$
\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}=W^{\prime{\scriptscriptstyle -1}}T\left(V\right)V^{{\scriptscriptstyle -1}}V^{\prime}
$$

***

$$
\begin{aligned}
W^{\prime}\left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}= & T\left(V\right)\\
\Downarrow\\
\left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}= & W^{\prime{\scriptscriptstyle -1}}T\left(V\right)
\end{aligned}
$$

***

$$
\begin{aligned}
\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}= & W^{\prime{\scriptscriptstyle -1}}T\left(V\right)V^{{\scriptscriptstyle -1}}V^{\prime}\\
= & \left\{ W^{\prime{\scriptscriptstyle -1}}T\left(V\right)\right\} \left\{ V^{{\scriptscriptstyle -1}}V^{\prime}\right\} \\
= & \left\{ \left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}\right\} \left\{ V^{{\scriptscriptstyle -1}}V^{\prime}\right\} \\
= & \left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}V^{{\scriptscriptstyle -1}}V^{\prime}\\
, & \begin{cases}
\left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}=W^{\prime{\scriptscriptstyle -1}}T\left(V\right) & :W^{\prime}\left[T\left(V\right)\right]_{{\scriptscriptstyle W^{\prime}}}=T\left(V\right)\\
V^{{\scriptscriptstyle -1}}V^{\prime} & :\left[V\middle|V^{\prime}\right]\overset{\text{Gauss-Jordan delimination}}{\longrightarrow}\left[I\middle|V^{{\scriptscriptstyle -1}}V^{\prime}\right]
\end{cases}
\end{aligned}
$$

***

$T^{\prime}\rightleftharpoons T$

$$
\begin{aligned}
\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}= & \left[W\right]_{{\scriptscriptstyle W^{\prime}}}\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\\
T^{\prime}=\left[T\left(V^{\prime}\right)\right]_{{\scriptscriptstyle W^{\prime}}}\Updownarrow & T=\left[T\left(V\right)\right]_{{\scriptscriptstyle W}}\\
T^{\prime}= & \left[W\right]_{{\scriptscriptstyle W^{\prime}}}T\left[V^{\prime}\right]_{{\scriptscriptstyle V}}\\
= & \left[W^{\prime{\scriptscriptstyle -1}}W\right]T\left[V^{{\scriptscriptstyle -1}}V^{\prime}\right]\\
T^{\prime}= & W^{\prime{\scriptscriptstyle -1}}WTV^{{\scriptscriptstyle -1}}V^{\prime}
\end{aligned}
$$

***

$$
T^{\prime}=W^{\prime{\scriptscriptstyle -1}}WTV^{{\scriptscriptstyle -1}}V^{\prime}
$$

### singular value decomposition

https://www.youtube.com/watch?v=oPHM-ZWWvlg&list=PLP-JUp2VR1LsFtHT-i_vZ3oNFIAc3t_Ju&index=42

https://www.youtube.com/watch?v=zagHmMPZyoo&list=PLP-JUp2VR1LsFtHT-i_vZ3oNFIAc3t_Ju&index=43

## Chi, Chen-Yu

https://www.youtube.com/playlist?list=PLJWAeYEa8SXBej3kuQMz8vV41VabZUILb