# Hilbert space

```{r}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

## Taylor expansion or Taylor series

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV15s4y1g7HQ&autoplay=0"></iframe>
:::

```{lemma}
Newton-Leibniz formula = N-LF, equivalent to first mean value theorem for derivatives / Lagrange mean value theorem = MVTdL / MVTd1
```

::: {show-in="html"}
牛頓-萊布尼茨公式 Newton-Leibniz formula = N-LF, equivalent to 第一微分均值定理 / 拉格朗日均值定理 first mean value theorem for derivatives / Lagrange mean
value theorem = MVTdL / MVTd1 \ref{thm:MVTd1} \label{thm:N-LF}
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
牛頓-萊布尼茨公式 Newton-Leibniz formula = N-LF, equivalent to 第一微分均值定理 / 拉格朗日均值定理 first mean value theorem for derivatives / Lagrange mean
value theorem = MVTdL / MVTd1 \ref{thm:MVTd1} \label{thm:N-LF}
\end{CJK}
:::

\begin{equation}
\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)\label{eq:N-LF}
\end{equation}

$$
\begin{aligned}
 & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\\
\begin{cases}
f^{\prime}:\left[a,b\right]\rightarrow\mathbb{R}\,\left(4\right) & a<b\,\left(3\right)\\
f^{\prime}\text{ continuous on }\left[a,b\right] & \left(5\right)
\end{cases}\Rightarrow\exists c\in\left(a,b\right)\left(f^{\prime}\left(c\right)=\dfrac{\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x}{b-a}\right) & \text{MVTi1 }\ref{thm:MVTi1}
\end{cases}\\
\Rightarrow & \begin{array}{c}
\exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right\} \,\ref{eq:MVTd1fromN-LFandMVTi1}\\
\Updownarrow\\
\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)\,\ref{eq:N-LFfromMVTd1}
\end{array}
\end{aligned}
$$

$\left(\Downarrow\right)$:

$$
\begin{aligned}
 & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\ref{thm:Rolle}}{\Rightarrow}\exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right\}  & \text{MVTd1 }\ref{thm:MVTd1}\\
\left[a,b\right]=\left[x_{0},x_{n}\right]=\left[x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right] & \ref{eq:-72}
\end{cases}\nonumber \\
 & \overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)\label{eq:N-LFfromMVTd1}
\end{aligned}
$$

$\left(\Uparrow\right)$:

$$
\begin{aligned}
 & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right) & \text{N-LF }\ref{eq:N-LF}\\
\begin{cases}
f^{\prime}:\left[a,b\right]\rightarrow\mathbb{R}\,\left(4\right) & a<b\,\left(3\right)\\
f^{\prime}\text{ continuous on }\left[a,b\right] & \left(5\right)
\end{cases}\Rightarrow\exists c\in\left(a,b\right)\left(f^{\prime}\left(c\right)=\dfrac{\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x}{b-a}\right) & \text{MVTi1 }\ref{thm:MVTi1}
\end{cases}\nonumber \\
\Rightarrow & \exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right\} \label{eq:MVTd1fromN-LFandMVTi1}
\end{aligned}
$$

Proof: $\left(\Downarrow\right)$


::: {show-in="html"}
重疊端點分割 $\left[a,b\right]$ 成 $n$ 部分聯集 $\left(n\in\mathbb{N}\right)$
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
重疊端點分割 $\left[a,b\right]$ 成 $n$ 部分聯集 $\left(n\in\mathbb{N}\right)$
\end{CJK}
:::

\begin{equation}
\left[a,b\right]=\left[x_{0},x_{n}\right]=\left[x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right]\label{eq:-72}
\end{equation}

$$
\begin{aligned}
 & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\ref{thm:Rolle}}{\Rightarrow}\exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right\}  & \text{MVTd1 }\ref{thm:MVTd1}\\
\left[a,b\right]=\left[x_{0},x_{n}\right]=\left[x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right] & \ref{eq:-72}
\end{cases}\\
\Rightarrow & \begin{cases}
\left(f:\left[x_{0},x_{n}\right]\rightarrow\mathbb{R}\right)=\left(f:\left[x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right]\rightarrow\mathbb{R}\right) & \Leftarrow\left(0\right)\\
f\text{ continuous on }\left[x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right] & \Leftarrow\left(1\right)\\
f\text{ differentiable on }\left(x_{0},x_{1}\right]\cup\left[x_{0},x_{2}\right]\cup\cdots\cup\left[x_{n-1},x_{n}\right) & \Leftarrow\left(2\right)
\end{cases}\\
\Rightarrow & \begin{cases}
f:\left[x_{0},x_{1}\right]\rightarrow\mathbb{R},f:\left[x_{0},x_{2}\right]\rightarrow\mathbb{R},\cdots,f:\left[x_{n-1},x_{n}\right]\rightarrow\mathbb{R} & \Leftarrow\left(0\right)\\
f\text{ continuous on }\left[x_{0},x_{1}\right],\left[x_{0},x_{2}\right],\cdots,\left[x_{n-1},x_{n}\right] & \Leftarrow\left(1\right)\\
f\text{ differentiable on }\left(x_{0},x_{1}\right],\left[x_{0},x_{2}\right],\cdots,\left[x_{n-1},x_{n}\right)\Rightarrow f\text{ differentiable on }\left(x_{0},x_{1}\right),\left(x_{0},x_{2}\right),\cdots,\left(x_{n-1},x_{n}\right) & \Leftarrow\left(2\right)
\end{cases}\\
 & \overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\begin{cases}
\exists c_{1}\in\left(x_{0},x_{1}\right)\left\{ f^{\prime}\left(c_{1}\right)\left(x_{1}-x_{0}\right)=f\left(x_{1}\right)-f\left(x_{0}\right)\right\} \\
\exists c_{2}\in\left(x_{1},x_{2}\right)\left\{ f^{\prime}\left(c_{2}\right)\left(x_{2}-x_{1}\right)=f\left(x_{2}\right)-f\left(x_{1}\right)\right\} \\
\vdots\\
\exists c_{k}\in\left(x_{k-1},x_{k}\right)\left\{ f^{\prime}\left(c_{k}\right)\left(x_{k}-x_{k-1}\right)=f\left(x_{k}\right)-f\left(x_{k-1}\right)\right\}  & \forall k\in\mathbb{N}\cap\left[1,n\right]\\
\vdots\\
\exists c_{n}\in\left(x_{n-1},x_{n}\right)\left\{ f^{\prime}\left(c_{n}\right)\left(x_{n}-x_{n-1}\right)=f\left(x_{n}\right)-f\left(x_{n-1}\right)\right\} 
\end{cases}\\
\Rightarrow & f^{\prime}\left(c_{k}\right)\left(x_{k}-x_{k-1}\right)=f\left(x_{k}\right)-f\left(x_{k-1}\right)\\
\Rightarrow & \sum_{k}f^{\prime}\left(c_{k}\right)\left(x_{k}-x_{k-1}\right)=\sum_{k}f\left(x_{k}\right)-f\left(x_{k-1}\right)\\
\Rightarrow & \sum_{k=1}^{n}f^{\prime}\left(c_{k}\right)\left(x_{k}-x_{k-1}\right)=\sum_{k=1}^{n}f\left(x_{k}\right)-f\left(x_{k-1}\right)\\
 & =\left[f\left(x_{1}\right)-f\left(x_{0}\right)\right]+\left[f\left(x_{2}\right)-f\left(x_{1}\right)\right]+\cdots+\left[f\left(x_{n}\right)-f\left(x_{n-1}\right)\right]\\
 & =f\left(x_{n}\right)-f\left(x_{0}\right)=f\left(b\right)-f\left(a\right)\\
 & \Rightarrow\sum_{k=1}^{n}f^{\prime}\left(c_{k}\right)\left(x_{k}-x_{k-1}\right)=f\left(b\right)-f\left(a\right)\overset{\Delta x_{k}=x_{k}-x_{k-1}}{\Leftrightarrow}\sum_{k=1}^{n}f^{\prime}\left(c_{k}\right)\Delta x_{k}=f\left(b\right)-f\left(a\right)\\
\Rightarrow & \lim_{n\rightarrow\infty}\lim_{\Delta x_{k}=\frac{x_{n}-x_{0}}{n}}\sum_{k=1}^{n}f^{\prime}\left(c_{k}\right)\Delta x_{k}=\lim_{n\rightarrow\infty}\lim_{\Delta x_{k}=\frac{b-a}{n}\rightarrow0}f\left(b\right)-f\left(a\right)=f\left(b\right)-f\left(a\right)\\
\Rightarrow & \int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=\int_{x_{0}}^{x_{n}}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)\\
\Rightarrow & \int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)
\end{aligned}
$$

$$
\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)
$$


::: {show-in="html"}
得到 牛頓-萊布尼茨公式 Newton-Leibniz formula = N-LF [eq:N-LF]
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
得到 牛頓-萊布尼茨公式 Newton-Leibniz formula = N-LF [eq:N-LF]
\end{CJK}
:::

$$
\tag*{$\Box$}
$$

Proof: $\left(\Uparrow\right)$

$$
\begin{aligned}
 & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right) & \text{N-LF }\ref{eq:N-LF}\\
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & a<b\,\left(3\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)
\end{cases}\Rightarrow\exists c\in\left(a,b\right)\left(f\left(c\right)=\dfrac{\int_{a}^{b}f\left(x\right)\mathrm{d}x}{b-a}\right) & \text{MVTi1 }\ref{thm:MVTi1}
\end{cases}\\
\Rightarrow & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right) & \text{N-LF }\ref{eq:N-LF}\\
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & a<b\,\left(3\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)
\end{cases}\Rightarrow\exists c\in\left(a,b\right)\left\{ f\left(c\right)\left(b-a\right)=\int_{a}^{b}f\left(x\right)\mathrm{d}x\right\}  & \text{MVTi1 }\ref{thm:MVTi1}
\end{cases}\\
\Rightarrow & \begin{cases}
\begin{cases}
f:\left[a,b\right]\rightarrow\mathbb{R} & \left(0\right)\\
f\text{ continuous on }\left[a,b\right] & \left(1\right)\\
f\text{ differentiable on }\left(a,b\right) & \left(2\right)
\end{cases}\overset{\stackrel{\ref{thm:Rolle}}{\ref{thm:MVTd1}}}{\Rightarrow}\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right) & \text{N-LF }\ref{eq:N-LF}\\
\begin{cases}
\text{if }f^{\prime}:\left[a,b\right]\rightarrow\mathbb{R} & \left(4\right)\\
\text{if }f^{\prime}\text{ continuous on }\left[a,b\right] & \left(5\right)
\end{cases}\Rightarrow\exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x\right\}  & \text{MVTi1 }\ref{thm:MVTi1}
\end{cases}\\
\Rightarrow & \exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x=f\left(b\right)-f\left(a\right)\right\} \\
\Rightarrow & \exists c\in\left(a,b\right)\left\{ f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right\} 
\end{aligned}
$$

$$
\exists c\in\left(a,b\right)\left[f^{\prime}\left(c\right)\left(b-a\right)=f\left(b\right)-f\left(a\right)\right]
$$

::: {show-in="html"}
得到 第一微分均值定理 / 拉格朗日均值定理 first mean value theorem for derivatives / Lagrange mean value theorem = MVTdL / MVTd1 [thm:MVTd1]
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
得到 第一微分均值定理 / 拉格朗日均值定理 first mean value theorem for derivatives / Lagrange mean value theorem = MVTdL / MVTd1 [thm:MVTd1]
\end{CJK}
:::

$$
\tag*{$\Box$}
$$

$$
\begin{aligned}
\int_{a}^{b}f^{\prime}\left(x\right)\mathrm{d}x= & f\left(b\right)-f\left(a\right)\\
\int_{x_{{\scriptscriptstyle 0}}}^{x}f^{\prime}\left(t\right)\mathrm{d}t= & f\left(x\right)-f\left(x_{{\scriptscriptstyle 0}}\right)\\
f\left(x\right)= & f\left(x_{{\scriptscriptstyle 0}}\right)+\int_{x_{{\scriptscriptstyle 0}}}^{x}f^{\prime}\left(t\right)\mathrm{d}t
\end{aligned}
$$

***

$$
f\left(x\right)=f\left(x_{{\scriptscriptstyle 0}}\right)+\int_{x_{{\scriptscriptstyle 0}}}^{x}f^{\prime}\left(t\right)\mathrm{d}t
$$

***

```{lemma}
univariable product rule
```

$$
\mathrm{d}\left(uv\right)=\left(\mathrm{d}u\right)v+u\mathrm{d}v=v\mathrm{d}u+u\mathrm{d}v
$$

$$
\frac{\mathrm{d}\left(uv\right)}{\mathrm{d}t}=v\frac{\mathrm{d}u}{\mathrm{d}t}+u\frac{\mathrm{d}v}{\mathrm{d}t}
$$

```{lemma}
integration by parts
```

\begin{eqnarray}
\int\frac{\mathrm{d}\left(uv\right)}{\mathrm{d}t}\thinspace\mathrm{d}t & = & \int v\frac{\mathrm{d}u}{\mathrm{d}t}+u\frac{\mathrm{d}v}{\mathrm{d}t}\mathrm{d}t\nonumber \\
\int\mathrm{d}\left(uv\right)= &  & =\int v\frac{\mathrm{d}u}{\mathrm{d}t}\mathrm{d}t+\int u\frac{\mathrm{d}v}{\mathrm{d}t}\mathrm{d}t\nonumber \\
uv= &  & =\int v\mathrm{d}u+\int u\mathrm{d}v\nonumber \\
\int u\frac{\mathrm{d}v}{\mathrm{d}t}\mathrm{d}t & = & uv-\int v\frac{\mathrm{d}u}{\mathrm{d}t}\mathrm{d}t\text{ "switching derivatives"}\label{eq:switchingDerivatives}\\
\int u\mathrm{d}v & = & uv-\int v\mathrm{d}u\text{ "switching differentials"}\label{eq:switchingDifferentials}
\end{eqnarray}

```{theorem}
univariable Taylor theorem with the remainder in integral form
```

$$
f\left(x\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t
$$

$$
f\left(x\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

Proof:

\begin{eqnarray}
 &  & f\left(x\right)-f\left(a\right)\\
 & \overset{\ref{thm:N-LF}}{=} & \int_{a}^{x}f'\left(t\right)\mathrm{d}t=\int_{a}^{x}f'\left(t\right)\frac{\mathrm{d}\left(t-x\right)}{\mathrm{d}t}\mathrm{d}t\overset{\ref{eq:switchingDerivatives}}{=}\left[f'\left(t\right)\left(t-x\right)\right]_{t=a}^{x}-\int_{a}^{x}\left(t-x\right)\frac{\mathrm{d}f'\left(t\right)}{\mathrm{d}t}\mathrm{d}t\\
 & = & \left[-f'\left(a\right)\left(a-x\right)\right]-\int_{a}^{x}f''\left(t\right)\left(t-x\right)\mathrm{d}t=f'\left(a\right)\left(x-a\right)-\int_{a}^{x}f''\left(t\right)\frac{\mathrm{d}\frac{\left(t-x\right)^{2}}{2}}{\mathrm{d}t}\mathrm{d}t\\
 & = & f'\left(a\right)\left(x-a\right)-\left(\left[f''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\right]_{t=a}^{x}-\int_{a}^{x}\frac{\left(t-x\right)^{2}}{2}\frac{\mathrm{d}f''\left(t\right)}{\mathrm{d}t}\mathrm{d}t\right)\\
 & = & f'\left(a\right)\left(x-a\right)-\left(\left[-f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}\right]-\int_{a}^{x}f'''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\mathrm{d}t\right)\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\int_{a}^{x}f'''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\mathrm{d}t\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\int_{a}^{x}f'''\left(t\right)\frac{\mathrm{d}\frac{\left(t-x\right)^{3}}{2\cdot3}}{\mathrm{d}t}\mathrm{d}t\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\left(\left[f'''\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\right]_{t=a}^{x}-\int_{a}^{x}\frac{\left(t-x\right)^{3}}{2\cdot3}\frac{\mathrm{d}f'''\left(t\right)}{\mathrm{d}t}\mathrm{d}t\right)\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\left(\left[f'''\left(a\right)\frac{\left(x-a\right)^{3}}{2\cdot3}\right]-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\mathrm{d}t\right)\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(t\right)\frac{\left(x-a\right)^{3}}{2\cdot3}-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\mathrm{d}t\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(t\right)\frac{\left(x-a\right)^{3}}{2\cdot3}-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\mathrm{d}\frac{\left(t-x\right)^{4}}{2\cdot3\cdot4}}{\mathrm{d}t}\mathrm{d}t\\
 & = & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(t\right)\frac{\left(x-a\right)^{3}}{2\cdot3}\\
 &  & -\left(\left[f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{4}}{2\cdot3\cdot4}\right]_{t=a}^{x}-\int_{a}^{x}\frac{\left(t-x\right)^{4}}{2\cdot3\cdot4}\frac{\mathrm{d}f^{\left(4\right)}\left(t\right)}{\mathrm{d}t}\mathrm{d}t\right)\\
 & = & f'\left(t\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(t\right)\frac{\left(x-a\right)^{3}}{2\cdot3}\\
 &  & -\left(\left[-f^{\left(4\right)}\left(a\right)\frac{\left(x-a\right)^{4}}{2\cdot3\cdot4}\right]-\int_{a}^{x}f^{\left(5\right)}\left(t\right)\frac{\left(t-x\right)^{4}}{2\cdot3\cdot4}\mathrm{d}t\right)\\
 & = & f'\left(t\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(a\right)\frac{\left(x-a\right)^{3}}{2\cdot3}+f^{\left(4\right)}\left(a\right)\frac{\left(x-a\right)^{4}}{2\cdot3\cdot4}\\
 & \vdots & +\int_{a}^{x}f^{\left(5\right)}\left(t\right)\frac{\left(t-x\right)^{4}}{2\cdot3\cdot4}\mathrm{d}t\\
 & = & \sum_{k=1}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t\textrm{ remainder in integral form}
\end{eqnarray}

$$
\tag*{$\Box$}
$$

Proof:

$$
\begin{aligned}
 & f\left(x\right)-f\left(a\right)\\
\overset{\ref{thm:N-LF}}{=} & \int_{a}^{x}f'\left(t\right)\mathrm{d}t=\int_{a}^{x}f'\left(t\right)\frac{\mathrm{d}\left(t-x\right)}{\mathrm{d}t}\mathrm{d}t\overset{\ref{eq:switchingDerivatives}}{=}\left[f'\left(t\right)\left(t-x\right)\right]_{t=a}^{x}-\int_{a}^{x}\left(t-x\right)\frac{\mathrm{d}f'\left(t\right)}{\mathrm{d}t}\mathrm{d}t\\
= & \left[-f'\left(a\right)\left(a-x\right)\right]-\int_{a}^{x}f''\left(t\right)\left(t-x\right)\mathrm{d}t=f'\left(a\right)\left(x-a\right)+\int_{a}^{x}f''\left(t\right)\left(x-t\right)\mathrm{d}t\\
= & f'\left(a\right)\left(x-a\right)-\int_{a}^{x}f''\left(t\right)\left(t-x\right)\mathrm{d}t=\sum\limits _{k=1}^{1}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f''\left(t\right)\left(x-t\right)\mathrm{d}t\\
= & f'\left(a\right)\left(x-a\right)-\int_{a}^{x}f''\left(t\right)\frac{\mathrm{d}\frac{\left(t-x\right)^{2}}{2}}{\mathrm{d}t}\mathrm{d}t\\
= & f'\left(a\right)\left(x-a\right)-\left(\left[f''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\right]_{t=a}^{x}-\int_{a}^{x}\frac{\left(t-x\right)^{2}}{2}\frac{\mathrm{d}f''\left(t\right)}{\mathrm{d}t}\mathrm{d}t\right)\\
= & f'\left(a\right)\left(x-a\right)-\left(\left[-f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}\right]-\int_{a}^{x}f'''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\mathrm{d}t\right)\\
= & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\int_{a}^{x}f'''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\mathrm{d}t\\
= & \sum\limits _{k=1}^{2}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f'''\left(t\right)\frac{\left(t-x\right)^{2}}{2}\mathrm{d}t=\sum\limits _{k=1}^{2}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f'''\left(t\right)\frac{\left(x-t\right)^{2}}{2!}\mathrm{d}t\\
= & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\int_{a}^{x}f'''\left(t\right)\frac{\mathrm{d}\frac{\left(t-x\right)^{3}}{2\cdot3}}{\mathrm{d}t}\mathrm{d}t\\
= & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\left(\left[f'''\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\right]_{t=a}^{x}-\int_{a}^{x}\frac{\left(t-x\right)^{3}}{2\cdot3}\frac{\mathrm{d}f'''\left(t\right)}{\mathrm{d}t}\mathrm{d}t\right)\\
= & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+\left(\left[f'''\left(a\right)\frac{\left(x-a\right)^{3}}{2\cdot3}\right]-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\mathrm{d}t\right)\\
= & f'\left(a\right)\left(x-a\right)+f''\left(a\right)\frac{\left(x-a\right)^{2}}{2}+f'''\left(t\right)\frac{\left(x-a\right)^{3}}{2\cdot3}-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\mathrm{d}t\\
= & \sum\limits _{k=1}^{3}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}-\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(t-x\right)^{3}}{2\cdot3}\mathrm{d}t=\sum\limits _{k=1}^{3}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(4\right)}\left(t\right)\frac{\left(x-t\right)^{3}}{3!}\mathrm{d}t\\
\overset{\vdots}{=} & \sum_{k=1}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
f\left(x\right)-f\left(a\right)= & \sum_{k=1}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
f\left(x\right)= & f\left(a\right)+\sum_{k=1}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \frac{f^{\left(0\right)}\left(a\right)}{0!}\left(x-a\right)^{0}+\sum_{k=1}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
\end{aligned}
$$

$$
\tag*{$\Box$}
$$

```{theorem}
univariable Taylor theorem with the remainder in Lagrange form
```

$$
f\left(x\right)=\sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\frac{f^{\left(n+1\right)}\left(\xi\right)}{\left(n+1\right)!}\left(x-a\right)^{n+1}
$$

Proof:


::: {show-in="html"}
by 連續函數 極值定理 / 最大最小值定理 / 最小最大值定理 continuous function extreme value theorem = CFEVT / extreme value theorem = EVT \ref{thm:EVT} and 連續函數
介值定理 / 中間值定理 continuous function intermediate value theorem = CFIVT / intermediate value theorem = IVT \ref{thm:IVT}
:::

::: {show-in="pdf"}
\begin{CJK}{UTF8}{bsmi}
by 連續函數 極值定理 / 最大最小值定理 / 最小最大值定理 continuous function extreme value theorem = CFEVT / extreme value theorem = EVT \ref{thm:EVT} and 連續函數
介值定理 / 中間值定理 continuous function intermediate value theorem = CFIVT / intermediate value theorem = IVT \ref{thm:IVT}
\end{CJK}
:::

\begin{eqnarray}
\textrm{let }f\left(x\right) & = & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+r_{n}\left(x\right)\\
r_{n}\left(x\right) & = & \int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t\\
\textrm{let }n & \in & \left\{ 2k-1|k\in\mathbb{N}\right\} \\
\textrm{let }f\left(t\right) & \overset{\ref{thm:EVT}}{\in} & \left[m,M\right]\subseteq f\left(\left(a,x\right)\right)\textrm{ when }a<x\\
\int_{a}^{x}m\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t\le & r_{n}\left(x\right) & \le\int_{a}^{x}M\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t\\
m\int_{a}^{x}\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t= &  & =M\int_{a}^{x}\frac{\left(t-x\right)^{n}}{n!}\mathrm{d}t\\
m\left[\frac{\left(t-x\right)^{n+1}}{\left(n+1\right)!}\right]_{t=a}^{x}= &  & =M\left[\frac{\left(t-x\right)^{n+1}}{\left(n+1\right)!}\right]_{t=a}^{x}\\
m\frac{\left(x-a\right)^{n+1}}{\left(n+1\right)!}= &  & =M\frac{\left(x-a\right)^{n+1}}{\left(n+1\right)!}\\
 & \Downarrow & \ref{thm:IVT}\\
r_{n}\left(x\right) & \overset{\exists\xi\in\left(a,x\right)}{=} & f^{\left(n+1\right)}\left(\xi\right)\frac{\left(x-a\right)^{n+1}}{\left(n+1\right)!}\\
f\left(x\right) & = & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+f^{\left(n+1\right)}\left(\xi\right)\frac{\left(x-a\right)^{n+1}}{\left(n+1\right)!}\\
 & = & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\frac{f^{\left(n+1\right)}\left(\xi\right)}{\left(n+1\right)!}\left(x-a\right)^{n+1}
\end{eqnarray}

$$
\tag*{$\Box$}
$$

```{theorem}
univariable Taylor theorem with the remainder in big O form
```

$$
f\left(a+h\right)=\sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}h^{k}+\mathrm{O}\left(h^{n+1}\right)
$$

Proof:

\begin{eqnarray}
\textrm{if }\left|f^{\left(n+1\right)}\left(t\right)\right| & \le & K\thinspace\forall t\in\left(a,x\right)\\
\left|r_{n}\left(x\right)\right|=\left|\frac{f^{\left(n+1\right)}\left(\xi\right)}{\left(n+1\right)!}\left(x-a\right)^{n+1}\right| & \le & \frac{K}{\left(n+1\right)!}\left|x-a\right|^{n+1}\\
r_{n}\left(x\right) & \in & \mathrm{O}\left(\left(x-a\right)^{n+1}\right)\\
\textrm{let }R_{n}\left(h\right) & = & r_{n}\left(a+h\right)\\
R_{n}\left(h\right) & \in & \mathrm{O}\left(h^{n+1}\right)\\
f\left(a+h\right) & = & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}h^{k}+\frac{f^{\left(n+1\right)}\left(\xi\right)}{\left(n+1\right)!}h^{n+1}\\
 & = & \sum_{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}h^{k}+\mathrm{O}\left(h^{n+1}\right)
\end{eqnarray}

$$
\tag*{$\Box$}
$$

***

$$
f\left(x\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$f\left(x\right)=\sin\left(x\right)$

$$
\sin\left(x\right)=\sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$a=0$

$$
\sin\left(x\right)=\sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(0\right)}{k!}\left(x-0\right)^{k}+\int_{0}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$$
\begin{aligned}
\sin\left(x\right)= & \sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(0\right)}{k!}\left(x-0\right)^{k}+\int_{0}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(0\right)}{k!}x^{k}+\int_{0}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \frac{\sin^{\left(0\right)}\left(0\right)}{0!}x^{0}+\frac{\sin^{\left(1\right)}\left(0\right)}{1!}x^{1}+\frac{\sin^{\left(2\right)}\left(0\right)}{2!}x^{2}+\frac{\sin^{\left(3\right)}\left(0\right)}{3!}x^{3}\\
 & +\int_{0}^{x}\sin^{\left(3+1\right)}\left(t\right)\frac{\left(x-t\right)^{3}}{3!}\mathrm{d}t\\
= & \frac{0}{0!}x^{0}+\frac{\cos\left(0\right)}{1!}x^{1}+\frac{-\sin\left(0\right)}{2!}x^{2}+\frac{-\cos\left(0\right)}{3!}x^{3}\\
 & +\int_{0}^{x}\sin^{\left(4\right)}\left(t\right)\frac{\left(x-t\right)^{3}}{3!}\mathrm{d}t\\
= & 0+\frac{1}{1}x+0-\frac{1}{6}x^{3}+\int_{0}^{x}\sin\left(t\right)\frac{\left(x-t\right)^{3}}{6}\mathrm{d}t\\
= & x-\frac{1}{6}x^{3}+\frac{1}{6}\int_{0}^{x}\left(x-t\right)^{3}\sin\left(t\right)\mathrm{d}t\\
\sin\left(x\right)= & x-\frac{1}{6}x^{3}+\frac{1}{6}\int_{0}^{x}\left(x-t\right)^{3}\sin\left(t\right)\mathrm{d}t
\end{aligned}
$$

$$
\begin{aligned}
\sin\left(x\right)= & \sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(0\right)}{k!}\left(x-0\right)^{k}+\int_{0}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{\sin^{\left(k\right)}\left(0\right)}{k!}x^{k}+\int_{0}^{x}\sin^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \frac{\sin^{\left(0\right)}\left(0\right)}{0!}x^{0}+\frac{\sin^{\left(1\right)}\left(0\right)}{1!}x^{1}+\frac{\sin^{\left(2\right)}\left(0\right)}{2!}x^{2}+\frac{\sin^{\left(3\right)}\left(0\right)}{3!}x^{3}+\frac{\sin^{\left(4\right)}\left(0\right)}{4!}x^{4}\\
 & +\int_{0}^{x}\sin^{\left(4+1\right)}\left(t\right)\frac{\left(x-t\right)^{4}}{4!}\mathrm{d}t\\
= & \frac{0}{0!}x^{0}+\frac{\cos\left(0\right)}{1!}x^{1}+\frac{-\sin\left(0\right)}{2!}x^{2}+\frac{-\cos\left(0\right)}{3!}x^{3}+\frac{\sin\left(0\right)}{4!}x^{4}\\
 & +\int_{0}^{x}\sin^{\left(5\right)}\left(t\right)\frac{\left(x-t\right)^{4}}{4!}\mathrm{d}t\\
= & 0+\frac{1}{1}x+0-\frac{1}{6}x^{3}+0+\int_{0}^{x}\cos\left(t\right)\frac{\left(x-t\right)^{4}}{24}\mathrm{d}t\\
= & x-\frac{1}{6}x^{3}+\frac{1}{24}\int_{0}^{x}\left(x-t\right)^{4}\cos\left(t\right)\mathrm{d}t\\
\sin\left(x\right)= & x-\frac{1}{6}x^{3}+\frac{1}{24}\int_{0}^{x}\left(x-t\right)^{4}\cos\left(t\right)\mathrm{d}t
\end{aligned}
$$

$$
\begin{aligned}
\sin\left(x\right)= & x-\frac{1}{6}x^{3}+\frac{1}{6}\int_{0}^{x}\left(x-t\right)^{3}\sin\left(t\right)\mathrm{d}t\\
= & x-\frac{1}{6}x^{3}+\frac{1}{24}\int_{0}^{x}\left(x-t\right)^{4}\cos\left(t\right)\mathrm{d}t\\
\Downarrow\\
\frac{1}{6}\int_{0}^{x}\left(x-t\right)^{3}\sin\left(t\right)\mathrm{d}t= & \frac{1}{24}\int_{0}^{x}\left(x-t\right)^{4}\cos\left(t\right)\mathrm{d}t\\
\Downarrow\\
\int_{0}^{x}\left(x-t\right)^{4}\cos\left(t\right)\mathrm{d}t= & 4\int_{0}^{x}\left(x-t\right)^{3}\sin\left(t\right)\mathrm{d}t
\end{aligned}
$$

***

$$
f\left(x\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}f^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$f\left(x\right)=\mathrm{e}^{x}=\exp\left(x\right)$

$$
\mathrm{e}^{x}=\sum\limits _{k=0}^{n}\frac{\exp^{\left(k\right)}\left(a\right)}{k!}\left(x-a\right)^{k}+\int_{a}^{x}\exp^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$a=0$

$$
\mathrm{e}^{x}=\sum\limits _{k=0}^{n}\frac{\exp^{\left(k\right)}\left(0\right)}{k!}\left(x-0\right)^{k}+\int_{0}^{x}\exp^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t
$$

$$
\begin{aligned}
\mathrm{e}^{x}= & \sum\limits _{k=0}^{n}\frac{\exp^{\left(k\right)}\left(0\right)}{k!}\left(x-0\right)^{k}+\int_{0}^{x}\exp^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{\exp^{\left(k\right)}\left(0\right)}{k!}x^{k}+\int_{0}^{x}\exp^{\left(n+1\right)}\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{\exp\left(0\right)}{k!}x^{k}+\int_{0}^{x}\exp\left(t\right)\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{\mathrm{e}^{0}}{k!}x^{k}+\int_{0}^{x}\mathrm{e}^{t}\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t=\sum\limits _{k=0}^{n}\frac{1}{k!}x^{k}+\int_{0}^{x}\mathrm{e}^{t}\frac{\left(x-t\right)^{n}}{n!}\mathrm{d}t\\
= & \sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\\
\mathrm{e}^{x}= & \sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t
\end{aligned}
$$

$$
\mathrm{e}^{x}=\sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t
$$

***

$\lim\limits _{n\rightarrow\infty}\dfrac{1+n+\dfrac{n^{2}}{2}+\cdots+\dfrac{n^{n}}{n!}}{\mathrm{e}^{n}}$

$$
\begin{aligned}
 & \lim\limits _{n\rightarrow\infty}\dfrac{1+n+\dfrac{n^{2}}{2}+\cdots+\dfrac{n^{n}}{n!}}{\mathrm{e}^{n}},\wedge\mathrm{e}^{x}=\sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\\
= & \lim\limits _{n\rightarrow\infty}\dfrac{1+n+\dfrac{n^{2}}{2}+\cdots+\dfrac{n^{n}}{n!}}{\left[\sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\right]_{x=n}}\\
= & \lim\limits _{n\rightarrow\infty}\dfrac{\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}}{\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}+\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t}=\lim\limits _{n\rightarrow\infty}\dfrac{1}{1+\dfrac{\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t}{\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}}}
\end{aligned}
$$

$$
\begin{aligned}
 & \lim\limits _{n\rightarrow\infty}\dfrac{1+n+\dfrac{n^{2}}{2}+\cdots+\dfrac{n^{n}}{n!}}{\mathrm{e}^{n}},\wedge\mathrm{e}^{x}=\sum\limits _{k=0}^{n}\frac{x^{k}}{k!}+\frac{1}{n!}\int_{0}^{x}\left(x-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\\
= & \lim\limits _{n\rightarrow\infty}\dfrac{\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}}{\mathrm{e}^{n}},\wedge\mathrm{e}^{n}=\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}+\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\Rightarrow\sum\limits _{k=0}^{n}\frac{n^{k}}{k!}=\mathrm{e}^{n}-\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t\\
= & \lim\limits _{n\rightarrow\infty}\dfrac{\mathrm{e}^{n}-\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t}\mathrm{d}t}{\mathrm{e}^{n}}=\lim\limits _{n\rightarrow\infty}\left[1-\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t\right],\text{ if }\lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}\left(t-n\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t\in\mathbb{R}\\
= & \lim\limits _{n\rightarrow\infty}1-\lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t=1-\lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t
\end{aligned}
$$

$\lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t$

$$
\begin{aligned}
 & \lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t,x=n-t\\
= & \lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{t=0}^{t=n}\left(n-t\right)^{n}\mathrm{e}^{t-n}\mathrm{d}t,t=n-x\\
= & \lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{n-x=0}^{n-x=n}\left(n-\left(n-x\right)\right)^{n}\mathrm{e}^{\left(n-x\right)-n}\mathrm{d}\left(n-x\right)\\
= & \lim\limits _{n\rightarrow\infty}\frac{1}{n!}\left(-\int_{x=n}^{x=0}x^{n}\mathrm{e}^{-x}\mathrm{d}x\right)=\lim\limits _{n\rightarrow\infty}\frac{1}{n!}\left(\int_{x=0}^{x=n}x^{n}\mathrm{e}^{-x}\mathrm{d}x\right)\\
= & \lim\limits _{n\rightarrow\infty}\frac{1}{n!}\int_{0}^{n}x^{n}\mathrm{e}^{-x}\mathrm{d}x
\end{aligned}
$$

$\int x^{n}\mathrm{e}^{-x}\mathrm{d}x$

difficult process

$$
\begin{aligned}
\int x^{n}\mathrm{e}^{-x}\mathrm{d}x= & -\int x^{n}\mathrm{d}\mathrm{e}^{-x}\\
= & -\left[x^{n}\mathrm{e}^{-x}-\int\mathrm{e}^{-x}\mathrm{d}x^{n}\right]\\
= & -\left[x^{n}\mathrm{e}^{-x}-\int\mathrm{e}^{-x}nx^{n-1}\mathrm{d}x\right]\\
= & -x^{n}\mathrm{e}^{-x}+n\int x^{n-1}\mathrm{e}^{-x}\mathrm{d}x
\end{aligned}
$$

$$
\int x^{n}\mathrm{e}^{-x}\mathrm{d}x=-x^{n}\mathrm{e}^{-x}+n\int x^{n-1}\mathrm{e}^{-x}\mathrm{d}x
$$

$$
\begin{aligned}
\int_{0}^{n}x^{n}\mathrm{e}^{-x}\mathrm{d}x= & \left[-x^{n}\mathrm{e}^{-x}\right]_{x=0}^{n}+n\int_{0}^{n}x^{n-1}\mathrm{e}^{-x}\mathrm{d}x\\
= & \left[-n^{n}\mathrm{e}^{-n}-\left(-0^{n}\mathrm{e}^{-0}\right)\right]+n\int_{0}^{n}x^{n-1}\mathrm{e}^{-x}\mathrm{d}x\\
= & -n^{n}\mathrm{e}^{-n}+n\int_{0}^{n}x^{n-1}\mathrm{e}^{-x}\mathrm{d}x
\end{aligned}
$$

$$
\int_{0}^{n}x^{n}\mathrm{e}^{-x}\mathrm{d}x=-n^{n}\mathrm{e}^{-n}+n\int_{0}^{n}x^{n-1}\mathrm{e}^{-x}\mathrm{d}x
$$

more formal process with gamma function

https://www.wolframalpha.com/input?i2d=true&i=Divide%5BIntegrate%5BPower%5Bx%2Cn%5DPower%5Be%2C-x%5D%2C%7Bx%2C0%2Cn%7D%5D%2Cn%21%5D

https://www.wolframalpha.com/input?i2d=true&i=Limit%5BDivide%5BIntegrate%5BPower%5Bx%2Cn%5DPower%5Be%2C-x%5D%2C%7Bx%2C0%2Cn%7D%5D%2Cn%21%5D%2Cn-%3E%E2%88%9E%5D

$$
\text{to be proved}
$$

***

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1du411a7qB&autoplay=0"></iframe>
:::

### Elliot Schneider

https://www.youtube.com/watch?v=HQsZG8Yxb7w

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://www.youtube.com/embed/HQsZG8Yxb7w"></iframe>
:::

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=12m25s

$$
f\left(x\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(x_{{\scriptscriptstyle 0}}\right)}{k!}\left(x-x_{{\scriptscriptstyle 0}}\right)^{k}+\frac{f^{\left(n+1\right)}\left(\xi\right)}{\left(n+1\right)!}\left(x-x_{{\scriptscriptstyle 0}}\right)^{n+1}
$$

$$
f\left(x_{{\scriptscriptstyle 0}}+\epsilon\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(x_{{\scriptscriptstyle 0}}\right)}{k!}\epsilon^{k}+\mathrm{O}\left(\epsilon^{n+1}\right),\epsilon=x-x_{{\scriptscriptstyle 0}}
$$

$$
f\left(x+\epsilon\right)=\sum\limits _{n=0}^{\infty}\frac{f^{\left(n\right)}\left(x\right)}{n!}\epsilon^{n}
$$

$$
\begin{aligned}
f\left(x+\epsilon\right)= & \sum\limits _{n=0}^{\infty}\frac{f^{\left(n\right)}\left(x\right)}{n!}\epsilon^{n}=\sum\limits _{n=0}^{\infty}\frac{\dfrac{\mathrm{d}^{n}}{\mathrm{d}x^{n}}f\left(x\right)}{n!}\epsilon^{n}=\sum\limits _{n=0}^{\infty}\frac{\left(\dfrac{\mathrm{d}}{\mathrm{d}x}\right)^{n}f\left(x\right)}{n!}\epsilon^{n}\\
= & \sum\limits _{n=0}^{\infty}\frac{\left(\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}\right)^{n}f\left(x\right)}{n!}=\left(\sum\limits _{n=0}^{\infty}\frac{\left(\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}\right)^{n}}{n!}\right)f\left(x\right)=\mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}f\left(x\right)
\end{aligned}
$$

***

univariable Taylor operator

$$
f\left(x+\epsilon\right)=\mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}f\left(x\right)
$$

$$
\mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}=\sum\limits _{n=0}^{\infty}\frac{\left(\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}\right)^{n}}{n!}=1+\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}+\dfrac{1}{2}\epsilon^{2}\dfrac{\mathrm{d}^{2}}{\mathrm{d}x^{2}}+\cdots
$$

***

$f\left(x\right)=mx+b$,

$$
\begin{aligned}
\mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}f\left(x\right)= & \left(1+\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}+\dfrac{1}{2}\epsilon^{2}\dfrac{\mathrm{d}^{2}}{\mathrm{d}x^{2}}+\cdots\right)\left(mx+b\right)\\
= & 1\left(mx+b\right)+\epsilon\dfrac{\mathrm{d}}{\mathrm{d}x}\left(mx+b\right)+\dfrac{1}{2}\epsilon^{2}\dfrac{\mathrm{d}^{2}}{\mathrm{d}x^{2}}\left(mx+b\right)+\cdots\\
= & \left(mx+b\right)+\left(\epsilon m\right)+\left(\dfrac{1}{2}\epsilon^{2}0\right)+\left(0+0+\cdots\right)\\
= & \left(mx+b\right)+\left(\epsilon m\right)+\left(0+0+\cdots\right)\\
= & \left(mx+b\right)+\left(\epsilon m\right)+0\\
= & mx+b+m\epsilon=m\left(x+\epsilon\right)+b=f\left(x+\epsilon\right)
\end{aligned}
$$

***

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=15m34s

multivariable ...

$$
\begin{aligned}
f\left(x+\epsilon\right)= & \mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}f\left(x\right)\\
f\left(x+\epsilon_{{\scriptscriptstyle x}},y+\epsilon_{{\scriptscriptstyle y}},z+\epsilon_{{\scriptscriptstyle z}}\right)= & f\left(x,y+\epsilon_{{\scriptscriptstyle y}},z+\epsilon_{{\scriptscriptstyle z}}\right)\\
+ & \epsilon\dfrac{\partial}{\partial x}f\left(x,y+\epsilon_{{\scriptscriptstyle y}},z+\epsilon_{{\scriptscriptstyle z}}\right)\\
+ & \dfrac{1}{2}\epsilon^{2}\dfrac{\partial^{2}}{\partial x^{2}}f\left(x,y+\epsilon_{{\scriptscriptstyle y}},z+\epsilon_{{\scriptscriptstyle z}}\right)\\
+ & \cdots
\end{aligned}
$$

$$
\begin{aligned}
f\left(x+\epsilon\right)= & \mathrm{e}^{{\scriptscriptstyle \epsilon\frac{\mathrm{d}}{\mathrm{d}x}}}f\left(x\right)\\
f\left(x+\epsilon_{{\scriptscriptstyle x}},y+\epsilon_{{\scriptscriptstyle y}},z+\epsilon_{{\scriptscriptstyle z}}\right)=f\left(\boldsymbol{r}+\boldsymbol{\epsilon}\right)= & \mathrm{e}^{{\scriptscriptstyle \boldsymbol{\epsilon}\cdot\boldsymbol{\nabla}}}f\left(\boldsymbol{r}\right),\begin{cases}
\boldsymbol{r}=\left(x,y,z\right)\\
\boldsymbol{\epsilon}=\left(\epsilon_{{\scriptscriptstyle x}},\epsilon_{{\scriptscriptstyle y}},\epsilon_{{\scriptscriptstyle z}}\right)\\
\boldsymbol{\nabla}=\left(\dfrac{\partial}{\partial x},\dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}\right)
\end{cases}
\end{aligned}
$$

$$
\boldsymbol{\epsilon}\cdot\boldsymbol{\nabla}=\left(\epsilon_{{\scriptscriptstyle x}},\epsilon_{{\scriptscriptstyle y}},\epsilon_{{\scriptscriptstyle z}}\right)\cdot\left(\dfrac{\partial}{\partial x},\dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}\right)=\epsilon_{{\scriptscriptstyle x}}\dfrac{\partial}{\partial x}+\epsilon_{{\scriptscriptstyle y}}\dfrac{\partial}{\partial y}+\epsilon_{{\scriptscriptstyle z}}\dfrac{\partial}{\partial z}
$$

***

multivariable Taylor operator

$$
f\left(\boldsymbol{r}+\boldsymbol{\epsilon}\right)=\mathrm{e}^{{\scriptscriptstyle \boldsymbol{\epsilon}\cdot\boldsymbol{\nabla}}}f\left(\boldsymbol{r}\right)
$$

***

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=17m36s

physics

#### making complicated equastion simple

linearize

simple pendulum

sin(x) to x linearization

potential energy

$$
f\left(x_{{\scriptscriptstyle 0}}+\epsilon\right)=\sum\limits _{k=0}^{n}\frac{f^{\left(k\right)}\left(x_{{\scriptscriptstyle 0}}\right)}{k!}\epsilon^{k}+\mathrm{O}\left(\epsilon^{n+1}\right)
$$

$$
U\left(x_{{\scriptscriptstyle 0}}+\epsilon\right)=\sum\limits _{k=0}^{n}\frac{U^{\left(k\right)}\left(x_{{\scriptscriptstyle 0}}\right)}{k!}\epsilon^{k}+\mathrm{O}\left(\epsilon^{n+1}\right)=U\left(x_{{\scriptscriptstyle 0}}\right)+U^{\prime}\left(x_{{\scriptscriptstyle 0}}\right)x+\dfrac{1}{2}U^{\prime\prime}\left(x_{{\scriptscriptstyle 0}}\right)x^{2}+\cdots
$$

$$
F=\dfrac{-\mathrm{d}U}{\mathrm{d}x}=-U^{\prime}\left(x_{{\scriptscriptstyle 0}}\right)-U^{\prime\prime}\left(x_{{\scriptscriptstyle 0}}\right)x-\cdots\overset{\text{if }U^{\prime}\left(x_{{\scriptscriptstyle 0}}\right)=0}{=}-U^{\prime\prime}\left(x_{{\scriptscriptstyle 0}}\right)x=-kx,k=U^{\prime\prime}\left(x_{{\scriptscriptstyle 0}}\right)
$$

#### non-relativistic limit

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=22m32s

$$
\begin{aligned}
E= & \sqrt{\left(mc^{2}\right)^{2}+\left(pc\right)^{2}}=\sqrt{m^{2}c^{4}+p^{2}c^{2}}\\
= & mc^{2}\sqrt{1+\left(\dfrac{pc}{mc^{2}}\right)^{2}}=mc^{2}\sqrt{1+\left(\dfrac{p}{mc}\right)^{2}}=mc^{2}\left(1+\left(\dfrac{p}{mc}\right)^{2}\right)^{\frac{1}{2}}\\
= & mc^{2}\left[\dfrac{1}{0!}+\dfrac{1}{1!}\left(\dfrac{1}{2}\right)\left(\left(\dfrac{p}{mc}\right)^{2}\right)^{1}+\dfrac{1}{2!}\left(\dfrac{1}{2}\right)\left(\dfrac{1}{2}-1\right)\left(\left(\dfrac{p}{mc}\right)^{2}\right)^{2}+\cdots\right]\\
= & mc^{2}\left[1+\dfrac{1}{2}\left(\dfrac{p}{mc}\right)^{2}-\dfrac{1}{8}\left(\dfrac{p}{mc}\right)^{4}+\cdots\right]=mc^{2}\left[1+\dfrac{1}{2}\dfrac{p^{2}}{m^{2}c^{2}}-\dfrac{1}{8}\dfrac{p^{4}}{m^{4}c^{4}}+\cdots\right]\\
= & mc^{2}+\dfrac{1}{2}\dfrac{p^{2}}{m}-\dfrac{1}{8}\dfrac{p^{4}}{m^{3}c^{2}}+\cdots=mc^{2}+\dfrac{p^{2}}{2m}-\dfrac{p^{4}}{8m^{3}c^{2}}+\cdots\\
= & E_{{\scriptscriptstyle 0}}+\dfrac{p^{2}}{2m}-\dfrac{p^{4}}{8m^{3}c^{2}}+\cdots,\begin{cases}
E_{{\scriptscriptstyle 0}}=mc^{2}\\
\dfrac{p^{2}}{2m}=\dfrac{\left(mv\right)^{2}}{2m}=\dfrac{mv^{2}}{2}=\dfrac{1}{2}mv^{2}
\end{cases}
\end{aligned}
$$

sluppy here $p=\gamma mv$

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=24m50s

leading relativistic correction

$$
-\dfrac{p^{4}}{8m^{3}c^{2}}
$$

binding energy of hydrogen atom

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=25m40s

fine-structure constant

#### quantum momentum operator

https://www.youtube.com/watch?v=HQsZG8Yxb7w&t=29m13s

## Dirac delta function

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1qu411578c&autoplay=0"></iframe>
:::

Dirac function = Dirac delta function

https://tikz.net/delta_function/

https://tikz.dev/tikz-arrows

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '90%', fig.cap='Dirac function = Dirac delta function', engine.opts=list(extra.preamble=c("\\usepackage{pgfplots}"))}
\usetikzlibrary {arrows.meta}
\begin{tikzpicture}[
  declare function={a(\x)=0;},
]
	\begin{axis}[
		axis x line = center,
		axis y line = center,
		xlabel = {$ x $}, xlabel style = {right},
		ylabel = {$ y=f\left(x\right) $}, ylabel style = {above},
		xmin = -1.1, xmax = 1.1,
		ymin = -0.1, ymax = 1.1,
		hide obscured x ticks=false, % for origin x tick label i.e. xtick = {0}
		%xtick= \empty,
		xticklabels= \empty,
		extra x ticks={0},
		%ytick = \empty,
		yticklabels= \empty,
		x = 5cm, y = 5cm, % x y scaling
		grid = major,
		domain = -1.2:1.2,
		%samples = 1000		
	]
  \draw [blue, thick, dashed, -{Latex}] (axis cs: 0,0)--(axis cs: 0,1) node [blue, below right] {$\delta\left(x\right)$};
  \addplot [blue, thick] {a(x)};
  \draw [blue, thick, fill=white] (axis cs: 0,0) circle (1.5);
	\end{axis}
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
:::

::: {.col data-latex="{0.49\textwidth}"}
$$
\delta\left(x\right)=\begin{cases}
0 & x\ne0\\
\infty & x=0
\end{cases},\int_{-\infty}^{\infty}\delta\left(x\right)\mathrm{d}x=1
$$
:::
::::::

$$
\mathrm{supp}\left(f\right)=\left\{ x\middle|\begin{cases}
x\in\mathcal{D}\\
f\left(x\right)\ne0
\end{cases}\right\} 
$$

$$
\mathrm{supp}\left(\delta\right)=\left\{ x\middle|\begin{cases}
x\in\mathcal{D}=\mathbb{R}\\
\delta\left(x\right)\ne0
\end{cases}\right\} =\left\{ 0\right\} 
$$

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '70%', fig.cap='$f(x)=1 \\text{ if } x=0 \\text{ else } 0$', engine.opts=list(extra.preamble=c("\\usepackage{pgfplots}"))}
\usetikzlibrary {arrows.meta}
\begin{tikzpicture}[
  declare function={a(\x)=0;},
]
	\begin{axis}[
		axis x line = center,
		axis y line = center,
		xlabel = {$ x $}, xlabel style = {right},
		ylabel = {$ y=f\left(x\right) $}, ylabel style = {above},
		xmin = -1.1, xmax = 1.1,
		ymin = -0.1, ymax = 1.1,
		hide obscured x ticks=false, % for origin x tick label i.e. xtick = {0}
		%xtick= \empty,
		xticklabels= \empty,
		extra x ticks={0},
		%ytick = \empty,
		yticklabels= \empty,
		x = 5cm, y = 5cm, % x y scaling
		grid = major,
		domain = -1.2:1.2,
		%samples = 1000		
	]
  \draw [blue, thick, fill=blue] (axis cs: 0,1) circle (1.5) node [blue, below right] {$f\left(x\right)$};
  \addplot [blue, thick] {a(x)};
  \draw [blue, thick, dashed] (axis cs: 0,0)--(axis cs: 0,1);
  \draw [blue, thick, fill=white] (axis cs: 0,0) circle (1.5);
	\end{axis}
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
:::

::: {.col data-latex="{0.49\textwidth}"}
$$
f\left(x\right)=\begin{cases}
0 & x\ne0\\
1 & x=0
\end{cases}
$$
:::
::::::

https://tex.stackexchange.com/questions/45275/tikz-get-values-for-predefined-dash-patterns

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '105%', fig.cap='$\\delta_{{\\scriptscriptstyle n}}\\left(x\\right)$', engine.opts=list(extra.preamble=c("\\usepackage{pgfplots}"))}
\usetikzlibrary {arrows.meta}
\begin{tikzpicture}[
  declare function={a(\x)=0;},
]
	\begin{axis}[
		axis x line = center,
		axis y line = center,
		xlabel = {$ x $}, xlabel style = {right},
		ylabel = {$ y=f\left(x\right) $}, ylabel style = {above},
		xmin = -1.1, xmax = 1.1,
		ymin = -0.1, ymax = 2.1,
		hide obscured x ticks=false, % for origin x tick label i.e. xtick = {0}
		%xtick= \empty,
		xticklabels= \empty,
		extra x ticks={0},
		%ytick = \empty,
		yticklabels= \empty,
		x = 5cm, y = 5cm, % x y scaling
		grid = major,
		domain = -1.2:1.2,
		%samples = 1000		
	]
  \addplot [blue, thick] {a(x)};
  \draw [blue, thick, fill=white] (axis cs: -0.25,0) circle (1.5);
  \draw [white, thick, fill=white] (axis cs: -0.25,0)--(axis cs: 0.25,0);
  \draw [blue, thick, fill=white] (axis cs: 0.25,0) circle (1.5);
  
  \draw [blue, thick, dashed] (axis cs: -0.25,0)--(axis cs: -0.25,2);
  \draw [blue, thick, dashed] (axis cs: 0.25,0)--(axis cs: 0.25,2);
  
  \draw [blue, thick, fill=blue] (axis cs: -0.25,2) circle (1.5);
  \draw [blue, thick, fill=blue] (axis cs: -0.25,2)--(axis cs: 0.25,2);
  \draw [blue, thick, fill=blue] (axis cs: 0.25,2) circle (1.5) node [blue, below right] {$\delta_{{\scriptscriptstyle n}}\left(x\right)$};
	\end{axis}
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
:::

::: {.col data-latex="{0.49\textwidth}"}
$$
\delta_{{\scriptscriptstyle n}}\left(x\right)=\begin{cases}
0 & \left|x\right|>\dfrac{1}{2n}\\
n & \left|x\right|\le\dfrac{1}{2n}
\end{cases},\forall n\in\mathbb{N}
$$

$$
\delta\left(x\right)=\left\{ \delta_{{\scriptscriptstyle n}}\left(x\right)\middle|n\in\mathbb{N}\right\} =\left\{ \delta_{{\scriptscriptstyle 1}}\left(x\right),\delta_{{\scriptscriptstyle 2}}\left(x\right),\cdots\right\} =\lim\limits _{n\rightarrow\infty}\delta_{{\scriptscriptstyle n}}\left(x\right)
$$
:::
::::::

$$
\delta_{{\scriptscriptstyle n}}\left(x\right)=\begin{cases}
0 & \left|x\right|>\dfrac{1}{2n}\Leftrightarrow\begin{cases}
x>\dfrac{1}{2n}\\
x<\dfrac{-1}{2n}
\end{cases}\\
n & \left|x\right|\le\dfrac{1}{2n}\Leftrightarrow\dfrac{-1}{2n}\le x\le\dfrac{1}{2n}\Leftrightarrow x\in\left[\dfrac{-1}{2n},\dfrac{1}{2n}\right]
\end{cases},\forall n\in\mathbb{N}
$$

$$
\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x=\int_{-\frac{1}{2n}}^{\frac{1}{2n}}n\mathrm{d}x=n\int_{-\frac{1}{2n}}^{\frac{1}{2n}}\mathrm{d}x=n\left[x\right]_{\frac{-1}{2n}}^{\frac{1}{2n}}=n\left[\frac{1}{2n}-\frac{-1}{2n}\right]=n\cdot\dfrac{1}{n}=1
$$

$$
\begin{aligned}
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x= & \lim\limits _{n\rightarrow\infty}\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle n}}\left(x\right)f\left(x\right)\mathrm{d}x\\
= & \lim\limits _{n\rightarrow\infty}\int_{-\frac{1}{2n}}^{\frac{1}{2n}}\delta_{{\scriptscriptstyle n}}\left(x\right)f\left(x\right)\mathrm{d}x\\
= & \lim\limits _{n\rightarrow\infty}\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x,\begin{cases}
f\left(x\right)\in\left[m,M\right]\subseteq f\left(\left[\dfrac{-1}{2n},\dfrac{1}{2n}\right]\right)\\
\Downarrow\\
m\le f\left(x\right)\le M
\end{cases}\\
\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nm\mathrm{d}x\le & \int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x\le\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nM\mathrm{d}x\\
m=m\cdot1=m\int_{-\frac{1}{2n}}^{\frac{1}{2n}}n\mathrm{d}x\le & \int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x\le M\int_{-\frac{1}{2n}}^{\frac{1}{2n}}n\mathrm{d}x=M\cdot1=M\\
\Downarrow & \ref{thm:IVT}\\
\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x\overset{\exists\xi_{{\scriptscriptstyle n}}\in\left(\frac{-1}{2n},\frac{1}{2n}\right)}{=} & f\left(\xi_{{\scriptscriptstyle n}}\right)\int_{-\frac{1}{2n}}^{\frac{1}{2n}}n\mathrm{d}x=f\left(\xi_{{\scriptscriptstyle n}}\right)\cdot1=f\left(\xi_{{\scriptscriptstyle n}}\right)\\
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x=\lim\limits _{n\rightarrow\infty}\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x= & \lim\limits _{n\rightarrow\infty}f\left(\xi_{{\scriptscriptstyle n}}\right),\xi_{{\scriptscriptstyle n}}\in\left(\frac{-1}{2n},\frac{1}{2n}\right)\\
= & f\left(0\right)\\
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x= & f\left(0\right)
\end{aligned}
$$

$$
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '95%', fig.cap='$\\int_{-\\infty}^{\\infty}\\delta\\left(x\\right)f\\left(x\\right)\\mathrm{d}x=f\\left(0\\right)$', engine.opts=list(extra.preamble=c("\\usepackage{pgfplots}"))}
\usetikzlibrary {arrows.meta}
\begin{tikzpicture}[
  declare function={a(\x)=0;},
  declare function={f(\x)=1/(1+exp(-\x));}
]
	\begin{axis}[
		axis x line = center,
		axis y line = center,
		xlabel = {$ x $}, xlabel style = {right},
		ylabel = {$ y $}, ylabel style = {above},
		xmin = -1.1, xmax = 1.1,
		ymin = -0.1, ymax = 2.1,
		hide obscured x ticks=false, % for origin x tick label i.e. xtick = {0}
		%xtick= \empty,
		xticklabels= \empty,
		extra x ticks={0},
		%ytick = \empty,
		yticklabels= \empty,
		x = 5cm, y = 5cm, % x y scaling
		grid = major,
		domain = -1.2:1.2,
		%samples = 1000		
	]
  \addplot [blue, thick] {a(x)};
  \draw [blue, thick, fill=white] (axis cs: -0.25,0) circle (1.5);
  \draw [white, thick, fill=white] (axis cs: -0.25,0)--(axis cs: 0.25,0);
  \draw [blue, thick, fill=white] (axis cs: 0.25,0) circle (1.5);
  
  \draw [blue, thick, dashed] (axis cs: -0.25,0)--(axis cs: -0.25,2);
  \draw [blue, thick, dashed] (axis cs: 0.25,0)--(axis cs: 0.25,2);
  
  \draw [blue, thick, fill=blue] (axis cs: -0.25,2) circle (1.5);
  \draw [blue, thick, fill=blue] (axis cs: -0.25,2)--(axis cs: 0.25,2);
  \draw [blue, thick, fill=blue] (axis cs: 0.25,2) circle (1.5) node [blue, below right] {$\delta_{{\scriptscriptstyle n}}\left(x\right)$};
  
  \addplot [red, thick] {f(x)};
  \node [text=red] at (axis cs: 0.5,0.5) {$f\left(x\right)=\dfrac{1}{1+\mathrm{e}^{-x}}$};
  \end{axis}
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
:::

::: {.col data-latex="{0.49\textwidth}"}
$$
\begin{aligned}
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x= & \lim\limits _{n\rightarrow\infty}\int_{-\frac{1}{2n}}^{\frac{1}{2n}}\delta_{{\scriptscriptstyle n}}\left(x\right)f\left(x\right)\mathrm{d}x\\
= & \lim\limits _{n\rightarrow\infty}\int_{-\frac{1}{2n}}^{\frac{1}{2n}}nf\left(x\right)\mathrm{d}x\\
= & \lim\limits _{n\rightarrow\infty}f\left(\xi_{{\scriptscriptstyle n}}\right),\xi_{{\scriptscriptstyle n}}\in\left(\frac{-1}{2n},\frac{1}{2n}\right)
\end{aligned}
$$

$$
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

$$
\int_{-\infty}^{\infty}\delta\left(x-0\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$
:::
::::::

***

$$
\int_{-\infty}^{\infty}\delta\left(x\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

$$
\int_{-\infty}^{\infty}\delta\left(x-0\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

$$
\int_{-\infty}^{\infty}\delta\left(x-x_{{\scriptscriptstyle 0}}\right)f\left(x\right)\mathrm{d}x=f\left(x_{{\scriptscriptstyle 0}}\right)
$$

$$
\int_{-\infty}^{\infty}\delta\left(x-x^{\prime}\right)f\left(x\right)\mathrm{d}x=f\left(x^{\prime}\right)
$$

***

https://www.youtube.com/watch?v=nDa3cqFk80o

$$
\begin{aligned}
 & \left\{ \left\{ f_{{\scriptscriptstyle n}}\left(x\right)\middle|n\in\mathbb{N}\right\} \middle|\begin{cases}
f_{{\scriptscriptstyle n}}:\mathbb{R}\rightarrow\mathbb{R}\\
\int_{-\infty}^{\infty}f_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x=1
\end{cases}\right\} \\
=\Biggl\{ & \left\{ \dfrac{n}{2}\mathrm{e}^{-n\left|x\right|}\middle|n\in\mathbb{N}\right\} ,\\
 & \left\{ \dfrac{1}{\pi}\dfrac{n}{n^{2}x^{2}+1}\middle|n\in\mathbb{N}\right\} ,\\
 & \left\{ \dfrac{n}{\sqrt{\pi}}\mathrm{e}^{-n^{2}x^{2}}\middle|n\in\mathbb{N}\right\} ,\\
 & \left\{ \dfrac{1}{\pi}\dfrac{\sin\left(nx\right)}{x}\middle|n\in\mathbb{N}\right\} ,\cdots\Biggr\}
\end{aligned}
$$

$$
\left\{ \delta\left(x\right)\middle|\begin{cases}
\delta:\mathbb{R}\rightarrow\mathbb{R}\\
\int_{-\infty}^{\infty}\delta\left(x\right)\cdot1\mathrm{d}x=1
\end{cases}\right\} 
$$

$$
\left\{ \delta\left(x\right)\middle|\begin{cases}
\delta:\mathbb{R}\rightarrow\mathbb{R}\\
\int_{-\infty}^{\infty}\delta\left(x-0\right)\cdot f\left(x\right)\mathrm{d}x=f\left(0\right)
\end{cases}\right\} 
$$

In measure theory, we can define the distance of two functions by

$$
d\left(f,g\right)=\sqrt{\int_{-\infty}^{\infty}\left[f\left(x\right)-g\left(x\right)\right]^{2}\mathrm{d}x}
$$

for real distance of two square delta function approximations,

$$
\begin{aligned}
d\left(\delta_{{\scriptscriptstyle m}},\delta_{{\scriptscriptstyle n}}\right)= & \sqrt{\int_{-\infty}^{\infty}\left[\delta_{{\scriptscriptstyle m}}\left(x\right)-\delta_{{\scriptscriptstyle n}}\left(x\right)\right]^{2}\mathrm{d}x}\\
= & \sqrt{\int_{-\infty}^{\infty}\left[\delta_{{\scriptscriptstyle m}}\left(x\right)\right]^{2}-2\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)+\left[\delta_{{\scriptscriptstyle n}}\left(x\right)\right]^{2}\mathrm{d}x}\\
= & \sqrt{\int_{-\infty}^{\infty}\left[\delta_{{\scriptscriptstyle m}}\left(x\right)\right]^{2}\mathrm{d}x-2\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x+\int_{-\infty}^{\infty}\left[\delta_{{\scriptscriptstyle n}}\left(x\right)\right]^{2}\mathrm{d}x}\\
= & \sqrt{\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle m}}\left(x\right)\mathrm{d}x-2\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x+\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle n}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x}\\
= & \sqrt{\delta_{{\scriptscriptstyle m}}\left(\xi_{{\scriptscriptstyle m}}\right)-2\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x+\delta_{{\scriptscriptstyle n}}\left(\xi_{{\scriptscriptstyle n}}\right)},\begin{cases}
\xi_{{\scriptscriptstyle m}}\in\left(\frac{-1}{2m},\frac{1}{2m}\right)\\
\xi_{{\scriptscriptstyle n}}\in\left(\frac{-1}{2n},\frac{1}{2n}\right)
\end{cases}\\
= & \sqrt{m-2\int_{-\infty}^{\infty}\delta_{{\scriptscriptstyle m}}\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x+n},\text{ if }m>n\\
= & \sqrt{m-2\delta_{{\scriptscriptstyle n}}\left(\xi_{{\scriptscriptstyle n}}\right)+n},\xi_{{\scriptscriptstyle n}}\in\left(\frac{-1}{2n},\frac{1}{2n}\right)\\
= & \sqrt{m-2n+n}=\sqrt{m-n}\in\mathbb{R}\\
d\left(\delta_{{\scriptscriptstyle m}},\delta_{{\scriptscriptstyle n}}\right)\overset{m>n}{=} & \sqrt{m-n}\in\mathbb{R}
\end{aligned}
$$

$$
\left\langle d\left(\delta_{{\scriptscriptstyle m}},\delta_{{\scriptscriptstyle n}}\right)\right\rangle _{{\scriptscriptstyle n\in\mathbb{N}}}=\left\langle \sqrt{m-n}\right\rangle _{{\scriptscriptstyle n\in\mathbb{N}}}\text{ is not a Cauchy series, not even mentioned convergence}
$$

### complex delta function

$$
\delta\left(k\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x
$$

$$
\begin{aligned}
 & \int_{-\infty}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\\
= & \lim\limits _{k^{\prime}\rightarrow0}\left[\int_{-\infty}^{k^{\prime-}}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k+\int_{k^{\prime-}}^{k^{\prime+}}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k+\int_{k^{\prime+}}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[\int_{-\infty}^{0}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k+\int_{k^{\prime-}}^{k^{\prime+}}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}0x}\mathrm{d}x\right)f\left(0,x\right)\mathrm{d}k+\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[\int_{-\infty}^{0}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k+\int_{k^{\prime-}}^{k^{\prime+}}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}1\mathrm{d}x\right)f\left(0,x\right)\mathrm{d}k+\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[-\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k+\int_{k^{\prime-}}^{k^{\prime+}}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}1\mathrm{d}x\right)f\left(0,x\right)\mathrm{d}k+\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[-\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\mathrm{d}x\right)f\left(x\right)\mathrm{d}k+\int_{k^{\prime-}}^{k^{\prime+}}\left(\frac{1}{2\pi}\infty\right)f\left(0,x\right)\mathrm{d}k+\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[\int_{k^{\prime-}}^{k^{\prime+}}\dfrac{1}{\mathrm{d}k}f\left(0,x\right)\mathrm{d}k+\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}-\mathrm{e}^{-\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\right]\\
\approx & \lim\limits _{k^{\prime}\rightarrow0}\left[f\left(0,x\right)+\int_{0}^{\infty}\left(\frac{1}{2\pi}0\right)f\left(k,x\right)\mathrm{d}k\right]\approx f\left(0,x\right)\\
 & f\left(0,x\right)=\int_{-\infty}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k=\int_{-\infty}^{\infty}\delta\left(k\right)f\left(k,x\right)\mathrm{d}k=\int_{-\infty}^{\infty}\delta\left(k-0\right)f\left(k,x\right)\mathrm{d}k
\end{aligned}
$$

https://tex.stackexchange.com/questions/150138/how-can-i-create-a-polar-plot-on-a-cartesian-grid

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
```{r, echo=FALSE, cache=TRUE, engine='tikz', fig.ext=if (knitr:::is_latex_output()) 'pdf' else 'png', out.width=if (knitr:::is_html_output()) '95%', fig.cap='$\\int_{-\\infty}^{\\infty}\\left(\\frac{1}{2\\pi}\\int_{-\\infty}^{\\infty}\\mathrm{e}^{\\mathrm{i}kx}\\mathrm{d}x\\right)f\\left(k,x\\right)\\mathrm{d}k$', engine.opts=list(extra.preamble=c("\\usepackage{pgfplots}"))}
\begin{tikzpicture}
	\begin{axis}[
		axis x line = center,
		axis y line = center,
		xlabel = {$ a $}, xlabel style = {right},
		ylabel = {$ b\mathrm{i} $}, ylabel style = {above},
		xmin = -1.1, xmax = 1.1,
		ymin = -1.1, ymax = 1.1,
		hide obscured x ticks=false, % for origin x tick label i.e. xtick = {0}
		%xtick= \empty,
		xticklabels= \empty,
		extra x ticks={0},
		%ytick = \empty,
		yticklabels= \empty,
		x = 5cm, y = 5cm, % x y scaling
		grid = major,
		domain = -1.2:1.2,
		%samples = 1000		
	]
  \draw [black, thick, fill=black] (axis cs: 0,0) circle (1.5);
  \node [text=black] at (axis cs: 0.0,0.2) {$f\left(0,x\right)$};
  \addplot[data cs=polar, blue, domain=0:360*4, samples=360*4, smooth] (x,{0.001*x});
  \node [text=blue] at (axis cs: -0.5,0.5) {$\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k$};
  \addplot[data cs=polar, red, domain=0:360*4, samples=360*4, smooth] (x,{-0.001*x});
  \node [text=red] at (axis cs: 0.5,-0.5) {$-\int_{0}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k$};
  \end{axis}
\end{tikzpicture}
```
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
:::

::: {.col data-latex="{0.49\textwidth}"}
$$
\begin{aligned}
f\left(0,x\right)= & \int_{-\infty}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{d}x\right)f\left(k,x\right)\mathrm{d}k\\
= & \int_{-\infty}^{\infty}\delta\left(k-0\right)f\left(k,x\right)\mathrm{d}k
\end{aligned}
$$
:::
::::::

***

According to Fourier transform and inverse transform, [Fourier analysis]^[\@ref(fourier-analysis)]^

$$
f\left(t\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}t\cdot\omega}\left(\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}\omega\cdot s}f\left(s\right)\mathrm{d}s\right)\mathrm{d}\omega
$$

$$
f\left(x\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}xk}\left(\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx^{\prime}}f\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k
$$

or

$$
f\left(t\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}t\cdot\omega}\left(\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}\imath\omega\cdot s}f\left(s\right)\mathrm{d}s\right)\mathrm{d}\omega
$$

$$
f\left(t\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}xk}\left(\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx^{\prime}}f\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k
$$

***

complex function if analytic always with better properties than real function

$$
\begin{aligned}
 & \left\{ \left\{ \delta\left(k\right)\middle|k\in\mathbb{R}\right\} \middle|\begin{cases}
\delta:\mathbb{R}\rightarrow\mathbb{C}\\
\int_{-\infty}^{\infty}\delta\left(k-0\right)f\left(k,x\right)\mathrm{d}k=f\left(0,x\right)
\end{cases}\right\} \\
= & \left\{ \left\{ K\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\left(\cdot\right)\mathrm{d}k\middle|k\in\mathbb{R}\right\} ,\cdots\right\} 
\end{aligned}
$$

***

$$
\delta\left(k\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\left(\cdot\right)\mathrm{d}x
$$

$$
\delta\left(x\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\left(\cdot\right)\mathrm{d}k
$$

$$
\delta\left(x^{\prime}\right)=\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx^{\prime}}\left(\cdot\right)\mathrm{d}k
$$

***

$\psi\left(x\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi\left(k\right)\mathrm{d}k$

$$
\psi\left(k\right)=\dfrac{1}{\sqrt{2\pi}}\int\mathrm{e}^{-\mathrm{i}kx}\psi\left(x\right)\mathrm{d}x
$$

$$
\begin{aligned}
 & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi\left(k\right)\mathrm{d}k\\
= & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\left(\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx^{\prime}}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k\\
= & \frac{1}{2\pi}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{e}^{-\mathrm{i}kx^{\prime}}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k\\
\overset{\text{Fubini}}{=} & \frac{1}{2\pi}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\psi\left(x^{\prime}\right)\mathrm{d}k\right)\mathrm{d}x^{\prime}\\
= & \int_{-\infty}^{\infty}\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\psi\left(x^{\prime}\right)\mathrm{d}k\right)\mathrm{d}x^{\prime}\\
= & \int_{-\infty}^{\infty}\psi\left(x^{\prime}\right)\left(\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\mathrm{d}k\right)\mathrm{d}x^{\prime}\\
= & \int_{-\infty}^{\infty}\psi\left(x^{\prime}\right)\delta\left(x-x^{\prime}\right)\mathrm{d}x^{\prime}=\int_{-\infty}^{\infty}\delta\left(x-x^{\prime}\right)\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}=\psi\left(x\right)
\end{aligned}
$$

$$
\psi\left(x\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi\left(k\right)\mathrm{d}k
$$

***

$\int_{-\infty}^{\infty}\left|\psi\left(k\right)\right|^{2}\mathrm{d}k=\int_{-\infty}^{\infty}\left|\psi\left(x\right)\right|^{2}\mathrm{d}x$

$$
\begin{cases}
\psi\left(x\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi\left(k\right)\mathrm{d}k\\
\psi\left(k\right)=\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\psi\left(x\right)\mathrm{d}x
\end{cases}
$$

$$
\begin{aligned}
\psi^{*}\left(k\right)=\overline{\psi}\left(k\right)= & \overline{\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\psi\left(x\right)\mathrm{d}x}\\
= & \dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\overline{\mathrm{e}^{-\mathrm{i}kx}\psi\left(x\right)}\mathrm{d}x\\
= & \dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\overline{\mathrm{e}^{-\mathrm{i}kx}}\overline{\psi\left(x\right)}\mathrm{d}x\\
= & \dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi^{*}\left(x\right)\mathrm{d}x
\end{aligned}
$$

$$
\begin{aligned}
 & \int_{-\infty}^{\infty}\left|\psi\left(k\right)\right|^{2}\mathrm{d}k=\int_{-\infty}^{\infty}\psi^{*}\left(k\right)\psi\left(k\right)\mathrm{d}k\\
= & \int_{-\infty}^{\infty}\left(\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi^{*}\left(x\right)\mathrm{d}x\right)\left(\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx^{\prime}}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k\\
= & \frac{1}{2\pi}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\psi^{*}\left(x\right)\mathrm{d}x\right)\left(\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx^{\prime}}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\right)\mathrm{d}k\\
= & \frac{1}{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\mathrm{e}^{-\mathrm{i}kx^{\prime}}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\mathrm{d}x\mathrm{d}k\\
= & \frac{1}{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\mathrm{d}x\mathrm{d}k\\
= & \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\psi\left(x^{\prime}\right)\mathrm{d}x^{\prime}\mathrm{d}x\mathrm{d}k\\
\overset{\text{Fubini}}{=} & \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\psi\left(x^{\prime}\right)\mathrm{d}k\mathrm{d}x^{\prime}\mathrm{d}x\\
= & \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\psi\left(x^{\prime}\right)\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k\left(x-x^{\prime}\right)}\mathrm{d}k\mathrm{d}x^{\prime}\mathrm{d}x\\
= & \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\psi^{*}\left(x\right)\psi\left(x^{\prime}\right)\delta\left(x-x^{\prime}\right)\mathrm{d}x^{\prime}\mathrm{d}x\\
= & \int_{-\infty}^{\infty}\psi^{*}\left(x\right)\int_{-\infty}^{\infty}\psi\left(x^{\prime}\right)\delta\left(x-x^{\prime}\right)\mathrm{d}x^{\prime}\mathrm{d}x\\
= & \int_{-\infty}^{\infty}\psi^{*}\left(x\right)\psi\left(x\right)\mathrm{d}x=\int_{-\infty}^{\infty}\left|\psi\left(x\right)\right|^{2}\mathrm{d}x
\end{aligned}
$$

$$
\int_{-\infty}^{\infty}\left|\psi\left(k\right)\right|^{2}\mathrm{d}k=\int_{-\infty}^{\infty}\left|\psi\left(x\right)\right|^{2}\mathrm{d}x
$$

***

convolution

$\psi\left(x\right)=\int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(y\right)\psi_{{\scriptscriptstyle 2}}\left(x-y\right)\mathrm{d}y\Rightarrow\psi\left(k\right)=\sqrt{2\pi}\psi_{{\scriptscriptstyle 1}}\left(k\right)\psi_{{\scriptscriptstyle 2}}\left(k\right)$

$$
\psi\left(x\right)=\int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(y\right)\psi_{{\scriptscriptstyle 2}}\left(x-y\right)\mathrm{d}y
$$

$$
\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 1}}y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}
$$

$$
\psi_{{\scriptscriptstyle 2}}\left(x-y\right)=\frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 2}}\left(x-y\right)}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 2}}
$$

$$
\begin{cases}
\psi\left(x\right)= & \int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(y\right)\psi_{{\scriptscriptstyle 2}}\left(x-y\right)\mathrm{d}y\\
\psi_{{\scriptscriptstyle 1}}\left(y\right)= & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 1}}y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\\
\psi_{{\scriptscriptstyle 2}}\left(x-y\right)= & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 2}}\left(x-y\right)}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 2}}
\end{cases}
$$

$$
\begin{aligned}
 & \psi\left(k\right)=\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\psi\left(x\right)\mathrm{d}x=\dfrac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(y\right)\psi_{{\scriptscriptstyle 2}}\left(x-y\right)\mathrm{d}y\mathrm{d}x\\
= & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\mathrm{e}^{-\mathrm{i}kx}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 1}}y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\right)\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 2}}\left(x-y\right)}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 2}}\right)\mathrm{d}y\mathrm{d}x\\
= & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 1}}y}\mathrm{e}^{-\mathrm{i}k_{{\scriptscriptstyle 2}}y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{e}^{-\mathrm{i}kx}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 2}}x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\mathrm{d}y\mathrm{d}x\\
= & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\mathrm{d}y\mathrm{d}x\\
\overset{\text{Fubini}}{=} & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}y\mathrm{d}x\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\\
= & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}y\right)\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}x\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}
\end{aligned}
$$

$$
\begin{aligned}
 & \dfrac{1}{\left(\sqrt{2\pi}\right)^{3}}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}y\right)\left(\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}x\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\\
= & \sqrt{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left[\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}y\right]\left[\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}x\right]\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\\
= & \sqrt{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left[\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)y}\mathrm{d}y\right]\left[\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\frac{1}{2\pi}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\left(k_{{\scriptscriptstyle 2}}-k\right)x}\mathrm{d}x\right]\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\\
= & \sqrt{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left[\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\delta\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)\right]\left[\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\delta\left(k_{{\scriptscriptstyle 2}}-k\right)\right]\mathrm{d}k_{{\scriptscriptstyle 1}}\mathrm{d}k_{{\scriptscriptstyle 2}}\\
\overset{\text{Fubini}}{=} & \sqrt{2\pi}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\left[\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\delta\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)\right]\left[\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\delta\left(k_{{\scriptscriptstyle 2}}-k\right)\right]\mathrm{d}k_{{\scriptscriptstyle 2}}\mathrm{d}k_{{\scriptscriptstyle 1}}\\
= & \sqrt{2\pi}\int_{-\infty}^{\infty}\left[\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\right]\left[\int_{-\infty}^{\infty}\delta\left(k_{{\scriptscriptstyle 1}}-k_{{\scriptscriptstyle 2}}\right)\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\delta\left(k_{{\scriptscriptstyle 2}}-k\right)\mathrm{d}k_{{\scriptscriptstyle 2}}\right]\mathrm{d}k_{{\scriptscriptstyle 1}}\\
= & \sqrt{2\pi}\int_{-\infty}^{\infty}\left[\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\right]\left[\delta\left(k_{{\scriptscriptstyle 1}}-k\right)\psi_{{\scriptscriptstyle 2}}\left(k\right)\right]\mathrm{d}k_{{\scriptscriptstyle 1}}=\sqrt{2\pi}\psi_{{\scriptscriptstyle 2}}\left(k\right)\int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\delta\left(k_{{\scriptscriptstyle 1}}-k\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\\
= & \sqrt{2\pi}\psi_{{\scriptscriptstyle 2}}\left(k\right)\psi_{{\scriptscriptstyle 1}}\left(k\right)=\sqrt{2\pi}\psi_{{\scriptscriptstyle 1}}\left(k\right)\psi_{{\scriptscriptstyle 2}}\left(k\right)
\end{aligned}
$$

$$
\begin{array}{c}
\begin{cases}
\psi\left(x\right)= & \int_{-\infty}^{\infty}\psi_{{\scriptscriptstyle 1}}\left(y\right)\psi_{{\scriptscriptstyle 2}}\left(x-y\right)\mathrm{d}y\\
\psi_{{\scriptscriptstyle 1}}\left(y\right)= & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 1}}y}\psi_{{\scriptscriptstyle 1}}\left(k_{{\scriptscriptstyle 1}}\right)\mathrm{d}k_{{\scriptscriptstyle 1}}\\
\psi_{{\scriptscriptstyle 2}}\left(x-y\right)= & \frac{1}{\sqrt{2\pi}}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}k_{{\scriptscriptstyle 2}}\left(x-y\right)}\psi_{{\scriptscriptstyle 2}}\left(k_{{\scriptscriptstyle 2}}\right)\mathrm{d}k_{{\scriptscriptstyle 2}}
\end{cases}\\
\Downarrow\\
\psi\left(k\right)=\sqrt{2\pi}\psi_{{\scriptscriptstyle 1}}\left(k\right)\psi_{{\scriptscriptstyle 2}}\left(k\right)
\end{array}
$$

### 3D delta function

https://www.youtube.com/watch?v=Y8y965ZAmQE

$$
\delta\left(\boldsymbol{r}-\boldsymbol{r}^{\prime}\right)=\delta\left(x-x^{\prime}\right)\delta\left(y-y^{\prime}\right)\delta\left(z-z^{\prime}\right)
$$

Poisson equation

$$
\begin{aligned}
\Delta\left(\dfrac{1}{r}\right)= & \boldsymbol{\nabla}^{2}\left(\dfrac{1}{\sqrt{\boldsymbol{r}^{2}}}\right)=\boldsymbol{\nabla}\cdot\boldsymbol{\nabla}\left(\dfrac{1}{\sqrt{\boldsymbol{r}\cdot\boldsymbol{r}}}\right)=\boldsymbol{\nabla}\cdot\left(\dfrac{\partial}{\partial x},\dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}\right)\left(\dfrac{1}{\sqrt{x^{2}+y^{2}+z^{2}}}\right)\\
= & \boldsymbol{\nabla}\cdot\left(\dfrac{-x}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}},\dfrac{-y}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}},\dfrac{-z}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}}\right)=\boldsymbol{\nabla}\cdot\dfrac{-\boldsymbol{r}}{r^{3}}\\
= & \left(\dfrac{\partial}{\partial x},\dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}\right)\cdot\left(\dfrac{-x}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}},\dfrac{-y}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}},\dfrac{-z}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}}\right)\\
= & \dfrac{\partial}{\partial x}\dfrac{-x}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}}+\dfrac{\partial}{\partial y}\dfrac{-y}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}}+\dfrac{\partial}{\partial z}\dfrac{-z}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{3}{2}}}\\
= & \dfrac{+2x^{2}-y^{2}-z^{2}}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{5}{2}}}+\dfrac{-x^{2}+2y^{2}-z^{2}}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{5}{2}}}+\dfrac{-x^{2}-y^{2}+2z^{2}}{\left(x^{2}+y^{2}+z^{2}\right)^{\frac{5}{2}}}=0
\end{aligned}
$$

$\Delta\left(\dfrac{1}{r}\right)=-4\pi\delta\left(\boldsymbol{r}\right)$

https://math.stackexchange.com/questions/3774483/derivatives-of-frac1r-and-dirac-delta-function

$$
\begin{aligned}
\Delta\dfrac{1}{r\left(\epsilon\right)}= & \boldsymbol{\nabla}^{2}\dfrac{1}{\sqrt{\boldsymbol{r}^{2}+\epsilon^{2}}}=\boldsymbol{\nabla}\cdot\boldsymbol{\nabla}\dfrac{1}{\sqrt{x^{2}+y^{2}+z^{2}+\epsilon^{2}}}\\
= & \boldsymbol{\nabla}\cdot\left(\dfrac{\partial}{\partial x},\dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}\right)\left(\dfrac{1}{\sqrt{x^{2}+y^{2}+z^{2}+\epsilon^{2}}}\right)\\
= & \boldsymbol{\nabla}\cdot\left(\dfrac{-x}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}},\dfrac{-y}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}},\dfrac{-z}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}}\right)\\
= & \dfrac{\partial}{\partial x}\dfrac{-x}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}}+\dfrac{\partial}{\partial y}\dfrac{-y}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}}+\dfrac{\partial}{\partial z}\dfrac{-z}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{3}{2}}}\\
= & \dfrac{+2x^{2}-y^{2}-z^{2}-\epsilon^{2}}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}+\dfrac{-x^{2}+2y^{2}-z^{2}-\epsilon^{2}}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}+\dfrac{-x^{2}-y^{2}+2z^{2}-\epsilon^{2}}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}\\
= & \dfrac{-3\epsilon^{2}}{\left(x^{2}+y^{2}+z^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}=\dfrac{-3\epsilon^{2}}{\left(\boldsymbol{r}^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}
\end{aligned}
$$

$$
\Delta\left(\dfrac{1}{r}\right)=\boldsymbol{\nabla}^{2}\left(\dfrac{1}{r}\right)=\boldsymbol{\nabla}^{2}\dfrac{1}{\left|\boldsymbol{r}\right|}=\boldsymbol{\nabla}^{2}\dfrac{1}{\left|\boldsymbol{r}-\boldsymbol{0}\right|}=\Delta\dfrac{1}{\left|\boldsymbol{r}-\boldsymbol{0}\right|}=\Delta\left(\boldsymbol{r}-\boldsymbol{0}\right)=\Delta\left(\boldsymbol{r}\right)
$$

$$
\Delta\dfrac{1}{r\left(\epsilon\right)}=\Delta_{\epsilon}\left(r\right)=\Delta_{\epsilon}\left(\boldsymbol{r}\right)=\Delta_{\epsilon}\left(\boldsymbol{r}-\boldsymbol{0}\right)
$$

$$
\lim\limits _{\epsilon\rightarrow0}\Delta_{\epsilon}\left(\boldsymbol{r}\right)=\lim\limits _{\epsilon\rightarrow0}\Delta\dfrac{1}{r\left(\epsilon\right)}=\Delta\left(\dfrac{1}{r}\right)=\Delta\left(\boldsymbol{r}\right)
$$

$$
K_{{\scriptscriptstyle 3}}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}=\lim\limits _{\epsilon\rightarrow0}K_{{\scriptscriptstyle 3}}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}=\lim\limits _{\epsilon\rightarrow0}K_{{\scriptscriptstyle 3}}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}-\boldsymbol{0}\right)\mathrm{d}\boldsymbol{r}=f\left(\boldsymbol{0}\right)
$$

$$
K_{{\scriptscriptstyle 1}}\int_{\mathbb{R}^{1}}f\left(x\right)\delta\left(x\right)\mathrm{d}x=\lim\limits _{n\rightarrow\infty}K_{{\scriptscriptstyle 1}}\int_{\mathbb{R}^{1}}f\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x\right)\mathrm{d}x=K_{{\scriptscriptstyle 1}}\int_{\mathbb{R}^{1}}f\left(x\right)\delta_{{\scriptscriptstyle n}}\left(x-0\right)\mathrm{d}x=f\left(0\right)
$$

$$
\begin{aligned}
 & \int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}=\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\dfrac{-3\epsilon^{2}}{\left(\boldsymbol{r}^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}\mathrm{d}\boldsymbol{r},\begin{cases}
\boldsymbol{r}=r\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\\
\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}=\begin{pmatrix}\sin\phi\cos\theta\\
\sin\phi\sin\theta\\
\cos\phi
\end{pmatrix}
\end{cases}\\
= & \int_{0}^{\infty}\int_{\mathbb{S}^{2}}f\left(r\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{-3\epsilon^{2}}{\left(r^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}r^{2}\sigma\left(\mathrm{d}\boldsymbol{n}_{{\scriptscriptstyle S}}\right)\mathrm{d}r,\begin{cases}
\mathbb{S}^{2}=\text{unit sphere centered at the origin}\\
\sigma\left(\mathrm{d}\boldsymbol{n}_{{\scriptscriptstyle S}}\right)\text{ is the surface measure of }\mathbb{S}^{2}\\
r^{2}\sigma\left(\mathrm{d}\boldsymbol{n}_{{\scriptscriptstyle S}}\right)=r^{2}\sin\phi\mathrm{d}\phi\mathrm{d}\theta\\
r=\epsilon s
\end{cases}\\
= & \int_{0}^{\infty}\int_{\mathbb{S}^{2}}f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{-3\epsilon^{2}}{\left(\left(\epsilon s\right)^{2}+\epsilon^{2}\right)^{\frac{5}{2}}}\left(\epsilon s\right)^{2}\sigma\left(\mathrm{d}\boldsymbol{n}_{{\scriptscriptstyle S}}\right)\mathrm{d}\epsilon s\\
= & \int_{0}^{\infty}\int_{\mathbb{S}^{2}}f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{-3\epsilon^{5}}{\epsilon^{5}\left(s^{2}+1\right)^{\frac{5}{2}}}s^{2}\sigma\left(\mathrm{d}\boldsymbol{n}_{{\scriptscriptstyle S}}\right)\mathrm{d}s=-\int_{0}^{\infty}\int_{\mathbb{S}^{2}}f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\sigma\left(\mathrm{d}\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\mathrm{d}s\\
= & -\int_{0}^{\infty}\int_{0}^{2\pi}\int_{0}^{2\pi}f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\sin\phi\mathrm{d}\phi\mathrm{d}\theta\mathrm{d}s=-\int_{0}^{\infty}4\pi f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s
\end{aligned}
$$
$$
\begin{aligned}
 & \lim\limits _{\epsilon\rightarrow0}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}\\
= & \lim\limits _{\epsilon\rightarrow0}\left[-\int_{0}^{\infty}4\pi f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s\right]\\
= & -4\pi f\left(\boldsymbol{0}\right)\int_{0}^{\infty}\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s
\end{aligned}
$$

https://www.integral-calculator.com/

$$
\begin{aligned}
 & \int\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s,s=\tan u\\
= & \int\dfrac{3\tan^{2}u}{\left(\tan^{2}u+1\right)^{\frac{5}{2}}}\mathrm{d}\tan u=\int\dfrac{3\tan^{2}u}{\left(\tan^{2}u+1\right)^{\frac{5}{2}}}\sec^{2}u\mathrm{d}u\\
= & \int\dfrac{3\tan^{2}u}{\left(\sec^{2}u\right)^{\frac{5}{2}}}\sec^{2}u\mathrm{d}u=\int\dfrac{3\tan^{2}u}{\sec^{5}u}\sec^{2}u\mathrm{d}u=\int\dfrac{3\tan^{2}u}{\sec^{3}u}\mathrm{d}u\\
= & \int\dfrac{3\left(\frac{\sin u}{\cos u}\right)^{2}}{\left(\frac{1}{\cos u}\right)^{3}}\mathrm{d}u=3\int\sin^{2}u\cos u\mathrm{d}u=3\int\sin^{2}u\mathrm{d}\sin u=3\dfrac{\sin^{3}u}{3}+C\\
= & \sin^{3}u+C=\left(\frac{s}{\sqrt{s^{2}+1}}\right)^{3}+C=\dfrac{s^{3}}{\left(s^{2}+1\right)^{\frac{3}{2}}}+C\overset{s>0}{=}\dfrac{1}{\left(1+s^{-2}\right)^{\frac{3}{2}}}+C
\end{aligned}
$$

$$
\int_{0}^{\infty}\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s=\left[\dfrac{s^{3}}{\left(s^{2}+1\right)^{\frac{3}{2}}}\right]_{0}^{\infty}=\left[\dfrac{1}{\left(1+s^{-2}\right)^{\frac{3}{2}}}\right]_{s=\infty}-\left[\dfrac{s^{3}}{\left(s^{2}+1\right)^{\frac{3}{2}}}\right]_{s=0}=1-0=1
$$

$$
\begin{aligned}
 & \lim\limits _{\epsilon\rightarrow0}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}\\
= & \lim\limits _{\epsilon\rightarrow0}\left[-\int_{0}^{\infty}4\pi f\left(\epsilon s\boldsymbol{n}_{\mathbb{{\scriptscriptstyle S}}}\right)\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s\right]\\
= & -4\pi f\left(\boldsymbol{0}\right)\int_{0}^{\infty}\dfrac{3s^{2}}{\left(s^{2}+1\right)^{\frac{5}{2}}}\mathrm{d}s=-4\pi f\left(\boldsymbol{0}\right)1=-4\pi f\left(\boldsymbol{0}\right)
\end{aligned}
$$

$$
\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}=\lim\limits _{\epsilon\rightarrow0}\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\Delta_{\epsilon}\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}=-4\pi f\left(\boldsymbol{0}\right)=-4\pi\int_{\mathbb{R}^{3}}f\left(\boldsymbol{r}\right)\delta\left(\boldsymbol{r}\right)\mathrm{d}\boldsymbol{r}
$$

$$
\Delta\left(\dfrac{1}{r}\right)=\Delta\left(\boldsymbol{r}\right)=-4\pi\delta\left(\boldsymbol{r}\right)
$$

$$
\tag*{$\Box$}
$$

***

$$
\Delta\left(\dfrac{1}{r}\right)=-4\pi\delta\left(\boldsymbol{r}\right)
$$

***

https://math.stackexchange.com/questions/368155/where-does-the-relation-nabla21-r-4-pi-delta3-bf-r-between-laplacian

***

$$
\begin{aligned}
 & \left\{ \left\{ \delta\left(k\right)\middle|k\in\mathbb{R}\right\} \middle|\begin{cases}
\delta:\mathbb{R}\rightarrow\mathbb{C}\\
\int_{-\infty}^{\infty}\delta\left(k-0\right)f\left(k,x\right)\mathrm{d}k=f\left(0,x\right)
\end{cases}\right\} \\
= & \left\{ \left\{ K\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}kx}\left(\cdot\right)\mathrm{d}k\middle|k\in\mathbb{R}\right\} ,\cdots\right\} 
\end{aligned}
$$

[\@ref(quaternion-1)]

$$
\begin{aligned}
 & \left\{ \left\{ \delta\left(\boldsymbol{k}\right)\middle|\boldsymbol{k}\in\mathbb{R}^{3}\right\} \middle|\begin{cases}
\delta:\mathbb{R}^{3}\rightarrow\mathbb{H}\\
\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\delta\left(\boldsymbol{k}-\boldsymbol{0}\right)f\left(\boldsymbol{k},\boldsymbol{r}\right)\mathrm{d}^{3}\boldsymbol{r}=f\left(\boldsymbol{0},\boldsymbol{r}\right)
\end{cases}\right\} \\
= & \left\{ \left\{ K\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\mathrm{e}^{\mathrm{i}\boldsymbol{k}\cdot\boldsymbol{r}}\left(\cdot\right)\mathrm{d}^{3}\boldsymbol{k}\middle|\boldsymbol{k}\in\mathbb{R}^{3}\right\} ,\cdots\right\} 
\end{aligned}
$$

***

Columb law

$$
F_{\text{e}}=k\dfrac{Qq}{r^{2}}
$$
$$
\boldsymbol{F}_{\text{e}}=k\dfrac{Qq}{r^{2}}\hat{\boldsymbol{r}}
$$

electric field from point electric charge

$$
\begin{aligned}
\boldsymbol{E}\left(\boldsymbol{r}\right)=\boldsymbol{E}\left(r\right)=\boldsymbol{E}=&k\dfrac{q}{r^{2}}\hat{\boldsymbol{r}}=k_{\text{e}}\dfrac{q}{\left|\boldsymbol{r}\right|^{2}}\dfrac{\boldsymbol{r}}{\left|\boldsymbol{r}\right|}=k_{\text{e}}q\dfrac{\boldsymbol{r}}{\left|\boldsymbol{r}\right|^{3}}=k_{\text{e}}q\dfrac{\boldsymbol{r}}{\left(\boldsymbol{r}^{2}\right)^{\frac{3}{2}}}\\=&k_{\text{e}}q\dfrac{\boldsymbol{r}}{\left(\boldsymbol{r}^{2}\right)^{\frac{3}{2}}}=k_{\text{e}}q\left(-\boldsymbol{\nabla}\frac{1}{\sqrt{\boldsymbol{r}\cdot\boldsymbol{r}}}\right)=-k_{\text{e}}q\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }
\end{aligned}
$$

$$
\boldsymbol{E}\left(\boldsymbol{r}\right)=-k_{\text{e}}q\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }
$$

electric field from tiny point electric charge

$$
\begin{aligned}
\mathrm{d}\boldsymbol{E}\left(\boldsymbol{r}\right)=\mathrm{d}\boldsymbol{E}\left(r\right)=\mathrm{d}\boldsymbol{E}= & k\dfrac{\mathrm{d}q}{r^{2}}\hat{\boldsymbol{r}}=k\dfrac{\mathrm{d}q\left(r\right)}{r^{2}}\hat{\boldsymbol{r}}=k_{\text{e}}\mathrm{d}q\left(\boldsymbol{r}\right)\dfrac{\boldsymbol{r}}{\left(\boldsymbol{r}^{2}\right)^{\frac{3}{2}}}\\
= & k_{\text{e}}\mathrm{d}q\left(\boldsymbol{r}\right)\dfrac{\boldsymbol{r}}{\left(\boldsymbol{r}^{2}\right)^{\frac{3}{2}}}=k_{\text{e}}\mathrm{d}q\left(\boldsymbol{r}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }
\end{aligned}
$$

$$
\mathrm{d}\boldsymbol{E}\left(\boldsymbol{r}\right)=-k_{\text{e}}\mathrm{d}q\left(\boldsymbol{r}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }
$$

$$
\mathrm{d}q=\rho\mathrm{d}V\Leftrightarrow\rho=\dfrac{\mathrm{d}q}{\mathrm{d}V}
$$

$$
\mathrm{d}q\left(\boldsymbol{r}\right)=\rho\left(\boldsymbol{r}\right)\mathrm{d}V\Leftrightarrow\rho\left(\boldsymbol{r}\right)=\dfrac{\mathrm{d}q\left(\boldsymbol{r}\right)}{\mathrm{d}V}
$$

$$
q=\int_{q}\mathrm{d}q\left(\boldsymbol{r}\right)=\int_{V}\rho\left(\boldsymbol{r}\right)\mathrm{d}V=\iiint_{V}\rho\left(\boldsymbol{r}\right)\mathrm{d}^{3}\boldsymbol{r}
$$

$$
\begin{aligned}
\boldsymbol{E}= & \iiint_{V}\boldsymbol{E}\left(\boldsymbol{r}\right)\mathrm{d}^{3}\boldsymbol{r}=\iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}\\
= & -k_{\text{e}}\iiint_{V}\left(\rho\left(\boldsymbol{r}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}\\
= & -k_{\text{e}}\iiint_{V}\left(\rho\left(\boldsymbol{r}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{0}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}=\boldsymbol{E}\left(\boldsymbol{0}\right)
\end{aligned}
$$

$$
\begin{aligned}
\Downarrow & \begin{cases}
Q\text{ at }\boldsymbol{r}\\
\mathrm{d}q\text{ or }\rho\text{ at }\boldsymbol{r}^{\prime},\text{ totally }V
\end{cases}\\
\boldsymbol{E}\left(\boldsymbol{r}\right)= & \iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\boldsymbol{\nabla}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}=\iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}\\
\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\cdot\boldsymbol{E}\left(\boldsymbol{r}\right)= & \boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\cdot\iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}=\iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\cdot\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}\\
= & \iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}^{2}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}=\iiint_{V}\left(-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\Delta_{{\scriptscriptstyle \boldsymbol{r}}}\dfrac{1}{\left\Vert \boldsymbol{r}-\boldsymbol{r}^{\prime}\right\Vert }\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime}\\
= & \iiint_{V}f\left(\boldsymbol{r}^{\prime}\right)\Delta\left(\boldsymbol{r}-\boldsymbol{r}^{\prime}\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime},f\left(\boldsymbol{r}^{\prime}\right)=-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\\
= & \iiint_{V}f\left(\boldsymbol{r}^{\prime}\right)\delta\left(\boldsymbol{r}-\boldsymbol{r}^{\prime}\right)\mathrm{d}^{3}\boldsymbol{r}^{\prime},f\left(\boldsymbol{r}^{\prime}\right)=-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}\right)\\
= & -4\pi f\left(\boldsymbol{r}^{\prime}=\boldsymbol{r}\right)=-4\pi\left[-k_{\text{e}}\rho\left(\boldsymbol{r}^{\prime}=\boldsymbol{r}\right)\right]\\
= & -4\pi f\left(\boldsymbol{r}\right)=-4\pi\left[-k_{\text{e}}\rho\left(\boldsymbol{r}\right)\right]=4\pi k_{\text{e}}\rho\left(\boldsymbol{r}\right)\\
\boldsymbol{\nabla}\cdot\boldsymbol{E}\left(\boldsymbol{r}\right)= & 4\pi k_{\text{e}}\rho\left(\boldsymbol{r}\right)
\end{aligned}
$$

$$
\boldsymbol{\nabla}\cdot\boldsymbol{E}\left(\boldsymbol{r}\right)=4\pi k_{\text{e}}\rho\left(\boldsymbol{r}\right)
$$

***

https://www.bilibili.com/video/BV1qu411578c/?t=7m26s

https://en.wikipedia.org/wiki/Helmholtz_equation

Helmholtz equation

$$
\left(\Delta+k^{2}\right)\dfrac{\mathrm{e}^{\pm\mathrm{i}\boldsymbol{k}\cdot\boldsymbol{r}}}{r}=-4\pi\delta\left(\boldsymbol{r}\right)
$$

$$
\left(\boldsymbol{\nabla}_{{\scriptscriptstyle \boldsymbol{r}}}^{2}+\boldsymbol{k}^{2}\right)\dfrac{\mathrm{e}^{\pm\mathrm{i}\boldsymbol{k}\cdot\boldsymbol{r}}}{\left\Vert \boldsymbol{r}\right\Vert }=-4\pi\delta\left(\boldsymbol{r}\right)
$$

### quantum state

https://www.bilibili.com/video/BV1qu411578c/?t=7m45s

### rigorous definition of delta function

$$
\int_{-\infty}^{\infty}\delta\left(x-0\right)\cdot f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

$$
\int\delta\left(x\right)f\left(x\right)\mathrm{d}x=f\left(0\right)
$$

$$
\int\delta\left(x\right)\left(\cdot\right)\mathrm{d}x:f\left(x\right)\rightarrow f\left(0\right)
$$

$$
f\left(x\right)\overset{\int\delta\left(x\right)\left(\cdot\right)\mathrm{d}x}{\rightarrow}f\left(0\right)
$$

$$
f\text{ must be with good properties}
$$

$$
\text{continuous linear functional}\begin{cases}
\text{function-like} & \eta_{{\scriptscriptstyle \varphi}}\left(f\right)=\int\varphi\left(x\right)f\left(x\right)\mathrm{d}x\\
\text{non-function-like},\text{ e.g.} & \delta_{{\scriptscriptstyle x^{\prime}}}\left(f\right)=f\left(x^{\prime}\right)
\end{cases}
$$

general function with "formally" integral notation

$$
\delta_{{\scriptscriptstyle x^{\prime}}}\left(f\right)=\int\delta\left(x-x^{\prime}\right)f\left(x\right)\mathrm{d}x=f\left(x^{\prime}\right)
$$

https://www.bilibili.com/video/BV1qu411578c/?t=11m36s

Quantum mechanics is founded on Hilbert space; Hoever, Dirac delta function is not a linear functional in Hilbert space.?

One of the reason is that we cannot find 1-1 of Dirac delta function, but many-to-one or one-to-many functions.

::: {show-in="html"}

量子力學本不需 Dirac delta function, 其 卻仍大行其道, 實 廣義函數 為其負重前行.

:::

::: {show-in="pdf"}

\begin{CJK}{UTF8}{bsmi}
量子力學本不需 Dirac delta function, 其 卻仍大行其道, 實 廣義函數 為其負重前行.
\end{CJK}

:::

## Fourier expansion or Fourier series

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1m24y1A74K&autoplay=0"></iframe>
:::

https://www.bilibili.com/video/BV1m24y1A74K/?=3m2s

Taylor vs. Fourier

$$
f\left(x\right)=x_{{\scriptscriptstyle 0}}x^{0}+x_{{\scriptscriptstyle 1}}x^{1}+x_{{\scriptscriptstyle 2}}x^{2}+\cdots=\sum\limits _{k=0}^{\infty}x_{{\scriptscriptstyle k}}x^{k},x_{{\scriptscriptstyle k}}=\dfrac{f^{\left(k\right)}\left(0\right)}{k!}
$$

$$
f\left(r\right)=r_{{\scriptscriptstyle 0}}r^{0}+r_{{\scriptscriptstyle 1}}r^{1}+r_{{\scriptscriptstyle 2}}r^{2}+\cdots=\sum\limits _{k=0}^{\infty}r_{{\scriptscriptstyle k}}r^{k},r_{{\scriptscriptstyle k}}=\dfrac{f^{\left(k\right)}\left(0\right)}{k!}
$$

$$
f\left(\theta\right)=a_{{\scriptscriptstyle 0}}\cos\left(0\theta\right)+a_{{\scriptscriptstyle 1}}\cos\left(1\theta\right)+a_{{\scriptscriptstyle 2}}\cos\left(2\theta\right)+\cdots=\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\cos\left(k\theta\right)
$$

$$
a_{{\scriptscriptstyle k}}=\int_{-\infty}^{\infty}\cos\left(k\theta\right)f\left(\theta\right)\mathrm{d}\theta
$$

$$
\begin{aligned}
 & f\left(\theta\right)=a_{{\scriptscriptstyle 0}}\cos\left(0\theta\right)+a_{{\scriptscriptstyle 1}}\cos\left(1\theta\right)+a_{{\scriptscriptstyle 2}}\cos\left(2\theta\right)+\cdots=\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\cos\left(k\theta\right)\\
= & a_{{\scriptscriptstyle 0}}\Re\left[\mathrm{e}^{\mathrm{i}0\theta}\right]+a_{{\scriptscriptstyle 1}}\Re\left[\mathrm{e}^{\mathrm{i}1\theta}\right]+a_{{\scriptscriptstyle 2}}\Re\left[\mathrm{e}^{\mathrm{i}2\theta}\right]+\cdots=\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\Re\left[\mathrm{e}^{\mathrm{i}k\theta}\right]=\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\Re\left[\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\right]\\
= & \Re\left[a_{{\scriptscriptstyle 0}}\mathrm{e}^{\mathrm{i}0\theta}+a_{{\scriptscriptstyle 1}}\mathrm{e}^{\mathrm{i}1\theta}+a_{{\scriptscriptstyle 2}}\mathrm{e}^{\mathrm{i}2\theta}+\cdots\right]=\Re\left[\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\mathrm{e}^{\mathrm{i}k\theta}\right]=\Re\left[\sum\limits _{k=0}^{\infty}a_{{\scriptscriptstyle k}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\right]
\end{aligned}
$$

$$
f\left(x\right)=x_{{\scriptscriptstyle 0}}x^{0}+x_{{\scriptscriptstyle 1}}x^{1}+x_{{\scriptscriptstyle 2}}x^{2}+\cdots=\sum\limits _{k=0}^{\infty}x_{{\scriptscriptstyle k}}x^{k}
$$

$$
f\left(z\right)=z_{{\scriptscriptstyle 0}}z^{0}+z_{{\scriptscriptstyle 1}}z^{1}+z_{{\scriptscriptstyle 2}}z^{2}+\cdots=\sum\limits _{k=0}^{\infty}z_{{\scriptscriptstyle k}}z^{k}
$$

$$
z=r\mathrm{e}^{\mathrm{i}\theta}
$$

$$
\begin{aligned}
f\left(z\right)= & z_{{\scriptscriptstyle 0}}z^{0}+z_{{\scriptscriptstyle 1}}z^{1}+z_{{\scriptscriptstyle 2}}z^{2}+\cdots=\sum\limits _{k=0}^{\infty}z_{{\scriptscriptstyle k}}z^{k}\\
f\left(r\mathrm{e}^{\mathrm{i}\theta}\right)= & z_{{\scriptscriptstyle 0}}\left(r\mathrm{e}^{\mathrm{i}\theta}\right)^{0}+z_{{\scriptscriptstyle 1}}\left(r\mathrm{e}^{\mathrm{i}\theta}\right)^{1}+z_{{\scriptscriptstyle 2}}\left(r\mathrm{e}^{\mathrm{i}\theta}\right)^{2}+\cdots=\sum\limits _{k=0}^{\infty}z_{{\scriptscriptstyle k}}\left(r\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\\
=f\left(r\right)= & z_{{\scriptscriptstyle 0}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{0}r^{0}+z_{{\scriptscriptstyle 1}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{1}r^{1}+z_{{\scriptscriptstyle 2}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{2}r^{2}+\cdots=\sum\limits _{k=0}^{\infty}z_{{\scriptscriptstyle k}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}r^{k}\\
= & r_{{\scriptscriptstyle 0}}r^{0}+r_{{\scriptscriptstyle 1}}r^{1}+r_{{\scriptscriptstyle 2}}r^{2}+\cdots=\sum\limits _{k=0}^{\infty}r_{{\scriptscriptstyle k}}r^{k},r_{{\scriptscriptstyle k}}=z_{{\scriptscriptstyle k}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\\
=f\left(\mathrm{e}^{\mathrm{i}\theta}\right)= & z_{{\scriptscriptstyle 0}}r^{0}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{0}+z_{{\scriptscriptstyle 1}}r^{1}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{1}+z_{{\scriptscriptstyle 2}}r^{2}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{2}+\cdots=\sum\limits _{k=0}^{\infty}z_{{\scriptscriptstyle k}}r^{k}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\\
= & \left(\mathrm{e}^{\mathrm{i}\theta}\right)_{{\scriptscriptstyle 0}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{0}+\left(\mathrm{e}^{\mathrm{i}\theta}\right)_{{\scriptscriptstyle 1}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{1}+\left(\mathrm{e}^{\mathrm{i}\theta}\right)_{{\scriptscriptstyle 2}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{2}+\cdots=\sum\limits _{k=0}^{\infty}\left(\mathrm{e}^{\mathrm{i}\theta}\right)_{{\scriptscriptstyle k}}\left(\mathrm{e}^{\mathrm{i}\theta}\right)^{k}\\
 & ,\left(\mathrm{e}^{\mathrm{i}\theta}\right)_{{\scriptscriptstyle k}}=z_{{\scriptscriptstyle k}}r^{k}
\end{aligned}
$$


## Hilbert space

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1cV411T7hb&autoplay=0"></iframe>
:::

```{definition}
sequence limit
```

$$
\lim_{n\rightarrow\infty}a_{n}=a\Leftrightarrow\forall\epsilon>0,\exists N\in\mathbb{N},\forall n\in\mathbb{N}\left(n>N\Rightarrow\left|a_{n}-a\right|<\epsilon\right)
$$

$$
x_{{\scriptscriptstyle 1}},x_{{\scriptscriptstyle 2}},\cdots,x_{{\scriptscriptstyle n}},x_{{\scriptscriptstyle n+1}},\cdots
$$

- to find convergence
  - monotone convergence theorem
  - split into known sequence
  - fraction: Stolz theorem
  - sequence itself
    - ratio $\dfrac{x_{{\scriptscriptstyle m}}}{x_{{\scriptscriptstyle n}}}$, but not good for some terms being zero 
    - difference $x_{{\scriptscriptstyle m}}-x_{{\scriptscriptstyle n}}$

```{theorem}
Cauchy convergence theorem
```

$$
\begin{array}{c}
\left\{ x_{{\scriptscriptstyle n}}\right\} _{{\scriptscriptstyle n\in\mathbb{N}}}\text{ is convergent}\\
\Updownarrow\\
\forall\epsilon>0,\exists N\in\mathbb{N},\forall m,n\in\mathbb{N}\left(m,n>N\Rightarrow\left|x_{{\scriptscriptstyle m}}-x_{{\scriptscriptstyle n}}\right|<\epsilon\right)\Leftrightarrow\left\{ x_{{\scriptscriptstyle n}}\right\} _{{\scriptscriptstyle n\in\mathbb{N}}}\text{ is a Cauchy sequence}
\end{array}
$$

$$
\text{convergence}\begin{cases}
\left|a_{n}-a\right|<\epsilon & \text{sequence-convergence}\\
\left|x_{{\scriptscriptstyle m}}-x_{{\scriptscriptstyle n}}\right|<\epsilon & \text{Cauchy-convergence}
\end{cases}
$$

$$
\text{convergence}\begin{cases}
\left|a_{n}-a\right|<\epsilon & \text{sequence-convergence}\\
\Downarrow\text{easy proof} & \Uparrow\text{hard proof}\\
\left|x_{{\scriptscriptstyle m}}-x_{{\scriptscriptstyle n}}\right|<\epsilon & \text{Cauchy-convergence}
\end{cases}
$$

## Dirac bracket

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1GN411m7A9&autoplay=0"></iframe>
:::

## quantum operator

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1za4y1F7xo&autoplay=0"></iframe>
:::

## Gram-Schmidt algorithm to find orthonormal basis

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1eu4y1S77C&autoplay=0"></iframe>
:::

## Hermite polynomials and Legendre polynomials

::: {show-in="html"}
<iframe width=500 height=300 frameborder="0" allowfullscreen src="https://player.bilibili.com/player.html?bvid=BV1Lu411V7Rp&autoplay=0"></iframe>
:::